### 2015-07-06 manuscript/ms.tex
adding ms

\documentclass[12pt,preprint]{aastex} 
%\documentclass[apjl]{emulateapj}
\usepackage{amsmath}
\usepackage{color}

\newcommand*{\blue}{\textcolor{blue}}
%\slugcomment{Draft version August 26, 2012}

\shortauthors{Kreidberg}

\begin{document}

\title{BATMAN}

\author{Laura Kreidberg\altaffilmark{1,2}}

\email{E-mail: laura.kreidberg@uchicago.edu}

\altaffiltext{1}{Department of Astronomy and Astrophysics, University of Chicago, 5640 S.~Ellis Ave, Chicago, IL 60637, USA}
\altaffiltext{2}{National Science Foundation Graduate Research Fellow}

\begin{abstract}
BATMAN is the best super-hero.
\end{abstract}

\keywords{planets and satellites: atmospheres --- planets and satellites:  composition --- planets and satellites: individual: WASP-12b}

\section{Introduction}
Previous methods --

Mandel and Agol 2002 - analytic, but uses brute force integration for nonlinear
Giminez 2006 -
Pal 2008 - 
Kipping 2011 -
Pal 2011 - multiple bodies, possibly lacking radial symmetry
	basically same as Mandel and Agol for linear/quadratic; more complex LD not implemented; code available (see discussion)
Abubekerov 
Kjurkchieva - brute force; radially symmetric; 
PyTransit

\section{Algorithm}
The light curve for a planet transiting a star can be described by the function $F(x,y)$, where $F$ is the relative flux of the system when the planet's center is a distance $(x,y)$ away from the stellar center.  $F$ is given by:

\begin{equation}
F = 1 - \int_S{I dS}
\end{equation}
where the stellar intensity $I$ is normalized such that the integrated intensity over the whole stellar surface is unity.

For a radially symmetric limb darkening profile, this surface integral can be reduced to a one-dimensional problem. The intensity is just $I(r)$. Everything in terms of $d = \sqrt{x^2+y^2}$.

There are three cases:\\
case 1: the planet center directly overlaps the stellar center ($d = 0$)\\
case 2: the planet eclipses the star but the centers do not directly overlap ($0<d<1$)\\
case 3: the planet does not eclipse the star ($d>=1$)

First case: planet center overlaps stellar center
\begin{equation}
F(0) = 1 - \int_0^{2\pi}{\int_0^{r_p}{I(r)r dr d\theta}}
\end{equation}

For the second case, where the center of the planet does not overlap the center of the star ($0<d<1$), I make use of the formula for the area of two overlapping circles with radii $r$ and $R$, with centers separated by a distance $d$:

\begin{equation}
A(r, R, d) = r^2\arccos{u} + R^2\arccos{v} - 0.5\sqrt{w}
\end{equation}
where
\begin{eqnarray}
u &=& (d^2+r^2-R^2)/(2dr)\\
v &=& (d^2 + R^2 -r^2)/(2dR) \\
w &=& (-d+r+R)(d+r-R)(d-r+R)(d+r+R)
\end{eqnarray}

Using this expression for the overlapping area, I approximate the integral in FIXME as
\begin{equation}
F(d) = 1 - \sum_{i=0}^{n-1}{\bar{I}(r_i, r_{i+1}) [A(r_{i+1}, r_p, d) - A(r_i, r_p, d)]}
\end{equation}

where $r_0 = d - r_p$ and $r_n = \textrm{MIN}(d+r_p, 1)$.  Works for negative $r$ because $\mu$ is still positive.

The area element is illustrated in Figure FIXME.

To make the sum more efficient, I make the distance $r_i+1 - r_i \propto \arccos{r_i}$  and use Simpson's rule to compute intensity over the area.

\section{Performance}

\section{Discussion}

TODO:
check performance of my alg relative to dA = r*dr*dtheta
xx check whether Simpson's rule is actually faster -- IT ISN'T

\acknowledgments
Props to Robin.

\bibliographystyle{apj}
\bibliography{ms.bib}

\end{document}

### 2015-07-09 c_src/_nonlinear_ld.c
more style clean-up

@@ -32,8 +32,8 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args);
 double intensity(double r, double u1, double u2, double u3, double u4, double norm)
 {
 	if(r > 0.99995) r = 0.99995;
-	double sqrtmu = pow(1. - r * r,0.25);
-	return (1. - u1 * (1. - sqrtmu) - u2 * (1. - pow(sqrtmu,2.)) - u3 * (1. - pow(sqrtmu, 3.)) - u4 * (1. - pow(sqrtmu,4.))) / norm; 	
+	double sqrtmu = pow(1. - r*r,0.25);
+	return (1. - u1*(1. - sqrtmu) - u2*(1. - pow(sqrtmu,2.)) - u3*(1. - pow(sqrtmu, 3.)) - u4*(1. - pow(sqrtmu,4.)))/norm; 	
 }
 
 double area(double d, double r, double R)
@@ -41,13 +41,13 @@ double area(double d, double r, double R)
 	/*
 	Returns area of overlapping circles with radii r and R; separated by a distance d
 	*/
-	double arg1 = (d * d + r * r - R * R) / (2. * d * r); 	
-	double arg2 = (d * d + R * R - r * r) / (2. * d * R); 
-	double arg3 = MAX((-d + r + R) * (d + r - R) * (d - r + R) * (d + r + R), 0.);
+	double arg1 = (d*d + r*r - R*R)/(2.*d*r); 	
+	double arg2 = (d*d + R*R - r*r)/(2.*d*R); 
+	double arg3 = MAX((-d + r + R)*(d + r - R)*(d - r + R)*(d + r + R), 0.);
 
-	if(r <= R - d) return M_PI * r * r;						//planet completely overlaps stellar circle
-	else if(r >= R + d) return M_PI * R * R;					//stellar circle completely overlaps planet
-	else return r * r * acos(arg1) + R * R * acos(arg2) - 0.5 * sqrt(arg3);		//partial overlap
+	if(r <= R - d) return M_PI*r*r;						//planet completely overlaps stellar circle
+	else if(r >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
+	else return r*r*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
 }
 
 static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
[…]

### 2015-07-09 c_src/_uniform_ld.c

@@ -40,7 +40,7 @@ static PyObject *_uniform_ld(PyObject *self, PyObject *args)
 	double *f_array = PyArray_DATA(flux);
 	double *z_array = PyArray_DATA(zs);
 
-	if(fabs(p - 0.5)<1.e-3) p = 0.5;
+	if(fabs(p - 0.5) < 1.e-3) p = 0.5;
 
 	#if defined (_OPENMP)
 	omp_set_num_threads(nthreads);
@@ -55,12 +55,12 @@ static PyObject *_uniform_ld(PyObject *self, PyObject *args)
 		
 		if(z >= 1. + p) f_array[i] = 1.;		//no overlap
 		if(p >= 1. && z <= p - 1.) f_array[i] = 0.;	//total eclipse of the star
-		else if(z <= 1. - p) f_array[i] = 1. - p * p;	//planet is fully in transit
+		else if(z <= 1. - p) f_array[i] = 1. - p*p;	//planet is fully in transit
 		else						//planet is crossing the limb
 		{
-			kap1=acos(fmin((1. - p * p + z * z) / 2. / z, 1.));
-			kap0=acos(fmin((p * p + z * z - 1.) / 2. / p / z, 1.));
-			f_array[i] = 1. - (p * p * kap0 + kap1 - 0.5 * sqrt(fmax(4. * z * z - pow(1. + z * z - p * p, 2.), 0.))) / M_PI;
+			kap1=acos(fmin((1. - p*p + z*z)/2./z, 1.));
+			kap0=acos(fmin((p*p + z*z - 1.)/2./p/z, 1.));
+			f_array[i] = 1. - (p*p*kap0 + kap1 - 0.5*sqrt(fmax(4.*z*z - pow(1. + z*z - p*p, 2.), 0.)))/M_PI;
 		}
 	}
 

### 2015-07-09 batman/plots.py
more style clean up

@@ -19,6 +19,8 @@ import numpy as np
 import math
 import matplotlib.pyplot as plt
 from .transitmodel import *
+from . import _quadratic_ld
+from . import _nonlinear_ld
 import timeit
 
 def wrapper(func, *args, **kwargs):
@@ -30,9 +32,14 @@ def wrapper(func, *args, **kwargs):
 def make_plots():
 	"""zs = np.linspace(0., 1., 1000)
 	rp = 0.1
+	wrapped = wrapper(_quadratic_ld._quadratic_ld, zs, rp, 0.1, 0.3, 1)
+	t = timeit.timeit(wrapped,number=10000)
+	print("time:", t)"""
+	"""zs = np.linspace(0., 1., 1000)
+	rp = 0.1
 	u = [0., 0.7, 0.0, -0.3]
-	f = occultnl.occultnl(zs, rp, u[0], u[1], u[2], u[3], 1.0e-2, 4)
-	fhi = occultnl.occultnl(zs, rp, u[0], u[1], u[2], u[3], 1.0e-4, 4)
+	f = _nonlinear_ld._nonlinear_ld(zs, rp, u[0], u[1], u[2], u[3], 1.0e-2, 4)
+	fhi = _nonlinear_ld._nonlinear_ld(zs, rp, u[0], u[1], u[2], u[3], 1.0e-4, 4)
 	fquad = occultquad.occultquad(zs, rp, 0.1, 0.3, 4)
 	#for i in range(len(f)): print "z, fnl, fquad", zs[i], f[i], fquad[i]
 
@@ -48,20 +55,20 @@ def make_plots():
 
 
 	#generates Figure FIXME: max err as a function of function call time
-	"""zs = np.linspace(0., 1., 1000)
+	zs = np.linspace(0., 1., 1000)
 	rp = 0.1
 	u = [0., 0.7, 0.0, -0.3]
 	n = 20
 	ts = []
 	errs = []
-	f_ref = occultnl.occultnl(zs, rp, u[0], u[1], u[2], u[3], 1.0e-4, 4)
+	f_ref = _nonlinear_ld._nonlinear_ld(zs, rp, u[0], u[1], u[2], u[3], 1.0e-4, 4)
 	fac = np.logspace(-3, -1, n)
 	for i in range(n):
-		wrapped = wrapper(occultnl.occultnl, zs, rp, u[0], u[1], u[2], u[3], fac[i], 12)
[…]

### 2015-07-09 batman/transitmodel.py

@@ -79,14 +79,14 @@ class TransitModel:
 		else:
 			if nthreads <= multiprocessing.cpu_count()and nthreads >1 and openmp.detect(): self.nthreads = nthreads
 			else: 
-				if nthreads>multiprocessing.cpu_count(): raise Exception("Maximum number of threads is "+'{0:d}'.format(multiprocessing.cpu_count()))
+				if nthreads > multiprocessing.cpu_count(): raise Exception("Maximum number of threads is "+'{0:d}'.format(multiprocessing.cpu_count()))
 				elif nthreads <= 1: raise Exception("Number of threads must be between 2 and {0:d}".format(multiprocessing.cpu_count()))
 				else: raise Exception("OpenMP not enabled: do not set the nthreads parameter")
 
 	def calc_err(self, plot = False):
 		"""
 
-		Returns maximum error (in parts per million) for transit light curve calculation.
+		Calculate maximum error (in parts per million) for transit light curve calculation.
 			
 		:param plot: 
 			If set to ``True``, plots the error in the light curve model as a function of separation of centers.
@@ -130,13 +130,13 @@ class TransitModel:
 				if err> self.max_err: fac_hi = fac	
 				else: fac_lo = fac
 				n += 1
-				if n>1e3: raise Exception("Convergence failure in calculation of scale factor for _nonlinear_ld")
+				if n > 1e3: raise Exception("Convergence failure in calculation of scale factor for _nonlinear_ld")
 			return fac
 		else: return 0.
 
 	def LightCurve(self, params):
 		"""
-		Calculates a model light curve.
+		Return a model light curve.
 
 		:param params:
 			Transit parameter object.
@@ -167,7 +167,7 @@ class TransitModel:
 
 class TransitParams(object):
 	"""
-	An object that stores the physical parameters of the transit.
+	Object to store the physical parameters of the transit.
 
 	:param t0:
 		Time of periastron passage (for eccentric orbits) or time of central transit (for circular orbits).

### 2015-07-10 manuscript/Figures/fig1.py
update figure 1

@@ -2,7 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 import math
 
-plt.figure(figsize = (4,4))
+plt.figure(figsize = (6,6))
 theta = np.linspace(0, 2*math.pi, 1000)
 
 r = 0.15
@@ -27,13 +27,16 @@ plt.plot(x,y, color='k')
 r1 = rs[5]
 r2 = rs[6]
 
-xb = x[(x>=r1)&(x<=r2)&(y<0.)] 
-yb = y[(x>=r1)&(x<=r2)&(y<0.)] 
-xt = x[(x>=r1)&(x<=r2)&(y>=0.)] 
-yt = y[(x>=r1)&(x<=r2)&(y>=0.)] 
+xmin = (d**2 + r1**2 - r**2)/(2.*d)
+xmax = (d**2 + r2**2 - r**2)/(2.*d)
+
+xb = x[(x>=xmin)&(x<xmax)&(y<0.)] 
+yb = y[(x>=xmin)&(x<xmax)&(y<0.)] 
+xt = x[(x>=xmin)&(x<xmax)&(y>0.)] 
+yt = y[(x>=xmin)&(x<xmax)&(y>0.)] 
 xl = r1*np.cos(theta)
 yl = r1*np.sin(theta)
-ind = (yl<=yt[0])&(yl>=yb[0])&(xl>=0.)
+ind = (yl<yt[0])&(yl>yb[0])&(xl>0.)
 xl = xl[ind]
 yl = yl[ind]
 ind0 = np.argsort(yl)
@@ -53,6 +56,14 @@ ind2 = np.argsort(yr)
 yr = yr[ind2]
 xr= xr[ind2]
 
+"""plt.clf()
+plt.plot(xb, yb, label="b")
+plt.plot(xr, yr, label="r")
+plt.plot(xt, yt, label="t")
+plt.plot(xl, yl, label="l")
+plt.legend()
+plt.show()"""
+
 xfill = np.append(xb,xr)
 yfill = np.append(yb,yr)
 xfill = np.append(xfill, xt)
@@ -60,7 +71,7 @@ yfill = np.append(yfill, yt)
 xfill = np.append(xfill, xl)
 yfill = np.append(yfill, yl)
 
-for i in range(len(xfill)): print(xfill[i], yfill[i])
+#for i in range(len(xfill)): print(xfill[i], yfill[i])
 #for i in range(len(xr)): print(xr[i], yr[i])
 
[…]

### 2015-07-12 manuscript/ms.tex
draft of section 2

@@
 \usepackage{color}
+%\usepackage{hyperref}
 
 \newcommand*{\blue}{\textcolor{blue}}
-%\slugcomment{Draft version August 26, 2012}
 
@@
 \section{Algorithm}
-The light curve for a planet transiting a star can be described by the function $F(x,y)$, where $F$ is the relative flux of the system when the planet's center is a distance $(x,y)$ away from the stellar center.
-$F$ is given by:
+\label{section:algorithm}
+To calculate the fraction $\delta$ of stellar flux blocked by a transiting planet, one must integrate the sky-projected intensity of the star ($I$) over the area obscured by the disk of the planet ($S$):
+\begin{equation}
+\delta = \iint_S{I \mathrm{d}S}
+\label{eqn:surfaceint}
+\end{equation}
+where $I$ is normalized such that the integrated intensity over the stellar disk is unity.
+This expression is valid for any general stellar surface brightness map; however, it is slow to evaluate numerically because the differential area elements must be small ($<10^{-6}$) in order to achieve better than one part per million (ppm) accuracy.
+
+On the other hand, if the stellar intensity profile is radially symmetric, the two-dimensional calculation in Equation\,\ref{eqn:surfaceint} can be reduced to one dimension and sped up greatly with the following algorithm:
+\begin{equation}
+\delta = \sum_{i=1}^{n} I\left(\frac{x_i+x_{i-1}}{2}\right) \left[A(x_{i}, r_p, d) - A(x_{i-1}, r_p, d)\right]
+\end{equation}
+where $I(x)$ is the 1D stellar intensity profile, $r_p$ is the planetary radius, $d$ is the separation of centers between the star and the planet, and $A(x, r_p, d)$ is the area of intersection between two circles of radii $x$ and $r_p$, separated by a distance $d$.
+The sum is carried out over the range $x_0 = d - r_p$ to $x_n = \textrm{MIN}(d + r_p, r_s)$, where $r_s$ is the stellar radius.
+The intersecting area is given by:
 
 \begin{equation}
-F = 1 - \int_S{I dS}
-\end{equation}
-where the stellar intensity $I$ is normalized such that the integrated intensity over the whole stellar surface is unity.
-
-For a radially symmetric limb darkening profile, this surface integral can be reduced to a one-dimensional problem. The intensity is just $I(r)$. Everything in terms of $d = \sqrt{x^2+y^2}$.
-
-There are three cases:\\
-case 1: the planet center directly overlaps the stellar center ($d = 0$)\\
-case 2: the planet eclipses the star but the centers do not directly overlap ($0<d<1$)\\
-case 3: the planet does not eclipse the star ($d>=1$)
-
-First case: planet center overlaps stellar center
-\begin{equation}
-F(0) = 1 - \int_0^{2\pi}{\int_0^{r_p}{I(r)r dr d\theta}}
-\end{equation}
-
-For the second case, where the center of the planet does not overlap the center of the star ($0<d<1$), I make use of the formula for the area of two overlapping circles with radii $r$ and $R$, with centers separated by a distance $d$:
-
-\begin{equation}
-A(r, R, d) = r^2\arccos{u} + R^2\arccos{v} - 0.5\sqrt{w}
+A(x, r_p, d) = 
+\begin{cases}
+x^2\cos^{-1}{u} + r_p^2\cos^{-1}{v} - 0.5\sqrt{w}, & r_p -d < x < r_p + d\\
+x^2, & x \le r_p - d\\
+r_p^2 & x \ge r_p +d \\
+\end{cases}
 \end{equation}
@@
 \begin{eqnarray}
-u &=& (d^2+r^2-R^2)/(2dr)\\
-v &=& (d^2 + R^2 -r^2)/(2dR) \\
-w &=& (-d+r+R)(d+r-R)(d-r+R)(d+r+R)
+u &=& (d^2+x^2-r_p^2)/(2dx)\\
+v &=& (d^2 + r_p^2 -x^2)/(2dr_p) \\
+w &=& (-d+x+r_p)(d+x-r_p)(d-x+r_p)(d+x+r_p)
 \end{eqnarray}
 
-Using this expression for the overlapping area, I approximate the integral in FIXME as
-\begin{equation}
-F(d) = 1 - \sum_{i=0}^{n-1}{\bar{I}(r_i, r_{i+1}) [A(r_{i+1}, r_p, d) - A(r_i, r_p, d)]}
-\end{equation}
+The advantage of this integration scheme is that the only error introduced is due to approximating the stellar intensity as constant over the differential area element $\Delta A = A(x_i, r_p, d) - A(x_{i-1}, r_p, d)$.
+Computation of $\Delta A$ is expensive, but it needs to be calculated relatively few times ($<<10^{6}$ for sub-ppm accuracy).
+This makes the integration faster than one with a simpler area element (e.g., $\Delta A = \Delta x \Delta y$), that requires a much smaller step size to achieve the same accuracy.
 
-where $r_0 = d - r_p$ and $r_n = \textrm{MIN}(d+r_p, 1)$.  Works for negative $r$ because $\mu$ is still positive.
+To further optimize the integration, I use a nonuniform step size.
+Typical stellar intensity profiles have larger gradients near the limb of the star than at the center (FIXME ref), so I choose a smaller steps at larger $x$ values: $x_i - x_{i-1} \propto \cos^{-1}{x_{i-1}}$.
+An alternative choice would be to make the step-size proportional to the intensity gradient; however, this is more expensive to compute and does not save time??.
 
-The area element is illustrated in Figure FIXME.
+See Figure\,\ref{fig:integration} for a schematic illustrating the geometry of the integration.
 
-To make the sum more efficient, I make the distance $r_i+1 - r_i \propto \arccos{r_i}$  and use Simpson's rule to compute intensity over the area.
+\begin{figure}
+\resizebox{\hsize}{!}{\includegraphics{f1.pdf}}
+\caption{Schematic illustration of the integration scheme.
+The star (large black circle; partially visible) has a radius of 1 and is centered in the plane of the sky at ($x$, $y$) = (0, 0).
+The planet (smaller black circle) is separated from the center of the stellar disk by a distance $d$ (marked by the solid black line).
+The star is partitioned into concentric circles (dotted lines) in order to calculate the integral over the planet disk.
+A single integration element $\Delta A$ is shaded in orange.
+The integration step size illustrated here is larger than for a typical calculation for visual clarity.}
+\label{fig:integration}
+\end{figure}
 
-\section{Performance}
+\section{The \texttt{batman} package}
+I wrote a Python package, \texttt{batman}, that implements the algorithm described in \S\,\ref{section:algorithm} to calculate model transit light curves.
+\texttt{batman} is being developed on GitHub and full documentation is available at \blue{\url{https://github.com/lkreidberg/batman}}.
+
+In brief, \texttt{batman} supports calculation of light curves for the following stellar intensity profiles:
+
+\begin{align}
+  I(\mu) &= I_0                                                                         & &\text{(uniform)}             \\
+  I(\mu) &= I_0[1 - u_1(1-\mu)]                                                         & &\text{(linear)}              \\
+  I(\mu) &= I_0[1 - u_1(1 - \mu) - u_2(1-\mu)^2]                                        & &\text{(quadratic)}           \\
+  I(\mu) &= I_0[1 - u_1(1-\mu^{1/2}) - u_2(1- \mu) - u_3(1-\mu^{3/2}) - u_4(1-\mu^2)]   & &\text{(nonlinear)}
+\end{align}
+
+as well as the option to define a custom profile.
+
+The code calculates the separation of centers between the star and the planet based on physical parameters of the transit (e.g., planet-to-star radius ratio, planet orbital period).
+(FIXME add ref to winn)
+
+-parallelized with OpenMP with a user-specified number of processors (automatically detects whether supported and fails gracefully)
+- compatible with Python 2-3.
+- listed on PyPI
+
+\subsection{Performance}
+\subsection{Comparison with other codes}
 
 \section{Discussion}
-
-TODO:
-check performance of my alg relative to dA = r*dr*dtheta
-xx check whether Simpson's rule is actually faster -- IT ISN'T
 

### 2015-07-12 manuscript/Figures/fig1.py

@@ -1,12 +1,15 @@
 import matplotlib.pyplot as plt
 import numpy as np
 import math
+from pylab import *
+
+rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
 
 plt.figure(figsize = (6,6))
 theta = np.linspace(0, 2*math.pi, 1000)
 
 r = 0.15
-fac = 7.0e-2 
+fac = 7.7e-2 
 rs = []
 
 while r < 1.0: 
@@ -23,6 +26,7 @@ d = 0.64
 x = r*np.cos(theta) + d
 y = r*np.sin(theta)
 plt.plot(x,y, color='k')
+
 
 r1 = rs[5]
 r2 = rs[6]
@@ -77,10 +81,27 @@ yfill = np.append(yfill, yl)
 plt.fill(xfill,yfill, color='orange')
 
 
-plt.xlim((0., 1.0))
-plt.ylim((-0.5, 0.5))
-plt.axhline(0., color='0.5', zorder=5, linestyle = 'dashed')
+r = 1.0
+x = r*np.cos(theta)
+y = r*np.sin(theta)
+
+plt.plot(x,y, color='k')
+
+plt.xlim((-0.06, 1.0))
+plt.ylim((-0.53, 0.53))
+
+plt.axhline(0., color='k', zorder=5, linestyle = 'dashed')
+plt.axvline(0., color='k', zorder=5, linestyle = 'dashed')
+
+plt.plot([0., d],[0., 0.], color='k', linewidth=1.5, zorder=6)
+plt.plot(0., 0., ms= 5, marker='o', color='k')
+plt.plot(d, 0., ms= 5, marker='o', color='k')
+plt.gca().text(0.3,0.04, "d", fontsize=16, weight='bold')
+#plt.gca().annotate("", xy = (0.05, 0.03), xycoords='data', xytext =  (0.6, 0.03), arrowprops = dict(arrowstyle="->", fc="0.4", ec="0.4"))
+#plt.gca().annotate("", xy = (0.05, 0.03), xycoords='data', xytext =  (0.6, 0.03), arrowprops = dict(arrowstyle="<-", fc="0.4", ec = "0.4"))
+
 plt.xlabel("x (stellar radii)")
 plt.ylabel("y (stellar radii)")
 
+plt.savefig("f1.pdf", dpi=300)
 plt.show()

### 2015-07-14 batman/plots.py
new diagnostic plots

@@ -22,6 +22,7 @@ from .transitmodel import *
 from . import _quadratic_ld
 from . import _nonlinear_ld
 import timeit
+import batman
 
 def wrapper(func, *args, **kwargs):
     def wrapped():

### 2015-07-14 batman/transitmodel.py
adding _set_fac()

@@ -133,6 +133,9 @@ class TransitModel:
 				if n > 1e3: raise Exception("Convergence failure in calculation of scale factor for _nonlinear_ld")
 			return fac
 		else: return 0.
+	
+	def _set_fac(self, fac):
+		self.fac = fac
 
 	def LightCurve(self, params):
 		"""
@@ -199,12 +202,12 @@ class TransitParams(object):
 
 	"""
 	def __init__(self):
-		self.t0 = 0.
-		self.per = 0.
-		self.rp = 0.
-		self.a = 0.
-		self.inc = 0.
-		self.ecc = 0.
-		self.w = 0. 
-		self.u = []
-		self.limb_dark = ""
+		self.t0 = None
+		self.per = None
+		self.rp = None
+		self.a = None
+		self.inc = None
+		self.ecc = None
+		self.w = None
+		self.u = None
+		self.limb_dark = None

### 2015-07-14 c_src/_nonlinear_ld.c
change dr for final step

@@ -98,10 +98,11 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 		else
 		{
 			delta = 0.;						//variable to store the integrated intensity, \int I dA
-
 			r = r_in;						//starting radius for integration
 			dr = fac*acos(r); 					//initial step size 
+
 			r += dr;						//first step
+
 			A_i = 0.;						//initial area
 	
 			while(r < r_out)
@@ -113,7 +114,7 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 				r = r + dr;					//stepping to next element
 				A_i = A_f;					//storing area
 			}
-			dr = r_out - r + dr;  					//calculating change in radius for last step
+			dr = r_out - r - dr;  					//calculating change in radius for last step  FIXME
 			r = r_out;						//final radius for integration
 			A_f = area(d, r, rprs);					//area for last integration step
 			I = intensity(r - dr/2.,u1,u2, u3, u4, norm); 		//intensity at the midpoint 

### 2015-07-14 manuscript/figs/fig1.py
adding fig 2

@@ -1,0 +1,107 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import math
+from pylab import *
+
+rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
+
+plt.figure(figsize = (6,6))
+theta = np.linspace(0, 2*math.pi, 1000)
+
+r = 0.15
+fac = 7.7e-2 
+rs = []
+
+while r < 1.0: 
+	rs.append(r)
+	x = r*np.cos(theta)
+	y = r*np.sin(theta)
+
+	dr = fac*math.acos(r)
+	r = r + dr
+	plt.plot(x,y, color='k', linestyle='dotted')
+
+r = 0.15
+d = 0.64
+x = r*np.cos(theta) + d
+y = r*np.sin(theta)
+plt.plot(x,y, color='k')
+
+
+r1 = rs[5]
+r2 = rs[6]
+
+xmin = (d**2 + r1**2 - r**2)/(2.*d)
+xmax = (d**2 + r2**2 - r**2)/(2.*d)
+
+xb = x[(x>=xmin)&(x<xmax)&(y<0.)] 
+yb = y[(x>=xmin)&(x<xmax)&(y<0.)] 
+xt = x[(x>=xmin)&(x<xmax)&(y>0.)] 
+yt = y[(x>=xmin)&(x<xmax)&(y>0.)] 
+xl = r1*np.cos(theta)
+yl = r1*np.sin(theta)
+ind = (yl<yt[0])&(yl>yb[0])&(xl>0.)
+xl = xl[ind]
+yl = yl[ind]
+ind0 = np.argsort(yl)
+yl = yl[ind0]
+xl = xl[ind0]
+ind4 = np.argsort(yl)
+ind4 = ind4[::-1]
+yl = yl[ind4]
+xl = xl[ind4]
+
+xr = r2*np.cos(theta)
+yr = r2*np.sin(theta)
+ind1 = (yr<yt[-1])&(yr>yb[-1])&(xr>0.)
+xr = xr[ind1]
+yr = yr[ind1]
+ind2 = np.argsort(yr)
+yr = yr[ind2]
+xr= xr[ind2]
+
+"""plt.clf()
+plt.plot(xb, yb, label="b")
+plt.plot(xr, yr, label="r")
+plt.plot(xt, yt, label="t")
+plt.plot(xl, yl, label="l")
+plt.legend()
+plt.show()"""
+
+xfill = np.append(xb,xr)
[…]

### 2015-07-14 manuscript/figs/fig2.py

@@ -1,0 +1,51 @@
+import batman
+import numpy as np
+import matplotlib.pyplot as plt
+from pylab import *
+from matplotlib import rc
+import matplotlib.gridspec as gridspec
+
+rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
+matplotlib.rcParams.update({'font.size':14})
+
+params = batman.TransitParams()	        #object to store transit parameters
+params.t0 = 0. 				#time of periastron passage (for eccentric orbits), OR
+					#mid-transit time (for circular orbits)
+params.per = 1.58			#orbital period	
+params.rp = 0.1				#planet radius (in units of stellar radii)
+params.a = 15.				#semi-major axis (in units of stellar radii)
+params.inc = 87.			#orbital inclination (in degrees)	
+params.ecc = 0.				#eccentricity	
+params.w = 90.				#longitude of periastron (in degrees)
+params.limb_dark = "nonlinear"          #limb darkening model
+params.u = [0.5, 0.1, 0.1, -0.1]       	#limb darkening coefficients
+   
+t = np.linspace(-0.025, 0.025, 1000)  	#times at which to calculate light curve	
+m = batman.TransitModel(params, t)      #initializes model
+
+flux = m.LightCurve(params)		#calculates light curve
+
+m._set_fac(1.0e-4)
+flux_precise = m.LightCurve(params)
+
+gs = gridspec.GridSpec(2, 1, height_ratios =[3,1], hspace=0.05)
+
+ax = plt.subplot(gs[0, 0])
+ax.plot(t*24., flux, color='k')
+ax.set_xticks([])
+plt.ylabel("Relative flux")
+plt.ylim((0.989, 1.001))
+plt.xlim((-0.5, 0.5))
+
+ax2 = plt.subplot(gs[1,0])
+ax2.plot(t*24., (flux - flux_precise)*1.0e6, color='k')
+plt.xlabel("Time from central transit (hours)")
+plt.ylabel("Error (ppm)")
+plt.xlim((-0.5, 0.5))
+plt.ylim((-1.1,0.1))
+
+plt.show()
+
+
+plt.savefig("f2.pdf", dpi=300)
+

### 2015-07-14 manuscript/Figures/fig1.py
changing dir name

@@ -1,107 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import math
-from pylab import *
-
-rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
-
-plt.figure(figsize = (6,6))
-theta = np.linspace(0, 2*math.pi, 1000)
-
-r = 0.15
-fac = 7.7e-2 
-rs = []
-
-while r < 1.0: 
-	rs.append(r)
-	x = r*np.cos(theta)
-	y = r*np.sin(theta)
-
-	dr = fac*math.acos(r)
-	r = r + dr
-	plt.plot(x,y, color='k', linestyle='dotted')
-
-r = 0.15
-d = 0.64
-x = r*np.cos(theta) + d
-y = r*np.sin(theta)
-plt.plot(x,y, color='k')
-
-
-r1 = rs[5]
-r2 = rs[6]
-
-xmin = (d**2 + r1**2 - r**2)/(2.*d)
-xmax = (d**2 + r2**2 - r**2)/(2.*d)
-
-xb = x[(x>=xmin)&(x<xmax)&(y<0.)] 
-yb = y[(x>=xmin)&(x<xmax)&(y<0.)] 
-xt = x[(x>=xmin)&(x<xmax)&(y>0.)] 
-yt = y[(x>=xmin)&(x<xmax)&(y>0.)] 
-xl = r1*np.cos(theta)
-yl = r1*np.sin(theta)
-ind = (yl<yt[0])&(yl>yb[0])&(xl>0.)
-xl = xl[ind]
-yl = yl[ind]
-ind0 = np.argsort(yl)
-yl = yl[ind0]
-xl = xl[ind0]
-ind4 = np.argsort(yl)
-ind4 = ind4[::-1]
-yl = yl[ind4]
-xl = xl[ind4]
-
-xr = r2*np.cos(theta)
-yr = r2*np.sin(theta)
-ind1 = (yr<yt[-1])&(yr>yb[-1])&(xr>0.)
-xr = xr[ind1]
-yr = yr[ind1]
-ind2 = np.argsort(yr)
-yr = yr[ind2]
-xr= xr[ind2]
-
-"""plt.clf()
-plt.plot(xb, yb, label="b")
-plt.plot(xr, yr, label="r")
-plt.plot(xt, yt, label="t")
-plt.plot(xl, yl, label="l")
-plt.legend()
-plt.show()"""
-
-xfill = np.append(xb,xr)
[…]

### 2015-07-14 manuscript/figs/fig2.py
add fig 2

@@ -5,6 +5,8 @@ from pylab import *
 from matplotlib import rc
 import matplotlib.gridspec as gridspec
 
+print(batman.__file__)
+
 rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
 matplotlib.rcParams.update({'font.size':14})
 
@@ -44,8 +46,6 @@ plt.ylabel("Error (ppm)")
 plt.xlim((-0.5, 0.5))
 plt.ylim((-1.1,0.1))
 
-plt.show()
-
-
 plt.savefig("f2.pdf", dpi=300)
+plt.show()
 

### 2015-07-16 manuscript/ms.tex
add fig 2
adding section 3

@@
 \end{equation}
-where $I(x)$ is the 1D stellar intensity profile, $r_p$ is the planetary radius, $d$ is the separation of centers between the star and the planet, and $A(x, r_p, d)$ is the area of intersection between two circles of radii $x$ and $r_p$, separated by a distance $d$.
+where $x$ is the normalized radial coordinate $0 < x < 1$, $I(x)$ is the 1D stellar intensity profile, $r_p$ is the planetary radius, $d$ is the separation of centers between the star and the planet, and $A(x, r_p, d)$ is the area of intersection between two circles of radii $x$ and $r_p$, separated by a distance $d$.
 The sum is carried out over the range $x_0 = d - r_p$ to $x_n = \textrm{MIN}(d + r_p, r_s)$, where $r_s$ is the stellar radius.
@@
 v &=& (d^2 + r_p^2 -x^2)/(2dr_p) \\
-w &=& (-d+x+r_p)(d+x-r_p)(d-x+r_p)(d+x+r_p)
+w &=& (-d+x+r_p)(d+x-r_p)(d-x+r_p)(d+x+r_p).
 \end{eqnarray}
@@
 To further optimize the integration, I use a nonuniform step size.
-Typical stellar intensity profiles have larger gradients near the limb of the star than at the center (FIXME ref), so I choose a smaller steps at larger $x$ values: $x_i - x_{i-1} \propto \cos^{-1}{x_{i-1}}$.
-An alternative choice would be to make the step-size proportional to the intensity gradient; however, this is more expensive to compute and does not save time??.
+Typical stellar intensity profiles have larger gradients near the limb of the star than at the center \citep[e.g.][]{claret00}, so smaller steps are required at larger $x$ values to achieve the same accuracy.
+I adopt the following step-size scaling:
+$$
+x_i - x_{i-1} \propto \cos^{-1}\left(x_{i-1}\right)
+$$  
+which is fast to compute and well-behaved at the limits $x=0$ and $x=1$.
 
@@
 
+\begin{figure}
+\resizebox{\hsize}{!}{\includegraphics{f2.pdf}}
+\caption{An example transit light curve for a nonlinear stellar intensity profile (top panel) and the error in the calculation (bottom panel).
+The error tolerance parameter \texttt{max\_err} was set to 1.0 parts per million.
+The error increases with distance from the center of the star up to around $\pm0.2$ hours from the time of mid-transit, because the stellar intensity gradient is larger at larger radii.
+The error decreases again during ingress and egress as the planet eclipses a smaller fraction of the stellar disk.}
+\label{fig:transit}
+\end{figure}
+
 \section{The \texttt{batman} package}
-I wrote a Python package, \texttt{batman}, that implements the algorithm described in \S\,\ref{section:algorithm} to calculate model transit light curves.
-\texttt{batman} is being developed on GitHub and full documentation is available at \blue{\url{https://github.com/lkreidberg/batman}}.
+I wrote a Python package, \texttt{batman}, to calculate model transit light curves.
+\texttt{batman} is an open source project and is being developed on GitHub.
+Full documentation is available at \blue{\url{https://github.com/lkreidberg/batman}}, but I summarize the main capabilities of the package here.
 
-In brief, \texttt{batman} supports calculation of light curves for the following stellar intensity profiles:
-
+\subsection{Features}
+\texttt{batman} supports calculation of exoplanet transit light curves for uniform, linear, quadratic, nonlinear, and custom stellar intensity profiles:
 \begin{align}
@@
 \end{align}
+where $\mu = \sqrt{1-x^2}$.
+A template is available in the source distribution for the creation of custom limb darkening profiles for any radially symmetric function.
+The nonlinear and custom models are calculated with the brute force integration scheme described in \S\,\ref{section:algorithm}, whereas models for quadratic and linear profiles are calculated analytically based on the widely-used formalism from \cite{mandel02}.
 
-as well as the option to define a custom profile.
+For models calculated by integration, the user can specify the maximum allowable error with the \texttt{max\_err} parameter.
+When a transit model is initialized, the integration step size is tuned such that the difference between the model and a template calculated with a fine step size (FIXME) is below the specified tolerance.
+Computation time increases FIXME with precision (see FIXME), so .
+FIXME lower limit on precision.
+Figure\,\ref{fig:transit} shows an example transit light curve with errors FIXME.
 
-The code calculates the separation of centers between the star and the planet based on physical parameters of the transit (e.g., planet-to-star radius ratio, planet orbital period).
-(FIXME add ref to winn)
+FIXME this needs work.
+One of the free parameters in the transit models is the separation of centers $d$ between the star and the planet.
+\texttt{batman} computes $d$ automatically from the following set of user-specified parameters: planet semi-major axis, orbital inclination, eccentricity, longitude of periastron, orbital period, and time of periastron/mid-transit time (for eccentric/circular orbits).
+The separation $d$ is calculated with equation (5) from Chapter FIXME and FIXME of \citep{seager10}.
 
--parallelized with OpenMP with a user-specified number of processors (automatically detects whether supported and fails gracefully)
-- compatible with Python 2-3.
-- listed on PyPI
+The computationally intensive sections of the code are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
+For all limb darkening models, the calculation can be parallelized at the C level with OpenMP.
+The number of processors is specified by the user with the parameter \texttt{nthreads}.
+\texttt{batman} will raise an exception if users attempt to parallelize a calculation on a system where OpenMP is not supported.
 

### 2015-07-16 batman/transitmodel.py
add set_fac method

@@ -134,7 +134,13 @@ class TransitModel:
 			return fac
 		else: return 0.
 	
-	def _set_fac(self, fac):
+	def set_fac(self, fac):
+		"""
+		Set scale factor for integration step size.
+
+		:params fac:
+			Scale factor.
+		"""
 		self.fac = fac
 
 	def LightCurve(self, params):

### 2015-07-16 c_src/_nonlinear_ld.c
revert to original calculation of dr

@@ -114,7 +114,7 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 				r = r + dr;					//stepping to next element
 				A_i = A_f;					//storing area
 			}
-			dr = r_out - r - dr;  					//calculating change in radius for last step  FIXME
+			dr = r_out - r + dr;  					//calculating change in radius for last step  FIXME
 			r = r_out;						//final radius for integration
 			A_f = area(d, r, rprs);					//area for last integration step
 			I = intensity(r - dr/2.,u1,u2, u3, u4, norm); 		//intensity at the midpoint 

### 2015-07-16 manuscript/ms.tex
add fig  3

@@
 \begin{abstract}
-BATMAN is the best super-hero.
+Batman -- the best super-hero.
 \end{abstract}
@@
 \section{Introduction}
-Previous methods --
+%Previous methods --
 
-Mandel and Agol 2002 - analytic, but uses brute force integration for nonlinear
-Giminez 2006 -
-Pal 2008 - 
-Kipping 2011 -
-Pal 2011 - multiple bodies, possibly lacking radial symmetry
-	basically same as Mandel and Agol for linear/quadratic; more complex LD not implemented; code available (see discussion)
-Abubekerov 
-Kjurkchieva - brute force; radially symmetric; 
-PyTransit
+%Mandel and Agol 2002 - analytic, but uses brute force integration for nonlinear
+%Giminez 2006 -
+%Pal 2008 - 
+%Kipping 2011 -
+%Pal 2011 - multiple bodies, possibly lacking radial symmetry
+%	basically same as Mandel and Agol for linear/quadratic; more complex LD not implemented; code available (see discussion)
+%Abubekerov 
+%Kjurkchieva - brute force; radially symmetric; 
+%PyTransit
 
@@
 $$
-x_i - x_{i-1} \propto \cos^{-1}\left(x_{i-1}\right)
+x_i - x_{i-1}  = f\cos^{-1}\left(x_{i-1}\right)
 $$  
-which is fast to compute and well-behaved at the limits $x=0$ and $x=1$.
+where $f$ is a constant scale factor. This prescription is fast to compute and well-behaved at the limits $x=0$ and $x=1$.
 
@@
 \label{fig:integration}
-\end{figure}
-
-\begin{figure}
-\resizebox{\hsize}{!}{\includegraphics{f2.pdf}}
-\caption{An example transit light curve for a nonlinear stellar intensity profile (top panel) and the error in the calculation (bottom panel).
-The error tolerance parameter \texttt{max\_err} was set to 1.0 parts per million.
-The error increases with distance from the center of the star up to around $\pm0.2$ hours from the time of mid-transit, because the stellar intensity gradient is larger at larger radii.
-The error decreases again during ingress and egress as the planet eclipses a smaller fraction of the stellar disk.}
-\label{fig:transit}
 \end{figure}
@@
 where $\mu = \sqrt{1-x^2}$.
-A template is available in the source distribution for the creation of custom limb darkening profiles for any radially symmetric function.
+A template is available in the source distribution for the creation of custom profiles for any radially symmetric function.
 The nonlinear and custom models are calculated with the brute force integration scheme described in \S\,\ref{section:algorithm}, whereas models for quadratic and linear profiles are calculated analytically based on the widely-used formalism from \cite{mandel02}.
 
-For models calculated by integration, the user can specify the maximum allowable error with the \texttt{max\_err} parameter.
-When a transit model is initialized, the integration step size is tuned such that the difference between the model and a template calculated with a fine step size (FIXME) is below the specified tolerance.
-Computation time increases FIXME with precision (see FIXME), so .
-FIXME lower limit on precision.
-Figure\,\ref{fig:transit} shows an example transit light curve with errors FIXME.
+For models calculated by integration, the user can specify the maximum allowable truncation error with the \texttt{max\_err} parameter.
+When a transit model is initialized, the integration step size is tuned such that the difference between the model and a high-accuracy template (calculated with scale factor $f = 10^{-4}$) is below the specified tolerance.
+The lower limit on achievable accuracy with this method is below $10^{-10}$ for typical limb darkening profiles.
+Figure\,\ref{fig:transit} shows an example transit light curve and its truncation error.
 
-FIXME this needs work.
-One of the free parameters in the transit models is the separation of centers $d$ between the star and the planet.
+\begin{figure}
+\resizebox{\hsize}{!}{\includegraphics{f2.pdf}}
+\caption{An example transit light curve for a nonlinear stellar intensity profile (top panel) and truncation error for the calculation (bottom panel).
+The error tolerance parameter \texttt{max\_err} was set to 1.0 parts per million.
+The truncation error increases with distance from the center of the star up to around $\pm0.2$ hours from the time of mid-transit, because the stellar intensity gradient is larger at larger radii.
+The error decreases again during ingress and egress as the planet eclipses a smaller fraction of the stellar disk.}
+\label{fig:transit}
+\end{figure}
+
+One of the required inputs for a transit model is the separation of centers $d$ between the star and the planet.
 \texttt{batman} computes $d$ automatically from the following set of user-specified parameters: planet semi-major axis, orbital inclination, eccentricity, longitude of periastron, orbital period, and time of periastron/mid-transit time (for eccentric/circular orbits).
-The separation $d$ is calculated with equation (5) from Chapter FIXME and FIXME of \citep{seager10}.
+FIXME add refs to eqns in Seager 10.
 
-The computationally intensive sections of the code are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
-For all limb darkening models, the calculation can be parallelized at the C level with OpenMP.
+The computationally intensive sections of the code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
+The code also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.
 The number of processors is specified by the user with the parameter \texttt{nthreads}.
-\texttt{batman} will raise an exception if users attempt to parallelize a calculation on a system where OpenMP is not supported.
+\texttt{batman} will raise an exception if the user attempts to parallelize a calculation on a system where OpenMP is not supported.
 
 \subsection{Performance}
+\begin{figure}
+\resizebox{\hsize}{!}{\includegraphics{f3.pdf}}
+\caption{Execution time as a function of truncation error tolerance.}
+\label{fig:timing}
+\end{figure}
+
 \subsection{Comparison with other codes}
@@
 \acknowledgments
-Props to Robin.
 

### 2015-07-16 manuscript/figs/fig3.py

@@ -1,0 +1,82 @@
+# The batman package: fast computation of exoplanet transit light curves
+# Copyright (C) 2015 Laura Kreidberg	 
+# 
+# This program is free software: you can redistribute it and/or modify
+# it under the terms of the GNU General Public License as published by
+# the Free Software Foundation, either version 3 of the License, or
+# (at your option) any later version.
+# 
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+# 
+# You should have received a copy of the GNU General Public License
+# along with this program.  If not, see <http://www.gnu.org/licenses/>.
+
+from __future__ import print_function
+import numpy as np
+import math
+import matplotlib.pyplot as plt
+import timeit
+import batman
+from pylab import *
+
+def wrapper(func, *args, **kwargs):
+    def wrapped():
+        return func(*args, **kwargs)
+    return wrapped
+
+rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
+#matplotlib.rcParams.update({'font.size':14})
+plt.figure(figsize=(4,4))
+
+params = batman.TransitParams()	        #object to store transit parameters
+params.t0 = 0. 				#time of periastron passage (for eccentric orbits), OR
+					#mid-transit time (for circular orbits)
+params.per = 1.58			#orbital period	
+params.rp = 0.1				#planet radius (in units of stellar radii)
+params.a = 15.				#semi-major axis (in units of stellar radii)
+params.inc = 87.			#orbital inclination (in degrees)	
+params.ecc = 0.				#eccentricity	
+params.w = 90.				#longitude of periastron (in degrees)
+params.limb_dark = "nonlinear"          #limb darkening model
+params.u = [0.5, 0.1, 0.1, -0.1]       	#limb darkening coefficients
+   
+t = np.linspace(-0.025, 0.025, 1000)  	#times at which to calculate light curve	
+m = batman.TransitModel(params, t)      #initializes model
+
+flux = m.LightCurve(params)		#calculates light curve
+
[…]

### 2015-07-16 manuscript/figs/fig2.py
change call from _set_fac to set_fac

@@ -27,7 +27,7 @@ m = batman.TransitModel(params, t)      #initializes model
 
 flux = m.LightCurve(params)		#calculates light curve
 
-m._set_fac(1.0e-4)
+m.set_fac(1.0e-4)
 flux_precise = m.LightCurve(params)
 
 gs = gridspec.GridSpec(2, 1, height_ratios =[3,1], hspace=0.05)

### 2015-07-16 batman/plots.py
update plots

@@ -24,6 +24,7 @@ from . import _nonlinear_ld
 import timeit
 import batman
 
+
 def wrapper(func, *args, **kwargs):
     def wrapped():
         return func(*args, **kwargs)
@@ -72,7 +73,8 @@ def make_plots():
 		f= _nonlinear_ld._nonlinear_ld(zs, rp, u[0], u[1], u[2], u[3], fac[i], 12)
 		err = np.max(np.abs(f - f_ref))
 		errs.append(err)
-	plt.plot(np.array(ts), np.array(errs)*1.0e6)
+	plt.plot(np.array(ts), np.array(errs)*1.0e6, color='k')
+	plt.xlim((1.0e-3, 1.0e-1))
 	plt.yscale('log')
 	plt.xscale('log')
 	plt.xlabel("Time (s)")

### 2015-07-16 manuscript/figs/fig3.py
update fig3

@@ -40,26 +40,36 @@ params.a = 15.				#semi-major axis (in units of stellar radii)
 params.inc = 87.			#orbital inclination (in degrees)	
 params.ecc = 0.				#eccentricity	
 params.w = 90.				#longitude of periastron (in degrees)
+   
+t = np.linspace(-0.013, 0.013, 100)  	#times at which to calculate light curve	
+
+params.limb_dark = "quadratic"          #limb darkening model
+params.u = [0.1, 0.3]       	#limb darkening coefficients
+m = batman.TransitModel(params, t)      #initializes model
+wrapped = wrapper(m.LightCurve, params)
+time = timeit.timeit(wrapped,number=100)/100.
+plt.axvline(time, color='0.5', linestyle='dashed')
+#print(t)
+
 params.limb_dark = "nonlinear"          #limb darkening model
 params.u = [0.5, 0.1, 0.1, -0.1]       	#limb darkening coefficients
-   
-t = np.linspace(-0.025, 0.025, 1000)  	#times at which to calculate light curve	
 m = batman.TransitModel(params, t)      #initializes model
-
 flux = m.LightCurve(params)		#calculates light curve
+#plt.plot(t, flux)
+#plt.show()
 
 #generates Figure FIXME: max err as a function of function call time
-zs = np.linspace(0., 1., 1000)
 rp = 0.1
 u = [0., 0.7, 0.0, -0.3]
 n = 20
 ts = []
 errs = []
-fac = np.logspace(-3, -1, n)
+fac = np.logspace(-2.2, 0.3, n)
 for i in range(n):
 	m.set_fac(fac[i])
 	wrapped = wrapper(m.LightCurve, params)
-	t = timeit.timeit(wrapped,number=10)/10.
+	if i<10: t = timeit.timeit(wrapped,number=200)/200.
+	else: t = timeit.timeit(wrapped,number=1000)/1000.
 	ts.append(t)
 	print(t)
 	err = m.calc_err() 
@@ -67,8 +77,13 @@ for i in range(n):
 plt.plot(np.array(ts), np.array(errs), color='k')
 print(np.min(errs), np.max(errs))
 print(np.max(ts), np.min(ts))
-plt.xlim((1.0e-3, 5.0e-2))
-plt.ylim((1.0e-3, 10.))
+
+
+plt.gca().tick_params('both', length=8, width=1.2, which='major')
[…]

### 2015-07-20 c_src/_exponential_ld.c
add exponential LD law

@@ -1,0 +1,165 @@
+/* The batman package: fast computation of exoplanet transit light curves
+ * Copyright (C) 2015 Laura Kreidberg	 
+ * 
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ * 
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+#define NPY_NO_DEPRECATED_API NPY_1_7_API_VERSION
+#include <Python.h>
+#include "numpy/arrayobject.h"
+#include <math.h>
+
+#if defined (_OPENMP)
+#  include <omp.h>
+#endif
+
+#define MIN(x, y) (((x) < (y)) ? (x) : (y))
+#define MAX(x, y) (((x) > (y)) ? (x) : (y))
+
+static PyObject *_exponential_ld(PyObject *self, PyObject *args);
+
+double intensity(double r, double u1, double u2, double norm)
+{
+	if(r > 0.99995) r = 0.99995;
+	if(r < 0.00005) r = 0.00005;
+	double mu = sqrt(1. - r*r);
+	return (1. - u1*(1. - mu) - u2/(1. - exp(mu)))/norm; 
+}
+
+double area(double d, double r, double R)
+{
+	/*
+	Returns area of overlapping circles with radii r and R; separated by a distance d
+	*/
+	double arg1 = (d*d + r*r - R*R)/(2.*d*r); 	
+	double arg2 = (d*d + R*R - r*r)/(2.*d*R); 
+	double arg3 = MAX((-d + r + R)*(d + r - R)*(d - r + R)*(d + r + R), 0.);
+
+	if(r <= R - d) return M_PI*r*r;						//planet completely overlaps stellar circle
+	else if(r >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
+	else return r*r*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
+}
+
[…]

### 2015-07-20 c_src/_logarithmic_ld.c
add logarithmic LD law

@@ -1,0 +1,163 @@
+/* The batman package: fast computation of exoplanet transit light curves
+ * Copyright (C) 2015 Laura Kreidberg	 
+ * 
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ * 
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+#define NPY_NO_DEPRECATED_API NPY_1_7_API_VERSION
+#include <Python.h>
+#include "numpy/arrayobject.h"
+#include <math.h>
+
+#if defined (_OPENMP)
+#  include <omp.h>
+#endif
+
+#define MIN(x, y) (((x) < (y)) ? (x) : (y))
+#define MAX(x, y) (((x) > (y)) ? (x) : (y))
+
+static PyObject *_logarithmic_ld(PyObject *self, PyObject *args);
+
+double intensity(double r, double u1, double u2, double norm)
+{
+	if(r > 0.99995) r = 0.99995;
+	double mu = sqrt(1. - r*r);
+	return (1. - u1*(1.-mu) - u2*mu*log(mu))/norm; 
+}
+
+double area(double d, double r, double R)
+{
+	/*
+	Returns area of overlapping circles with radii r and R; separated by a distance d
+	*/
+	double arg1 = (d*d + r*r - R*R)/(2.*d*r); 	
+	double arg2 = (d*d + R*R - r*r)/(2.*d*R); 
+	double arg3 = MAX((-d + r + R)*(d + r - R)*(d - r + R)*(d + r + R), 0.);
+
+	if(r <= R - d) return M_PI*r*r;						//planet completely overlaps stellar circle
+	else if(r >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
+	else return r*r*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
+}
+
+static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
+{
[…]

### 2015-07-20 c_src/_exponential_ld.c
add exponential LD

@@ -85,7 +85,6 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 	#endif
 
 	double norm = 2.*M_PI*(0.5 - 0.1666666667*u1 + 0.77750463*u2); 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
-	printf("NEWBIE\n");
 
 	#if defined (_OPENMP)
 	#pragma omp parallel for private(d, r_in, r_out, delta, r, dr, A_i, A_f, I)

### 2015-07-20 setup.py
add exponential and logarithmic LD extensions

@@ -67,9 +67,10 @@ parallel_libraries = ['gomp'] if needs_gomp else []
 _nonlinear_ld = Extension('batman._nonlinear_ld', ['c_src/_nonlinear_ld.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
 _quadratic_ld = Extension('batman._quadratic_ld', ['c_src/_quadratic_ld.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
 _uniform_ld   = Extension('batman._uniform_ld', ['c_src/_uniform_ld.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
+_logarithmic_ld   = Extension('batman._logarithmic_ld', ['c_src/_logarithmic_ld.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
+_exponential_ld   = Extension('batman._exponential_ld', ['c_src/_exponential_ld.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
 _custom_ld   = Extension('batman._custom_ld', ['c_src/_custom_ld.c', 'c_src/_custom_intensity.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
 _rsky = Extension('batman._rsky', ['c_src/_rsky.c'])
-
 
 setup(	name='batman-package', 
 	version="0.9.1", 
@@ -87,5 +88,5 @@ setup(	name='batman-package',
 		],
 	include_dirs = [np.get_include()],
 	install_requires = ['numpy'],
-	ext_modules=[_nonlinear_ld, _quadratic_ld, _uniform_ld, _custom_ld, _rsky]
+	ext_modules=[_nonlinear_ld, _quadratic_ld, _uniform_ld, _logarithmic_ld, _exponential_ld, _custom_ld, _rsky]
 )

### 2015-07-20 c_src/_mandelagol_nonlinear_ld.c
add original Mandel/Agol nonlinear alg.

@@ -42,7 +42,7 @@ static PyObject *_mandelagol_nonlinear_ld(PyObject *self, PyObject *args)
 
 	double pi = acos(-1.);
 	double c1, c2, c3, c4, p;
-  	if(!PyArg_ParseTuple(args,"Odddddi", &z0, &c1, &c2, &c3, & c4, &p, &nz)) return NULL;
+  	if(!PyArg_ParseTuple(args,"Odddddi", &z0, &c1, &c2, &c3, &c4, &p, &nz)) return NULL;
 	rl = p;
 
 	dims[0] = z0->dimensions[0];
@@ -89,8 +89,8 @@ static PyObject *_mandelagol_nonlinear_ld(PyObject *self, PyObject *args)
 	nr=2;
 	dmumax=1.;
 
-//	while(dmumax > fac*1.e-3)
-	while(dmumax > fac*3.7e-5)
+	while(dmumax > fac*2.e-4)
+	//while(dmumax > fac*3.7e-5)
 	{
 		//printf("%0.10f\n", dmumax);
 		for(i = i1; i <= i2; i++) mulimbp[i] = mulimb[i];	

### 2015-07-20 batman/transitmodel.py
adding logarithmic, exponential, and squareroot LD

@@ -19,8 +19,11 @@ import matplotlib.pyplot as plt
 from . import _nonlinear_ld
 from . import _quadratic_ld
 from . import _uniform_ld
+from . import _logarithmic_ld
+from . import _exponential_ld
 from . import _custom_ld
 from . import _rsky
+#from . import _mandelagol_nonlinear_ld
 from math import pi
 import multiprocessing
 from . import openmp
@@ -49,17 +52,19 @@ class TransitModel:
 	"""
 	def __init__(self, params, t, max_err=1.0, nthreads = 1):
 		#checking for invalid input
-		if (params.limb_dark == "uniform" and len(params.u) != 0) or (params.limb_dark == "linear" and len(params.u) != 1) or \
-		    (params.limb_dark == "quadratic" and len(params.u) != 2) or (params.limb_dark == "nonlinear" and len(params.u) != 4):
+		if  (params.limb_dark == "uniform" and len(params.u) != 0) or (params.limb_dark == "linear" and len(params.u) != 1) or \
+		    (params.limb_dark == "quadratic" and len(params.u) != 2) or (params.limb_dark == "logarithmic" and len(params.u) != 2) or \
+		    (params.limb_dark == "exponential" and len(params.u) != 2) or (params.limb_dark == "squareroot" and len(params.u) != 2) or \
+		    (params.limb_dark == "nonlinear" and len(params.u) != 4):
 			raise Exception("Incorrect number of coefficients for " +params.limb_dark + " limb darkening; u should have the form:\n \
 			 u = [] for uniform LD\n \
 			 u = [u1] for linear LD\n \
-  			 u = [u1, u2] for quadratic LD\n \
+  			 u = [u1, u2] for quadratic, logarithmic, exponential, and squareroot LD\n \
 			 u = [u1, u2, u3, u4] for nonlinear LD, or\n \
 		         u = [u1, ..., un] for custom LD") 
-		if params.limb_dark not in ["uniform", "linear", "quadratic", "nonlinear", "custom"]: 
+		if params.limb_dark not in ["uniform", "linear", "quadratic", "logarithmic", "exponential", "squareroot", "nonlinear", "custom"]: 
 			raise Exception("\""+params.limb_dark+"\""+" limb darkening not supported; allowed options are:\n \
[…]

### 2015-07-20 manuscript/ms.tex
update references, add fig3

@@
 \section{Introduction}
+When an extrasolar planet transits in front of its host star, a small fraction of the stellar flux is blocked by the planet.
+Precise monitoring of stellar brightness as a function of time can therefore reveal transiting planets, and thousands of planets have been discovered with this technique (FIXME refs).
+The transit light curves can be used to characterize the planets in detail, enabling measurements of planetary radius, orbital period, and even atmospheric temperature structure and composition (FIXME refs).
+%This characterization depends on a foundation of accurate transit light curve models.
+
+Even though transit light curve models are a very basic part of the toolkit for exoplanet characterization, they are not trivial to compute.
+The model must account for the planet's size and position on the sky, as well as stellar limb darkening, a phenomenon that causes the apparent brightness of the stellar disk to decrease from center to edge \citep{milne21}.
+
+Transit light curve models can be calculated analytically for several limb darkening profiles.
+\citep{mandel02} presented analytic solutions for quadratic and nonlinear limb darkening laws.
+\citep{gimenez06} present Other limb darkening profiles do not have analytic solutions, and transit models must be calculated by numeric integration of the stellar intensity over the disk of the planet.
+Gimenez.
+
+These methods are implemented in several software packages.
+
 %Previous methods --
 
-%Mandel and Agol 2002 - analytic, but uses brute force integration for nonlinear
+%Mandel and Agol 2002 - analytic, but uses numeric integration for nonlinear
 %Giminez 2006 -
@@
 %Abubekerov 
-%Kjurkchieva - brute force; radially symmetric; 
+%Kjurkchieva - numeric; radially symmetric; 
 %PyTransit
@@
 
+See Figure\,\ref{fig:integration} for a schematic illustrating the geometry of the integration.
 The advantage of this integration scheme is that the only error introduced is due to approximating the stellar intensity as constant over the differential area element $\Delta A = A(x_i, r_p, d) - A(x_{i-1}, r_p, d)$.
@@
 
-See Figure\,\ref{fig:integration} for a schematic illustrating the geometry of the integration.
-
 \begin{figure}
@@
 A single integration element $\Delta A$ is shaded in orange.
-The integration step size illustrated here is larger than for a typical calculation for visual clarity.}
+The integration step size illustrated here is larger than for a typical calculation for visual clarity.
+Note that because the stellar intensity profile is radially symmetric, the coordinate system can be chosen such that the planet lies on the x-axis as shown.}
 \label{fig:integration}
@@
 \texttt{batman} is an open source project and is being developed on GitHub.
-Full documentation is available at \blue{\url{https://github.com/lkreidberg/batman}}, but I summarize the main capabilities of the package here.
+Full documentation is available at \blue{\url{https://github.com/lkreidberg/batman}}.
+I summarize the main capabilities of the package here.
 
@@
 \end{align}
-where $\mu = \sqrt{1-x^2}$.
+where $\mu = \sqrt{1-x^2}$ and $u_1, ..., u_n$ are limb darkening coefficients.
+The four-parameter nonlinear law provides a better fit to stellar models than simpler parameterizations \citep{claret00}.
+The \texttt{batman} source distribution also includes a template for the creation of a custom profile for any radially symmetric function, such as the logarithmic law of \cite{klinglesmith70} or the exponential law from \cite{claret03}.
+
 A template is available in the source distribution for the creation of custom profiles for any radially symmetric function.
-The nonlinear and custom models are calculated with the brute force integration scheme described in \S\,\ref{section:algorithm}, whereas models for quadratic and linear profiles are calculated analytically based on the widely-used formalism from \cite{mandel02}.
+The nonlinear and custom models are calculated with the numeric integration scheme described in \S\,\ref{section:algorithm}, whereas models for quadratic and linear profiles are calculated analytically based on the widely-used formalism from \cite{mandel02}.
 
@@
 \resizebox{\hsize}{!}{\includegraphics{f3.pdf}}
-\caption{Execution time as a function of truncation error tolerance.}
-\label{fig:timing}
+\caption{Execution time as a function of truncation error tolerance for a nonlinear model (black curve), compared to the execution time for a quadratic model (dashed gray line). FIXME finish.}
+\label{fig:performance}
 \end{figure}
+I tested \texttt{batman}'s performance over a range of typical use cases using a 1.7 GHz Intel Core i5 processor.
+In Figure\,\ref{fig:performance}, I show the execution time versus error tolerance for a single transit light curve calculation using a nonlinear intensity profile, compared to the execution time for a quadratic model computed analytically.
+The test case consists of 100 points evenly sampled in time during the planet's transit.
+I used physical parameters from the GJ 1214b system \citep{kreidberg14a}.
+The stellar intensity profile is the same for the nonlinear and quadratic models: the nonlinear limb darkening coefficients are $[0.0, 0.7, 0.0, -0.3]$ and the quadratic coefficients are $[0.1, 0.3]$.
+For this case, reaching a truncation error of 1.0 ppm via integration is just 2.5 times slower than the analytic calculation.
+FIXME add what the power law relation for timing is.
 
-\subsection{Comparison with other codes}
+\subsection{Comparison with Analytic Models for Nonlinear Limb Darkening}
+I explored using an analytic model to calculate transit light curves for nonlinear limb darkening profiles.
+The analytic solution for nonlinear limb darkening was presented in \cite{mandel02}, but it is not used in any published software packages; the original code provided by \cite{mandel02} uses numeric integration.
+I tested a Python/C implementation of the \cite{mandel02} algorithm and found that the new integration scheme in \S\,\ref{section:algorithm} is over 20 times faster (for a truncation error tolerance of 1 ppm).
 
-\section{Discussion}
+The analytic solution is challenging to compute because it uses the Appell F1 hypergeometric function.
+This function is only convergent for certain regions of parameter space and must be calculated with analytic continuation for other cases.
+\cite{colavecchia04} provide a Fortran library for computing Appell F1.
+I used this library to implement the analytic model for nonlinear limb darkening but did not achieve good results.
+The returned Appell F1 values are not accurate for all input parameters (based on a comparison with \texttt{Mathematica} and the pure Python library \texttt{mpmath}).
+Even for cases where F1 is correct, the computation is FIXME times slower than numeric integration (for an error tolerance of FIXME).
+I concluded that integration is a faster and easier solution than analytic models for nonlinear limb darkening.
+
+\section{Summary}
+%In this paper I introduced the \texttt{batman} package.
+%in preparation for ultra-precise light curves from future telescopes.
+%now have transit model that is good enough for JWST
 

### 2015-07-20 manuscript/figs/fig3.py

@@ -49,19 +49,18 @@ m = batman.TransitModel(params, t)      #initializes model
 wrapped = wrapper(m.LightCurve, params)
 time = timeit.timeit(wrapped,number=100)/100.
 plt.axvline(time, color='0.5', linestyle='dashed')
-#print(t)
+print("quadratic", time)
 
 params.limb_dark = "nonlinear"          #limb darkening model
-params.u = [0.5, 0.1, 0.1, -0.1]       	#limb darkening coefficients
+params.u = [0.0, 0.7, 0.0, -0.3]       	#limb darkening coefficients
 m = batman.TransitModel(params, t)      #initializes model
 flux = m.LightCurve(params)		#calculates light curve
 #plt.plot(t, flux)
 #plt.show()
 
 #generates Figure FIXME: max err as a function of function call time
-rp = 0.1
-u = [0., 0.7, 0.0, -0.3]
-n = 20
+#u = [0., 0.7, 0.0, -0.3]
+n = 100
 ts = []
 errs = []
 fac = np.logspace(-2.2, 0.3, n)
@@ -71,8 +70,8 @@ for i in range(n):
 	if i<10: t = timeit.timeit(wrapped,number=200)/200.
 	else: t = timeit.timeit(wrapped,number=1000)/1000.
 	ts.append(t)
-	print(t)
 	err = m.calc_err() 
+	print(t, err)
 	errs.append(err)
 plt.plot(np.array(ts), np.array(errs), color='k')
 print(np.min(errs), np.max(errs))

### 2015-07-22 manuscript/ms.tex
add introduction

@@
 \begin{abstract}
-Batman -- the best super-hero.
+I introduce \texttt{batman}, a Python package for calculating exoplanet transit light curve models.
 \end{abstract}
@@
 When an extrasolar planet transits in front of its host star, a small fraction of the stellar flux is blocked by the planet.
-Precise monitoring of stellar brightness as a function of time can therefore reveal transiting planets, and thousands of planets have been discovered with this technique (FIXME refs).
-The transit light curves can be used to characterize the planets in detail, enabling measurements of planetary radius, orbital period, and even atmospheric temperature structure and composition (FIXME refs).
+Precise measurements of stellar brightness as a function of time (light curves) can therefore reveal transiting planets.
+Over a thousand planets have been discovered with this technique \citep[e.g.][]{rowe2015}.
+Transit light curves can be used to characterize the planets in detail, enabling measurements of planetary radius, orbital elements, and even atmospheric temperature structure and composition \citep{seager00, seager03}.
 %This characterization depends on a foundation of accurate transit light curve models.
@@
 Even though transit light curve models are a very basic part of the toolkit for exoplanet characterization, they are not trivial to compute.
-The model must account for the planet's size and position on the sky, as well as stellar limb darkening, a phenomenon that causes the apparent brightness of the stellar disk to decrease from center to edge \citep{milne21}.
+The model must account for the planet's size and position on the sky, as well as stellar limb darkening, which causes the apparent brightness of the stellar disk to decrease from center to edge.
+The  stellar intensity profile can be fit with several functional forms, including a linear limb darkening law \citep{schwarzschild06}, quadratic \citep{kopal50}, square-root \citep{diaz-cordoves92}, logarithmic \citep{klinglesmith70}, exponential \citep{claret03}, and four-parameter nonlinear \citep{claret00}.
+For some of these profiles, model transit light curves can be calculated analytically \citep{mandel02, gimenez06, abubekerov13}.
+Other profiles do not have analytic solutions, and models must be calculated by numeric integration of the stellar intensity over the disk of the planet.
 
-Transit light curve models can be calculated analytically for several limb darkening profiles.
-\citep{mandel02} presented analytic solutions for quadratic and nonlinear limb darkening laws.
-\citep{gimenez06} present Other limb darkening profiles do not have analytic solutions, and transit models must be calculated by numeric integration of the stellar intensity over the disk of the planet.
-Gimenez.
+A number of codes are available to calculate transit light curves.
+\cite{mandel02} provide Fortran/IDL routines to compute models for quadratic and nonlinear limb darkening laws.
+The software packages \texttt{TAP} \citep{gazak12} and \texttt{EXOFAST} \citep{eastman13} include IDL implementations of the \cite{mandel02} algorithm for quadratic limb darkening.
+\texttt{JKTEBOP} calculates models in Fortran for a broad range of limb darkening laws \citep{southworth04}.
+\cite{kjurkchieva13} introduce the pure Python code \texttt{TAC-maker}, which performs numeric integration for arbitrary limb darkening profiles.
+There are also routines available to model simultaneous transits by one or more bodies \citep{kipping11, pal12}.
+Most recently, \cite{parvianinen15} released the Python package \texttt{PyTransit}, which implements analytic models from \cite{mandel02} and \cite{gimenez06}.
 
-These methods are implemented in several software packages.
+In this paper, I introduce the Python package \texttt{batman}.
+This package enables fast computation of transit light curves for uniform, linear, quadratic, logarithmic, exponential, and nonlinear limb darkening laws.
+Light curves for the first three of these are calculated analytically based on the formalism from \cite{mandel02}.
+Models for the remaining limb darkening cases are computed with an efficient new integration scheme, described in \S\,\ref{section:algorithm}.
+I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:conclusion}.
 
-%Previous methods --
-
-%Mandel and Agol 2002 - analytic, but uses numeric integration for nonlinear
-%Giminez 2006 -
-%Pal 2008 - 
-%Kipping 2011 -
-%Pal 2011 - multiple bodies, possibly lacking radial symmetry
-%	basically same as Mandel and Agol for linear/quadratic; more complex LD not implemented; code available (see discussion)
-%Abubekerov 
-%Kjurkchieva - numeric; radially symmetric; 
-%PyTransit
+%This package implements an efficient new algorithm (described in \S\,\ref{section:algorithm}) to calculate transit light curve models for any radially symmetric limb darkening profile.
+The transit models are computed with C extension modules written with the Python/C API, and the code is parallelized at the C level with OpenMP.
+The package therefore   mostly widely used programming language
 
@@
 where $I$ is normalized such that the integrated intensity over the stellar disk is unity.
-This expression is valid for any general stellar surface brightness map; however, it is slow to evaluate numerically because the differential area elements must be small ($<10^{-6}$) in order to achieve better than one part per million (ppm) accuracy.
+This expression is valid for any general stellar surface brightness map; however, it is slow to evaluate numerically because the differential area elements must be small ($\lesssim 10^{-6}$) in order to achieve better than one part per million (ppm) accuracy.
 
@@
 
-To further optimize the integration, I use a nonuniform step size.
+The integration can be further optimized by using a nonuniform step size.
 Typical stellar intensity profiles have larger gradients near the limb of the star than at the center \citep[e.g.][]{claret00}, so smaller steps are required at larger $x$ values to achieve the same accuracy.
@@
 The integration step size illustrated here is larger than for a typical calculation for visual clarity.
-Note that because the stellar intensity profile is radially symmetric, the coordinate system can be chosen such that the planet lies on the x-axis as shown.}
+Note that because the stellar intensity profile is radially symmetric, the coordinate system can be chosen such that the planet lies on the x-axis, as shown.}
 \label{fig:integration}
@@
 \section{The \texttt{batman} package}
-I wrote a Python package, \texttt{batman}, to calculate model transit light curves.
+The Python package \texttt{batman} implements the algorithm described in \S\,\ref{section:algorithm} to calculate model transit light curves.
 \texttt{batman} is an open source project and is being developed on GitHub.
@@
 
-\subsection{Features}
-\texttt{batman} supports calculation of exoplanet transit light curves for uniform, linear, quadratic, nonlinear, and custom stellar intensity profiles:
+\subsection{Limb-Darkening Models}
+\texttt{batman} supports calculation of exoplanet transit light curves for uniform, linear \citep{schwarzschild06}, quadratic \citep{kopal50}, square-root \citep{diaz-cordoves92}, logarithmic \citep{klinglesmith70}, exponential \citep{claret03}, and nonlinear \citep{claret00} stellar intensity profiles:
 \begin{align}
   I(\mu) &= I_0                                                                         & &\text{(uniform)}             \\
-  I(\mu) &= I_0[1 - u_1(1-\mu)]                                                         & &\text{(linear)}              \\
-  I(\mu) &= I_0[1 - u_1(1 - \mu) - u_2(1-\mu)^2]                                        & &\text{(quadratic)}           \\
-  I(\mu) &= I_0[1 - u_1(1-\mu^{1/2}) - u_2(1- \mu) - u_3(1-\mu^{3/2}) - u_4(1-\mu^2)]   & &\text{(nonlinear)}
+  I(\mu) &= I_0[1 - c_1(1-\mu)]                                                         & &\text{(linear)}              \\
+  I(\mu) &= I_0[1 - c_1(1 - \mu) - c_2(1-\mu)^2]                                        & &\text{(quadratic)}           \\
+  I(\mu) &= I_0[1 - c_1(1 - \mu) - c_2(1-\sqrt{\mu})]                                   & &\text{(square-root)}         \\
+  I(\mu) &= I_0[1 - c_1(1 - \mu) - c_2\mu\ln{\mu}]                                      & &\text{(logarithmic)}         \\
+  I(\mu) &= I_0\left[1 - c_1(1 - \mu) - c_2/(1-\exp{\mu})\right]                  	& &\text{(exponential)}         \\
+  I(\mu) &= I_0[1 - c_1(1-\mu^{1/2}) - c_2(1- \mu) - c_3(1-\mu^{3/2}) - c_4(1-\mu^2)]   & &\text{(nonlinear)}
 \end{align}
-where $\mu = \sqrt{1-x^2}$ and $u_1, ..., u_n$ are limb darkening coefficients.
-The four-parameter nonlinear law provides a better fit to stellar models than simpler parameterizations \citep{claret00}.
-The \texttt{batman} source distribution also includes a template for the creation of a custom profile for any radially symmetric function, such as the logarithmic law of \cite{klinglesmith70} or the exponential law from \cite{claret03}.
+where $\mu = \sqrt{1-x^2}$ and $c_1, ..., c_n$ are limb darkening coefficients.
+The \texttt{batman} source distribution also includes a template for the creation of a custom profile for any radially symmetric function.
 
-A template is available in the source distribution for the creation of custom profiles for any radially symmetric function.
-The nonlinear and custom models are calculated with the numeric integration scheme described in \S\,\ref{section:algorithm}, whereas models for quadratic and linear profiles are calculated analytically based on the widely-used formalism from \cite{mandel02}.
+The square-root, logarithmic, exponential, nonlinear, and custom models are computed with the numeric integration scheme from \S\,\ref{section:algorithm}.
+The uniform, linear, and quadratic models are calculated analytically, with code based on the Fortran routines \texttt{occultquad.f} and \texttt{occultuniform.f} provided by \cite{mandel02}.
 
+\subsection{Accuracy}
 For models calculated by integration, the user can specify the maximum allowable truncation error with the \texttt{max\_err} parameter.
@@
 
+\subsection{Separation of Centers}
 One of the required inputs for a transit model is the separation of centers $d$ between the star and the planet.
@@
 FIXME add refs to eqns in Seager 10.
-
-The computationally intensive sections of the code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
-The code also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.
-The number of processors is specified by the user with the parameter \texttt{nthreads}.
-\texttt{batman} will raise an exception if the user attempts to parallelize a calculation on a system where OpenMP is not supported.
 
@@
 \end{figure}
+The computationally intensive sections of the code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
+The code also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.
+The number of processors is specified by the user with the parameter \texttt{nthreads}.
+\texttt{batman} will raise an exception if the user attempts to parallelize a calculation on a system where OpenMP is not supported.
+
 I tested \texttt{batman}'s performance over a range of typical use cases using a 1.7 GHz Intel Core i5 processor.
@@
 \subsection{Comparison with Analytic Models for Nonlinear Limb Darkening}
-I explored using an analytic model to calculate transit light curves for nonlinear limb darkening profiles.
+I explored using an analytic model to calculate transit light curves for the four-parameter nonlinear limb darkening profile.
 The analytic solution for nonlinear limb darkening was presented in \cite{mandel02}, but it is not used in any published software packages; the original code provided by \cite{mandel02} uses numeric integration.

### 2015-07-22 c_src/_quadratic_ld.c
switch from rj() to ellpic_bulirsch()

@@ -29,6 +29,7 @@
 
 double rc(double x, double y);
 double rj(double x, double y, double z, double p);
+double ellpic_bulirsch(double n, double k);
 double ellec(double k);
 double ellk(double k);
 double rf(double x, double y, double z);
@@ -177,7 +178,8 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 			Kk = ellk(q);
 			Ek = ellec(q);
 			n = 1.0/x1 - 1.0;
-			Pk = Kk - n/3.0*rj(0.0, 1.0 - q*q, 1.0, 1.0 + n);
+			//Pk = Kk - n/3.0*rj(0.0, 1.0 - q*q, 1.0, 1.0 + n);
+			Pk = ellpic_bulirsch(n, q);
 			lambdad[i] = 1.0/9.0/M_PI/sqrt(p*z)*(((1.0 - x2)*(2.0*x2 +  \
 			        x1 - 3.0) - 3.0*x3*(x2 - 2.0))*Kk + 4.0*p*z*(z*z +  \
 			        7.0*p*p - 4.0)*Ek - 3.0*x3/x1*Pk);
@@ -196,7 +198,9 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 			Kk = ellk(q);
 			Ek = ellec(q);
 			n = x2/x1 - 1.0;
-			Pk = Kk - n/3.0*rj(0.0, 1.0 - q*q, 1.0, 1.0 + n);
+			//Pk = Kk - n/3.0*rj(0.0, 1.0 - q*q, 1.0, 1.0 + n);
+			Pk = ellpic_bulirsch(n, q);
+			
 			lambdad[i] = 2.0/9.0/M_PI/sqrt(1.0 - x1)*((1.0 - 5.0*z*z + p*p +  \
 			         x3*x3)*Kk + (1.0 - x1)*(z*z + 7.0*p*p - 4.0)*Ek - 3.0*x3/x1*Pk);
 			if(z < p) lambdad[i] += 2.0/3.0;
@@ -328,6 +332,44 @@ double rj(double x, double y, double z, double p)
 	return rj;  
 }
 	
+double ellpic_bulirsch(double n, double k)
+{
+	double kc = sqrt(1.-k*k);	
+	double p = n + 1.;
+	double m0 = 1.;
+	double c = 1.;
+	p = sqrt(p);
+	double d = 1./p;
[…]

### 2015-07-22 manuscript/ms.tex
update intro

@@
 A number of codes are available to calculate transit light curves.
-\cite{mandel02} provide Fortran/IDL routines to compute models for quadratic and nonlinear limb darkening laws.
+\cite{mandel02} provide Fortran and IDL routines to compute models for quadratic and nonlinear limb darkening laws.
 The software packages \texttt{TAP} \citep{gazak12} and \texttt{EXOFAST} \citep{eastman13} include IDL implementations of the \cite{mandel02} algorithm for quadratic limb darkening.
@@
 In this paper, I introduce the Python package \texttt{batman}.
-This package enables fast computation of transit light curves for uniform, linear, quadratic, logarithmic, exponential, and nonlinear limb darkening laws.
+This package enables fast computation of transit light curves for any radially symmetric limb darkening law, and currently supports uniform, linear, quadratic, logarithmic, exponential, and nonlinear limb darkening.
 Light curves for the first three of these are calculated analytically based on the formalism from \cite{mandel02}.
-Models for the remaining limb darkening cases are computed with an efficient new integration scheme, described in \S\,\ref{section:algorithm}.
+Models for the remaining cases are computed with an efficient new integration scheme, described in \S\,\ref{section:algorithm}.
 I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:conclusion}.
@@
 The uniform, linear, and quadratic models are calculated analytically, with code based on the Fortran routines \texttt{occultquad.f} and \texttt{occultuniform.f} provided by \cite{mandel02}.
+FIXME updated elliptic intregrals??
 
 \subsection{Accuracy}
+FIXME accurate transitss importantntttt!!
 For models calculated by integration, the user can specify the maximum allowable truncation error with the \texttt{max\_err} parameter.

### 2015-07-22 c_src/_quadratic_ld.c
add docstring to ellpic_bulirsch function

@@ -332,13 +332,32 @@ double rj(double x, double y, double z, double p)
 	return rj;  
 }
 	
+/*
+
+   Computes the complete elliptical integral of the third kind using
+   the algorithm of Bulirsch (1965):
+
+   Bulirsch 1965, Numerische Mathematik, 7, 78
+   Bulirsch 1965, Numerische Mathematik, 7, 353
+
+ INPUTS:
+
+    n,k - int(dtheta/((1-n*sin(theta)^2)*sqrt(1-k^2*sin(theta)^2)),0, pi/2)
+
+ RESULT:
+
+    The complete elliptical integral of the third kind
+
+ -- translated from the ellpic_bulirsch.pro routine from EXOFAST
+ -- Eastman et al. 2013, PASP 125, 83
+*/
+
 double ellpic_bulirsch(double n, double k)
 {
 	double kc = sqrt(1.-k*k);	
-	double p = n + 1.;
+	double p = sqrt(n + 1.);
 	double m0 = 1.;
 	double c = 1.;
-	p = sqrt(p);
 	double d = 1./p;
 	double e = kc;
 	double f, g;

### 2015-07-22 c_src/_custom_intensity.c
label LD coefficients c1...cn

@@ -19,16 +19,16 @@
 
 /*
 	- The intensity function returns the stellar intensity at a radius r; where 0 <= r <= 1
-	- The function arguments are the normalized radius (r), and limb darkening coefficients u1, ..., un
-	- see below for an example intensity profile for I(r) \propto 1 - u1*(1-sqrt(1-r^2)) - u2*ln((sqrt(1-r^2)+c)/(1+c))
+	- The function arguments are the normalized radius (r), and limb darkening coefficients c1, ..., un
+	- see below for an example intensity profile for I(r) \propto 1 - c1*(1-sqrt(1-r^2)) - c2*ln((sqrt(1-r^2)+c)/(1+c))
 	- The normalization constant is calculated by constraining the integrated intensity to equal 1:
 		\int_r \int_theta {I(r)*r*dr*dtheta}/norm = 1
 */
 
-double intensity(double r, double u1, double u2, double u3, double u4, double u5, double u6) 
+double intensity(double r, double c1, double c2, double c3, double c4, double c5, double c6) 
 {
 	if(r > 0.99995) r = 0.99995;
 	double mu = sqrt(1. - r*r);
-	double norm = 2.*M_PI*(-u1/6. - u2*u3/2. + u2/4. + 0.5 + u2*u3*u3*log(1. + 1./u3)/2.); 
-	return (1. - u1*(1. - mu) - u2*log((mu + u3)/(1. + u3)))/norm;
+	double norm = 2.*M_PI*(-c1/6. - c2*c3/2. + c2/4. + 0.5 + c2*c3*c3*log(1. + 1./c3)/2.); 
+	return (1. - c1*(1. - mu) - c2*log((mu + c3)/(1. + c3)))/norm;
 }

### 2015-07-22 c_src/_custom_ld.c

@@ -29,7 +29,7 @@
 
 static PyObject *_custom_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double u1, double u2, double u3, double u4, double u5, double u6);
+double intensity(double r, double c1, double c2, double c3, double c4, double c5, double c6);
 
 double area(double d, double r, double R)
 {
@@ -47,12 +47,12 @@ double area(double d, double r, double R)
 
 static PyObject *_custom_ld(PyObject *self, PyObject *args)
 {
-	double rprs, d, fac, A_i, r, I, dr, A_f, r_in, r_out, delta, u1, u2, u3, u4, u5, u6;
+	double rprs, d, fac, A_i, r, I, dr, A_f, r_in, r_out, delta, c1, c2, c3, c4, c5, c6;
 	int nthreads;
 	npy_intp i, dims[1];
 	PyArrayObject *zs, *flux;
 
-  	if(!PyArg_ParseTuple(args,"Oddddddddi", &zs, &rprs, &u1, &u2, &u3, &u4, &u5, &u6, &fac, &nthreads)) return NULL; //parses input arguments
+  	if(!PyArg_ParseTuple(args,"Oddddddddi", &zs, &rprs, &c1, &c2, &c3, &c4, &c5, &c6, &fac, &nthreads)) return NULL; //parses input arguments
 	
 	dims[0] = PyArray_DIMS(zs)[0]; 
 	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
@@ -99,7 +99,7 @@ static PyObject *_custom_ld(PyObject *self, PyObject *args)
 			while(r < r_out)
 			{
 				A_f = area(d, r, rprs);					//calculates area of overlapping circles
-				I = intensity(r - dr/2.,u1,u2, u3, u4, u5, u6); 	//intensity at the midpoint
+				I = intensity(r - dr/2.,c1,c2, c3, c4, c5, c6); 	//intensity at the midpoint
 				delta += (A_f - A_i)*I;					//increase in transit depth for this integration step
 				dr = fac*acos(r);  					//updating step size
 				r = r + dr;						//stepping to next element
@@ -108,7 +108,7 @@ static PyObject *_custom_ld(PyObject *self, PyObject *args)
 			dr = r_out - r + dr;  						//calculating change in radius for last step
 			r = r_out;							//final radius for integration
[…]

### 2015-07-22 c_src/_exponential_ld.c

@@ -29,12 +29,12 @@
 
 static PyObject *_exponential_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double u1, double u2, double norm)
+double intensity(double r, double c1, double c2, double norm)
 {
 	if(r > 0.99995) r = 0.99995;
 	if(r < 0.00005) r = 0.00005;
 	double mu = sqrt(1. - r*r);
-	return (1. - u1*(1. - mu) - u2/(1. - exp(mu)))/norm; 
+	return (1. - c1*(1. - mu) - c2/(1. - exp(mu)))/norm; 
 }
 
 double area(double d, double r, double R)
@@ -56,10 +56,10 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 	double rprs, d, fac, A_i, r, I; 
 	int nthreads;
 	npy_intp i, dims[1];
-	double dr, A_f, r_in, r_out, delta, u1, u2;
+	double dr, A_f, r_in, r_out, delta, c1, c2;
 	
 	PyArrayObject *zs, *flux;
-  	if(!PyArg_ParseTuple(args,"Oddddi", &zs, &rprs, &u1, &u2, &fac, &nthreads)) return NULL;
+  	if(!PyArg_ParseTuple(args,"Oddddi", &zs, &rprs, &c1, &c2, &fac, &nthreads)) return NULL;
 
 	dims[0] = PyArray_DIMS(zs)[0]; 
 	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
@@ -84,7 +84,7 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 	omp_set_num_threads(nthreads);	//specifies number of threads (if OpenMP is supported)
 	#endif
 
-	double norm = 2.*M_PI*(0.5 - 0.1666666667*u1 + 0.77750463*u2); 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
+	double norm = 2.*M_PI*(0.5 - 0.1666666667*c1 + 0.77750463*c2); 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
 
 	#if defined (_OPENMP)
 	#pragma omp parallel for private(d, r_in, r_out, delta, r, dr, A_i, A_f, I)
@@ -109,7 +109,7 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 			while(r < r_out)
 			{
[…]

### 2015-07-22 c_src/_logarithmic_ld.c

@@ -29,11 +29,11 @@
 
 static PyObject *_logarithmic_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double u1, double u2, double norm)
+double intensity(double r, double c1, double c2, double norm)
 {
 	if(r > 0.99995) r = 0.99995;
 	double mu = sqrt(1. - r*r);
-	return (1. - u1*(1.-mu) - u2*mu*log(mu))/norm; 
+	return (1. - c1*(1.-mu) - c2*mu*log(mu))/norm; 
 }
 
 double area(double d, double r, double R)
@@ -55,10 +55,10 @@ static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 	double rprs, d, fac, A_i, r, I; 
 	int nthreads;
 	npy_intp i, dims[1];
-	double dr, A_f, r_in, r_out, delta, u1, u2;
+	double dr, A_f, r_in, r_out, delta, c1, c2;
 	
 	PyArrayObject *zs, *flux;
-  	if(!PyArg_ParseTuple(args,"Oddddi", &zs, &rprs, &u1, &u2, &fac, &nthreads)) return NULL;
+  	if(!PyArg_ParseTuple(args,"Oddddi", &zs, &rprs, &c1, &c2, &fac, &nthreads)) return NULL;
 
 	dims[0] = PyArray_DIMS(zs)[0]; 
 	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
@@ -83,7 +83,7 @@ static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 	omp_set_num_threads(nthreads);	//specifies number of threads (if OpenMP is supported)
 	#endif
 
-	double norm = (-3.*u1 + 2.*u2 + 9.)*M_PI/9.0; 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
+	double norm = (-3.*c1 + 2.*c2 + 9.)*M_PI/9.0; 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
 
 	#if defined (_OPENMP)
 	#pragma omp parallel for private(d, r_in, r_out, delta, r, dr, A_i, A_f, I)
@@ -108,7 +108,7 @@ static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 			while(r < r_out)
 			{
 				A_f = area(d, r, rprs);				//calculates area of overlapping circles
[…]

### 2015-07-22 c_src/_nonlinear_ld.c

@@ -29,11 +29,11 @@
 
 static PyObject *_nonlinear_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double u1, double u2, double u3, double u4, double norm)
+double intensity(double r, double c1, double c2, double c3, double c4, double norm)
 {
 	if(r > 0.99995) r = 0.99995;
 	double sqrtmu = pow(1. - r*r,0.25);
-	return (1. - u1*(1. - sqrtmu) - u2*(1. - pow(sqrtmu,2.)) - u3*(1. - pow(sqrtmu, 3.)) - u4*(1. - pow(sqrtmu,4.)))/norm; 	
+	return (1. - c1*(1. - sqrtmu) - c2*(1. - pow(sqrtmu,2.)) - c3*(1. - pow(sqrtmu, 3.)) - c4*(1. - pow(sqrtmu,4.)))/norm; 	
 }
 
 double area(double d, double r, double R)
@@ -55,10 +55,10 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 	double rprs, d, fac, A_i, r, I; 
 	int nthreads;
 	npy_intp i, dims[1];
-	double dr, A_f, r_in, r_out, delta, u1, u2, u3, u4;
+	double dr, A_f, r_in, r_out, delta, c1, c2, c3, c4;
 	
 	PyArrayObject *zs, *flux;
-  	if(!PyArg_ParseTuple(args,"Oddddddi", &zs, &rprs, &u1, &u2, &u3, &u4, &fac, &nthreads)) return NULL;
+  	if(!PyArg_ParseTuple(args,"Oddddddi", &zs, &rprs, &c1, &c2, &c3, &c4, &fac, &nthreads)) return NULL;
 
 	dims[0] = PyArray_DIMS(zs)[0]; 
 	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
@@ -83,7 +83,7 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 	omp_set_num_threads(nthreads);	//specifies number of threads (if OpenMP is supported)
 	#endif
 
-	double norm = (-u1/10. - u2/6. - 3.*u3/14. - u4/4. + 0.5)*2.*M_PI; 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
[…]

### 2015-07-22 c_src/_quadratic_ld.c

@@ -40,8 +40,8 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 /*	Input: *************************************
  
 	 zs   	impact parameter in units of rs
-	 u1   	linear    limb-darkening coefficient (gamma_1 in Mandel & Agol 2002)
-	 u2   	quadratic limb-darkening coefficient (gamma_2)
+	 c1   	linear    limb-darkening coefficient (gamma_1 in Mandel & Agol 2002)
+	 c2   	quadratic limb-darkening coefficient (gamma_2)
 	 p    	occulting star size in units of rs
 	
 	 Output: ***********************************
@@ -49,16 +49,16 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 	 flux 	fraction of flux at each zs for a limb-darkened source
 	
 	 Limb darkening has the form:
-	 I(r) = [1 - u1 * (1 - sqrt(1 - (r/rs)^2)) - u2*(1 - sqrt(1 - (r/rs)^2))^2]/(1 - u1/3 - u2/6)/pi
+	 I(r) = [1 - c1 * (1 - sqrt(1 - (r/rs)^2)) - c2*(1 - sqrt(1 - (r/rs)^2))^2]/(1 - c1/3 - c2/6)/pi
 */
 	int nz, nthreads;
-	double u1, u2, p, *mu,  *lambdad, *etad, \
+	double c1, c2, p, *mu,  *lambdad, *etad, \
 		*lambdae, lam, x1, x2, x3, z, omega, kap0 = 0.0, kap1 = 0.0, \
 		q, Kk, Ek, Pk, n;
 	PyArrayObject *zs, *flux;
 	npy_intp i, dims[1];
 
-  	if(!PyArg_ParseTuple(args,"Odddi", &zs, &p, &u1, &u2, &nthreads)) return NULL;
+  	if(!PyArg_ParseTuple(args,"Odddi", &zs, &p, &c1, &c2, &nthreads)) return NULL;
 
 	dims[0] = PyArray_DIMS(zs)[0]; 
[…]

### 2015-07-23 manuscript/ms.tex
add abstract and conclusion

@@
 \begin{abstract}
-I introduce \texttt{batman}, a Python package for calculating exoplanet transit light curve models.
+I introduce \texttt{batman}, a Python package for modeling exoplanet transit light curves.
+\texttt{batman} supports calculation of light curves for any radially symmetric stellar limb darkening law, using a new integration algorithm for models that cannot be calculated analytically.
+The code uses C extension modules to speed up model calculation and is parallelized with OpenMP.
+(FIXME: say how long it takes to compute a million models).
+Model light curves can be computed with accuracy better than 0.0001 parts per million, ensuring that the community is prepared for the extraordinarily precise transit light curves we anticipate measuring with upcoming facilities.
 \end{abstract}
 
-\keywords{planets and satellites: atmospheres --- planets and satellites:  composition --- planets and satellites: individual: WASP-12b}
+\keywords{}
 
@@
 
-Even though transit light curve models are a very basic part of the toolkit for exoplanet characterization, they are not trivial to compute.
+Transit light curve models are a very basic part of the toolkit for exoplanet characterization, but they are not trivial to compute.
 The model must account for the planet's size and position on the sky, as well as stellar limb darkening, which causes the apparent brightness of the stellar disk to decrease from center to edge.
@@
 Models for the remaining cases are computed with an efficient new integration scheme, described in \S\,\ref{section:algorithm}.
-I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:conclusion}.
+I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:summary}.
 
@@
 where $x$ is the normalized radial coordinate $0 < x < 1$, $I(x)$ is the 1D stellar intensity profile, $r_p$ is the planetary radius, $d$ is the separation of centers between the star and the planet, and $A(x, r_p, d)$ is the area of intersection between two circles of radii $x$ and $r_p$, separated by a distance $d$.
-The sum is carried out over the range $x_0 = d - r_p$ to $x_n = \textrm{MIN}(d + r_p, r_s)$, where $r_s$ is the stellar radius.
+The sum is carried out over the range $x_0 = d - r_p$ to $x_n = \textrm{MIN}(d + r_p, r_s)$, where $r_s = 1$ is the stellar radius.
 The intersecting area is given by:
@@
 \section{The \texttt{batman} package}
+\label{section:batman}
 The Python package \texttt{batman} implements the algorithm described in \S\,\ref{section:algorithm} to calculate model transit light curves.
@@
 
-\subsection{Limb-Darkening Models}
-\texttt{batman} supports calculation of exoplanet transit light curves for uniform, linear \citep{schwarzschild06}, quadratic \citep{kopal50}, square-root \citep{diaz-cordoves92}, logarithmic \citep{klinglesmith70}, exponential \citep{claret03}, and nonlinear \citep{claret00} stellar intensity profiles:
+\subsection{Limb Darkening Models}
+\texttt{batman} supports calculation of exoplanet transit light curves for uniform, linear, quadratic, square-root, logarithmic, exponential, and four-parameter nonlinear stellar intensity profiles:
 \begin{align}
@@
 The uniform, linear, and quadratic models are calculated analytically, with code based on the Fortran routines \texttt{occultquad.f} and \texttt{occultuniform.f} provided by \cite{mandel02}.
-FIXME updated elliptic intregrals??
+For the analytic models, I follow \cite{eastman13} and use the algorithm from \cite{bulirsch65} to improve the calculation speed and accuracy for elliptic integrals of the third kind.
+
+\subsection{Utilities}
+\texttt{batman} includes a utility function to calculate the separation of centers $d$ between the star and the planet based on orbital parameters of the system.
+The input parameters are the planet semi-major axis $a$, inclination $i$, eccentricity $e$, longitude of periastron $\omega$, period $P$, and transit ephemeris $t_0$.
+For eccentric orbits, the transit ephemeris is the time of periastron; for circular orbits, $t_0$ is the mid-transit time.
+The separation of centers is given by:
+$$
+d = \frac{a(1-e^2)}{1+e\cos{f}}\sqrt{1-\sin^2{(\omega+f)}\sin^2{i}}
+$$
+where $f$ is defined as the orbital phase (for circular orbits) or the true anomaly (for eccentric orbits).
+The true anomaly is calculated with the algorithm provided by Murray \& Correia in Chapter 1 of \cite{seager10}.
 
 \subsection{Accuracy}
-FIXME accurate transitss importantntttt!!
-For models calculated by integration, the user can specify the maximum allowable truncation error with the \texttt{max\_err} parameter.
-When a transit model is initialized, the integration step size is tuned such that the difference between the model and a high-accuracy template (calculated with scale factor $f = 10^{-4}$) is below the specified tolerance.
-The lower limit on achievable accuracy with this method is below $10^{-10}$ for typical limb darkening profiles.
+Recent transit observations have yielded signal-to-noise greater than 1000 per exposure \citep[e.g.][]{kreidberg14a, knutson14b}.
+Accurate transit light curve calculation is essential for modeling such high precision measurements and will be increasingly important for data obtained with next-generation facilities.
+\texttt{batman} therefore enables the user to specify the maximum allowable truncation error for numeric integration.
 Figure\,\ref{fig:transit} shows an example transit light curve and its truncation error.
+
+To ensure that the truncation error is below the specified threshold, the integration step size is tuned during model initialization.
+Truncation error is measured relative to a model calculated with a very small step size ($f = 10^{-4}$).
+For typical limb darkening profiles, this method is reliable for truncation errors down to $\sim 10^{-4}$ ppm.
+However, tuning the step size is a slow operation because it requires computing several light curve models ($\sim10$).
+As an alternative, methods are available to set the step size directly and calculate the corresponding truncation error.
+
+I tested the accuracy of the analytic model for quadratic limb darkening by comparing it to a numerically integrated model with an error tolerance of 0.001 ppm.
+The analytic model is accurate to 0.03 ppm for a test case with $r_p = 0.1$, $(c_1, c_2) = (0.1, 0.3)$, sampled at $10^6$ evenly spaced points over the interval $0 < d < 1$.
+The accuracy is somewhat worse than machine epsilon because of error tolerance in the computation of special functions.
+
+I also tested the accuracy of the widely-used \cite{mandel02} code that uses Numerical Recipes functions to calculate elliptic integrals \citep{press92}.
+The accuracy was better than 0.005 ppm for most input values; however, for the case $r_p - d < \epsilon$, the error in the light curve exceeded 2 ppm.
+By contast, the \cite{bulirsch65} algorithm for elliptic integrals is well-behaved for this case and also faster.
 
@@
 \caption{An example transit light curve for a nonlinear stellar intensity profile (top panel) and truncation error for the calculation (bottom panel).
-The error tolerance parameter \texttt{max\_err} was set to 1.0 parts per million.
+The error tolerance parameter was set to 1.0 ppm.
 The truncation error increases with distance from the center of the star up to around $\pm0.2$ hours from the time of mid-transit, because the stellar intensity gradient is larger at larger radii.
@@
 
-\subsection{Separation of Centers}
-One of the required inputs for a transit model is the separation of centers $d$ between the star and the planet.
-\texttt{batman} computes $d$ automatically from the following set of user-specified parameters: planet semi-major axis, orbital inclination, eccentricity, longitude of periastron, orbital period, and time of periastron/mid-transit time (for eccentric/circular orbits).
-FIXME add refs to eqns in Seager 10.
-
 \subsection{Performance}
@@
 \resizebox{\hsize}{!}{\includegraphics{f3.pdf}}
-\caption{Execution time as a function of truncation error tolerance for a nonlinear model (black curve), compared to the execution time for a quadratic model (dashed gray line). FIXME finish.}
+\caption{Execution time as a function of truncation error tolerance for a light curve modeled with the nonlinear limb darkening law (black line).
+The execution time for a quadratic model (computed analytically to better than 0.05 ppm accuracy) is indicated by the arrow.
+Calculations were made with a 1.7 GHz Intel Core i5 processor.}
 \label{fig:performance}
 \end{figure}
+
 The computationally intensive sections of the code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
 The code also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.
-The number of processors is specified by the user with the parameter \texttt{nthreads}.
+The number of processors is specified by the user.
 \texttt{batman} will raise an exception if the user attempts to parallelize a calculation on a system where OpenMP is not supported.
@@
 The test case consists of 100 points evenly sampled in time during the planet's transit.
-I used physical parameters from the GJ 1214b system \citep{kreidberg14a}.
-The stellar intensity profile is the same for the nonlinear and quadratic models: the nonlinear limb darkening coefficients are $[0.0, 0.7, 0.0, -0.3]$ and the quadratic coefficients are $[0.1, 0.3]$.
-For this case, reaching a truncation error of 1.0 ppm via integration is just 2.5 times slower than the analytic calculation.
-FIXME add what the power law relation for timing is.
+I used physical parameters for the transiting planet GJ\,1214b \citep{kreidberg14a}.
+The stellar intensity profile is the same for the nonlinear and quadratic models: the nonlinear limb darkening coefficients are $(0.0, 0.7, 0.0, -0.3)$ and the quadratic coefficients are $(0.1, 0.3)$.
+For this case, reaching a truncation error of 1.0 ppm via integration is just 3 times slower than the analytic calculation.
+Decreasing the truncation error by a factor of 10 increases the computation time by a factor of three.
 
@@
 I explored using an analytic model to calculate transit light curves for the four-parameter nonlinear limb darkening profile.
-The analytic solution for nonlinear limb darkening was presented in \cite{mandel02}, but it is not used in any published software packages; the original code provided by \cite{mandel02} uses numeric integration.
-I tested a Python/C implementation of the \cite{mandel02} algorithm and found that the new integration scheme in \S\,\ref{section:algorithm} is over 20 times faster (for a truncation error tolerance of 1 ppm).
+The analytic solution for nonlinear limb darkening was presented in \cite{mandel02}, but it is not used in any published software packages.
+The original code provided by \cite{mandel02} uses a numeric integration scheme that is 20 times slower than the algorithm presented in \S\,\ref{section:algorithm} (for an error tolerance of 1 ppm).
 
@@
 \cite{colavecchia04} provide a Fortran library for computing Appell F1.
-I used this library to implement the analytic model for nonlinear limb darkening but did not achieve good results.
-The returned Appell F1 values are not accurate for all input parameters (based on a comparison with \texttt{Mathematica} and the pure Python library \texttt{mpmath}).
+I used this library to implement the analytic model for nonlinear limb darkening.
+However, the returned Appell F1 values are not accurate for all input parameters (based on a comparison with \texttt{Mathematica} and the pure Python library \texttt{mpmath}).
 Even for cases where F1 is correct, the computation is FIXME times slower than numeric integration (for an error tolerance of FIXME).
@@
 \section{Summary}
-%In this paper I introduced the \texttt{batman} package.
-%in preparation for ultra-precise light curves from future telescopes.
-%now have transit model that is good enough for JWST
+\label{section:summary}
+I introduced a new algorithm for computing transit light curves for any radially symmetric stellar limb darkening law.
+I also described the Python package \texttt{batman}, a fast and versatile code to generate transit light curve models.
+\texttt{batman} can calculate light curves with accuracy better than 0.0001 ppm, ensuring that we are prepared for the extraordinarily precise data we anticipate from upcoming facilities.
+(FIXME don't repeat abstract)
 

### 2015-07-23 c_src/_quadratic_ld.c
use ellpic_bulirsch instead of rj

@@ -349,7 +349,7 @@ double rj(double x, double y, double z, double p)
     The complete elliptical integral of the third kind
 
  -- translated from the ellpic_bulirsch.pro routine from EXOFAST
- -- Eastman et al. 2013, PASP 125, 83
+    (Eastman et al. 2013, PASP 125, 83) by Laura Kreidberg (7/22/15)
 */
 
 double ellpic_bulirsch(double n, double k)

### 2015-07-23 manuscript/figs/fig3.py
use arrow for analytic model

@@ -28,7 +28,7 @@ def wrapper(func, *args, **kwargs):
     return wrapped
 
 rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
-#matplotlib.rcParams.update({'font.size':14})
+matplotlib.rcParams.update({'font.size':11})
 plt.figure(figsize=(4,4))
 
 params = batman.TransitParams()	        #object to store transit parameters
@@ -48,22 +48,21 @@ params.u = [0.1, 0.3]       	#limb darkening coefficients
 m = batman.TransitModel(params, t)      #initializes model
 wrapped = wrapper(m.LightCurve, params)
 time = timeit.timeit(wrapped,number=100)/100.
-plt.axvline(time, color='0.5', linestyle='dashed')
+#plt.axvline(time, color='0.5', linestyle='dashed')
+plt.gca().annotate("", xy=(time, .1), xycoords='data', xytext=(time, 2), textcoords='data', arrowprops=dict(arrowstyle="->")) 
+
 print("quadratic", time)
 
 params.limb_dark = "nonlinear"          #limb darkening model
 params.u = [0.0, 0.7, 0.0, -0.3]       	#limb darkening coefficients
 m = batman.TransitModel(params, t)      #initializes model
 flux = m.LightCurve(params)		#calculates light curve
-#plt.plot(t, flux)
-#plt.show()
 
 #generates Figure FIXME: max err as a function of function call time
-#u = [0., 0.7, 0.0, -0.3]
 n = 100
 ts = []
 errs = []
-fac = np.logspace(-2.2, 0.3, n)
+fac = np.logspace(-2, 0.3, n)
 for i in range(n):
 	m.set_fac(fac[i])
 	wrapped = wrapper(m.LightCurve, params)
@@ -81,7 +80,7 @@ print(np.max(ts), np.min(ts))
 plt.gca().tick_params('both', length=8, width=1.2, which='major')
 plt.gca().tick_params('both', length=4, width=0.8, which='minor')
 
-plt.xlim((1.0e-5, 2.0e-3))
+plt.xlim((2.0e-5, 1.0e-3))
 plt.ylim((1.0e-1, 300.))
 plt.yscale('log')
 plt.xscale('log')

### 2015-07-23 manuscript/ms.tex
update text and fig3

@@
 
-\title{BATMAN}
+\title{\texttt{batman}: BAsic Transit Model cAlculatioN in Python}
 
@@
 I introduce \texttt{batman}, a Python package for modeling exoplanet transit light curves.
-\texttt{batman} supports calculation of light curves for any radially symmetric stellar limb darkening law, using a new integration algorithm for models that cannot be calculated analytically.
+The \texttt{batman} package supports calculation of light curves for any radially symmetric stellar limb darkening law, using a new integration algorithm for models that cannot be calculated analytically.
 The code uses C extension modules to speed up model calculation and is parallelized with OpenMP.
-(FIXME: say how long it takes to compute a million models).
-Model light curves can be computed with accuracy better than 0.0001 parts per million, ensuring that the community is prepared for the extraordinarily precise transit light curves we anticipate measuring with upcoming facilities.
+For a typical light curve with 100 data points in transit, \texttt{batman} can calculate one million quadratic limb-darkened models in three minutes with a single 1.7 GHz Intel Core i5 processor.
+Models can be computed with accuracy better than 0.001 parts per million, ensuring that the community is prepared for the precise transit light curves we anticipate measuring with upcoming facilities.
+The code is open source and publicly available at \texttt{\blue{\url{https://github.com/lkreidberg/batman}}}.
 \end{abstract}
 
-\keywords{}
+\keywords{methods: data analysis -- methods: numerical}
 
 \section{Introduction}
-When an extrasolar planet transits in front of its host star, a small fraction of the stellar flux is blocked by the planet.
-Precise measurements of stellar brightness as a function of time (light curves) can therefore reveal transiting planets.
-Over a thousand planets have been discovered with this technique \citep[e.g.][]{rowe2015}.
-Transit light curves can be used to characterize the planets in detail, enabling measurements of planetary radius, orbital elements, and even atmospheric temperature structure and composition \citep{seager00, seager03}.
-%This characterization depends on a foundation of accurate transit light curve models.
+Exoplanet transit light curves can be used to characterize planetary systems in detail, enabling measurements of the planet's radius, orbital elements, and even atmospheric temperature structure and composition \citep{seager00, seager03}.
+A number of current and planned observational facilities -- including \emph{K2}, \emph{TESS}, \emph{CHEOPS}, \emph{JWST}, and \emph{PLATO} -- will measure precise transit light curves for thousands of exoplanets that will shed light on planetary formation, evolution, and habitability.
 
-Transit light curve models are a very basic part of the toolkit for exoplanet characterization, but they are not trivial to compute.
+Transit light curve models are a fundamental part of the toolkit for exoplanet characterization, but they are not trivial to compute.
 The model must account for the planet's size and position on the sky, as well as stellar limb darkening, which causes the apparent brightness of the stellar disk to decrease from center to edge.
@@
 
-In this paper, I introduce the Python package \texttt{batman}.
+In this paper, I introduce the open source Python package \texttt{batman}.
 This package enables fast computation of transit light curves for any radially symmetric limb darkening law, and currently supports uniform, linear, quadratic, logarithmic, exponential, and nonlinear limb darkening.
@@
 I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:summary}.
-
-%This package implements an efficient new algorithm (described in \S\,\ref{section:algorithm}) to calculate transit light curve models for any radially symmetric limb darkening profile.
-The transit models are computed with C extension modules written with the Python/C API, and the code is parallelized at the C level with OpenMP.
-The package therefore   mostly widely used programming language
 
@@
 Computation of $\Delta A$ is expensive, but it needs to be calculated relatively few times ($<<10^{6}$ for sub-ppm accuracy).
-This makes the integration faster than one with a simpler area element (e.g., $\Delta A = \Delta x \Delta y$), that requires a much smaller step size to achieve the same accuracy.
+This makes the integration faster than a scheme with a simpler area element (e.g., $\Delta A = \Delta x \Delta y$), that requires a much smaller step size to achieve the same accuracy.
 
@@
 \label{section:batman}
-The Python package \texttt{batman} implements the algorithm described in \S\,\ref{section:algorithm} to calculate model transit light curves.
+The Python package \texttt{batman} implements the algorithm described in \S\,\ref{section:algorithm} and several analytic models to calculate transit light curves.
 \texttt{batman} is an open source project and is being developed on GitHub.
@@
 The uniform, linear, and quadratic models are calculated analytically, with code based on the Fortran routines \texttt{occultquad.f} and \texttt{occultuniform.f} provided by \cite{mandel02}.
-For the analytic models, I follow \cite{eastman13} and use the algorithm from \cite{bulirsch65} to improve the calculation speed and accuracy for elliptic integrals of the third kind.
+For the analytic models, I follow \cite{eastman13} and use the algorithm from \cite{bulirsch65} to improve calculation speed and accuracy for elliptic integrals of the third kind.
 
@@
 The accuracy was better than 0.005 ppm for most input values; however, for the case $r_p - d < \epsilon$, the error in the light curve exceeded 2 ppm.
-By contast, the \cite{bulirsch65} algorithm for elliptic integrals is well-behaved for this case and also faster.
+By contrast, the \cite{bulirsch65} algorithm for elliptic integrals is well-behaved for this case and also faster.
 
@@
 
-The computationally intensive sections of the code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
-The code also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.
+Computationally intensive sections of code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
+\texttt{batman} also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.
 The number of processors is specified by the user.
@@
 
-I tested \texttt{batman}'s performance over a range of typical use cases using a 1.7 GHz Intel Core i5 processor.
+I tested \texttt{batman}'s performance over a range of typical use cases with a 1.7 GHz Intel Core i5 processor.
 In Figure\,\ref{fig:performance}, I show the execution time versus error tolerance for a single transit light curve calculation using a nonlinear intensity profile, compared to the execution time for a quadratic model computed analytically.
@@
 However, the returned Appell F1 values are not accurate for all input parameters (based on a comparison with \texttt{Mathematica} and the pure Python library \texttt{mpmath}).
-Even for cases where F1 is correct, the computation is FIXME times slower than numeric integration (for an error tolerance of FIXME).
+Even for cases where F1 is correct, the computation is over an order of magnitude slower than numeric integration (for an error tolerance of 0.1 ppm).
 I concluded that integration is a faster and easier solution than analytic models for nonlinear limb darkening.
@@
 I introduced a new algorithm for computing transit light curves for any radially symmetric stellar limb darkening law.
-I also described the Python package \texttt{batman}, a fast and versatile code to generate transit light curve models.
-\texttt{batman} can calculate light curves with accuracy better than 0.0001 ppm, ensuring that we are prepared for the extraordinarily precise data we anticipate from upcoming facilities.
-(FIXME don't repeat abstract)
+I also described the open-source Python package \texttt{batman}, a versatile code to generate model light curves.
+Uniform, linear, quadratic, logarithmic, exponential, and four-parameter nonlinear limb darkening laws are currently supported.
+\texttt{batman} uses C extension modules to compute light curves and is parallelized with OpenMP to optimize performance.
+Light curves can be calculated with accuracy better than 0.001 ppm, ensuring that the community is prepared to model the extraordinarily precise data we anticipate from upcoming facilities.
+\texttt{batman} is available at \texttt{\blue{\url{https://github.com/lkreidberg/batman}}} and is also hosted on the Python Package Index under the name \texttt{batman-package}.
 
 \acknowledgments
+I thank Jacob Bean, Kevin Stevenson, Eric Agol, Ethan Kruse, Geert Jan Talens, Thomas Beatty, Brett Morris, and Karl Fogel for their support in developing \texttt{batman}.
+I also thank contributors to SciPy, Matplotlib, and the Python Programming Language for software and services.
+Support for this work was provided by a grant from the National Science Foundation through a Graduate Research Fellowship to L.K.
 

### 2015-07-23 manuscript/figs/fig3.py

@@ -49,7 +49,7 @@ m = batman.TransitModel(params, t)      #initializes model
 wrapped = wrapper(m.LightCurve, params)
 time = timeit.timeit(wrapped,number=100)/100.
 #plt.axvline(time, color='0.5', linestyle='dashed')
-plt.gca().annotate("", xy=(time, .1), xycoords='data', xytext=(time, 2), textcoords='data', arrowprops=dict(arrowstyle="->")) 
+plt.gca().annotate("", xy=(time, .1), xycoords='data', xytext=(time, 1), textcoords='data', arrowprops=dict(arrowstyle="->")) 
 
 print("quadratic", time)
 

### 2015-07-25 c_src/_rsky.c
fix true anomaly calculation

@@ -25,11 +25,8 @@ static PyObject *_rsky(PyObject *self, PyObject *args);
 double getE(double M, double e)	//calculates the eccentric anomaly (see Seager Exoplanets book:  Murray & Correia eqn. 5 -- see section 3)
 {
 	double E = M, eps = 1.0e-7;
-	
-	while(fabs(E - e*sin(E) - M) > eps)
-	{
-		E = E - (E - e*sin(E) - M)/(1.0 - e*cos(E));
-	}
+
+	while(fabs(E - e*sin(E) - M) > eps) E = E - (E - e*sin(E) - M)/(1.0 - e*cos(E));
 	return E;
 }
 
@@ -42,17 +39,17 @@ static PyObject *_rsky(PyObject *self, PyObject *args)
 		(see the section by Murray, and Winn eq. 5).  In the Mandel & Agol 
 		(2002) paper, this quantity is denoted d.
 	*/
-	double ecc, E, inc, a, r, d, f, omega, per, M, n, t0, eps, t;
+	double ecc, E, inc, a, d, f, omega, per, M, n, t0, eps, t, BIGD = 100.;
 	npy_intp i, dims[1];
-	PyArrayObject *ts, *zs;
+	PyArrayObject *ts, *ds;
 
   	if(!PyArg_ParseTuple(args,"Odddddd", &ts, &t0, &per, &a, &inc, &ecc, &omega)) return NULL; 
 
 	dims[0] = PyArray_DIMS(ts)[0]; 
-	zs = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ts));
+	ds = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ts));
 	
 	double *t_array = PyArray_DATA(ts);
-	double *z_array = PyArray_DATA(zs);
+	double *d_array = PyArray_DATA(ds);
 
 	n = 2.*M_PI/per;	// mean motion
 	eps = 1.0e-7;
@@ -61,19 +58,21 @@ static PyObject *_rsky(PyObject *self, PyObject *args)
 	{
 		t = t_array[i];
 
-		if(ecc > 1.e-5)						//calculates f for eccentric orbits
+		if(ecc < 1.0e-5)
+		{
+			f = ((t - t0)/per - (int)((t - t0)/per))*2.*M_PI;			//calculates f for a circular orbit
+		}
+		else
 		{
[…]

### 2015-07-25 manuscript/ms.tex
tweak abstract

@@
 I introduce \texttt{batman}, a Python package for modeling exoplanet transit light curves.
-The \texttt{batman} package supports calculation of light curves for any radially symmetric stellar limb darkening law, using a new integration algorithm for models that cannot be calculated analytically.
+The \texttt{batman} package supports calculation of light curves for any radially symmetric stellar limb darkening law, using a new integration algorithm for models that cannot be quickly calculated analytically.
 The code uses C extension modules to speed up model calculation and is parallelized with OpenMP.

### 2015-07-28 manuscript/ms.tex
add refs

@@
 For a typical light curve with 100 data points in transit, \texttt{batman} can calculate one million quadratic limb-darkened models in three minutes with a single 1.7 GHz Intel Core i5 processor.
-Models can be computed with accuracy better than 0.001 parts per million, ensuring that the community is prepared for the precise transit light curves we anticipate measuring with upcoming facilities.
-The code is open source and publicly available at \texttt{\blue{\url{https://github.com/lkreidberg/batman}}}.
+The same calculation takes seven minutes using the four-parameter nonlinear limb darkening model (computed to 1 ppm accuracy).
+Maximum truncation error for integrated models is an input parameter that can be set as low as 0.001 ppm, ensuring that the community is prepared for the precise transit light curves we anticipate measuring with upcoming facilities.
+The \texttt{batman} package is open source and publicly available at \texttt{\blue{\url{https://github.com/lkreidberg/batman}}}.
 \end{abstract}
@@
 \section{Introduction}
-Exoplanet transit light curves can be used to characterize planetary systems in detail, enabling measurements of the planet's radius, orbital elements, and even atmospheric temperature structure and composition \citep{seager00, seager03}.
-A number of current and planned observational facilities -- including \emph{K2}, \emph{TESS}, \emph{CHEOPS}, \emph{JWST}, and \emph{PLATO} -- will measure precise transit light curves for thousands of exoplanets that will shed light on planetary formation, evolution, and habitability.
+The transit technique has revolutionized the study of exoplanetary systems.
+Thanks largely to the \emph{Kepler} mission, over a thousand planets have been discovered with this method \citep{rowe15}.
+These discoveries have yielded transformative constraints on planet occurrence rates over a wide range of planet sizes, orbital periods, and host star properties \citep{borucki11, youdin11, howard12,fressin13, dong13, morton14}.
+The have also enabled the first estimates of the occurrence of habitable planets \citep{traub12, dressing13, petigura13, kopparapu13, foremanmackey14, dressing15}.
+Transit light curves can even reveal planets' atmospheric temperature structure and composition \citep[e.g.][]{seager00, lecavelier08, sing11, deming13, kreidberg14b, knutson14a, fraine14, kreidberg15}.
+A number of current and planned observational facilities -- including \emph{K2}, \emph{TESS}, \emph{CHEOPS}, \emph{JWST}, and \emph{PLATO} -- will measure precise transit light curves for thousands of exoplanets that will further transform our understanding of planet formation, evolution, and habitability.
 
-Transit light curve models are a fundamental part of the toolkit for exoplanet characterization, but they are not trivial to compute.
-The model must account for the planet's size and position on the sky, as well as stellar limb darkening, which causes the apparent brightness of the stellar disk to decrease from center to edge.
+Light curve models are a fundamental tool for transiting exoplanet science, but they are not trivial to compute quickly and accurately.
+Accurate calculation is challenging because the model must account for the planet's size and position on the sky, as well as stellar limb darkening, which causes the apparent brightness of the stellar disk to decrease from center to edge.
 The  stellar intensity profile can be fit with several functional forms, including a linear limb darkening law \citep{schwarzschild06}, quadratic \citep{kopal50}, square-root \citep{diaz-cordoves92}, logarithmic \citep{klinglesmith70}, exponential \citep{claret03}, and four-parameter nonlinear \citep{claret00}.
@@
 Other profiles do not have analytic solutions, and models must be calculated by numeric integration of the stellar intensity over the disk of the planet.
+In addition, speed is an important consideration because a large number of models must typically be calculated to make a robust estimation of transit parameters and their uncertainties.
 
@@
 In this paper, I introduce the open source Python package \texttt{batman}.
-This package enables fast computation of transit light curves for any radially symmetric limb darkening law, and currently supports uniform, linear, quadratic, logarithmic, exponential, and nonlinear limb darkening.
+This package is based on code that was used to model high-precision light curves obtained for atmosphere characterization \citep{kreidberg14a, kreidberg14b, stevenson14a, kreidberg15}.
+The \texttt{batman} package enables fast computation of transit light curves for any radially symmetric limb darkening law, and currently supports uniform, linear, quadratic, logarithmic, exponential, and nonlinear limb darkening.
 Light curves for the first three of these are calculated analytically based on the formalism from \cite{mandel02}.
@@
 
+For the case $\pi <= \omega + |f| <= 2\pi$, $d$ is set to a large value so that the code does not produce a transit model when the planet is being eclipsed by the star.
+To model a secondary eclipse, use a uniform limb darkening profile and set $t_0$ equal to the mid-eclipse time.
+
 \subsection{Accuracy}
@@
 The stellar intensity profile is the same for the nonlinear and quadratic models: the nonlinear limb darkening coefficients are $(0.0, 0.7, 0.0, -0.3)$ and the quadratic coefficients are $(0.1, 0.3)$.
-For this case, reaching a truncation error of 1.0 ppm via integration is just 3 times slower than the analytic calculation.
+For this case, reaching a truncation error of 1.0 ppm via integration is just two times slower than the analytic calculation.
 Decreasing the truncation error by a factor of 10 increases the computation time by a factor of three.
@@
 I used this library to implement the analytic model for nonlinear limb darkening.
-However, the returned Appell F1 values are not accurate for all input parameters (based on a comparison with \texttt{Mathematica} and the pure Python library \texttt{mpmath}).
+However, the returned Appell F1 values are not accurate for all input parameters, based on a comparison with \texttt{Mathematica} and the pure Python library \texttt{mpmath}.
 Even for cases where F1 is correct, the computation is over an order of magnitude slower than numeric integration (for an error tolerance of 0.1 ppm).
@@
 I also thank contributors to SciPy, Matplotlib, and the Python Programming Language for software and services.
-Support for this work was provided by a grant from the National Science Foundation through a Graduate Research Fellowship to L.K.
+Support for this work was provided by a grant from the National Science Foundation through a Graduate Research Fellowship to the author.
 

### 2015-07-28 c_src/_quadratic_ld.c
remove unused rf(), rc(), and rj() functions

@@ -27,12 +27,9 @@
 #define MIN(x, y) (((x) < (y)) ? (x) : (y))
 #define MAX(x, y) (((x) > (y)) ? (x) : (y))
 
-double rc(double x, double y);
-double rj(double x, double y, double z, double p);
 double ellpic_bulirsch(double n, double k);
 double ellec(double k);
 double ellk(double k);
-double rf(double x, double y, double z);
 static PyObject *_quadratic_ld(PyObject *self, PyObject *args);
 
 static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
@@ -64,9 +61,20 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
 	nz = (int)dims;
 
-
 	double *f_array = PyArray_DATA(flux);
 	double *z_array = PyArray_DATA(zs);
+
+	/*
+		NOTE:  the safest way to access numpy arrays is to use the PyArray_GETITEM and PyArray_SETITEM functions.
+		Here we use a trick for faster access and more convenient access, where we set a pointer to the 
+		beginning of the array with the PyArray_DATA (e.g., f_array) and access elements with e.g., f_array[i].
+		Success of this operation depends on the numpy array storing data in blocks equal in size to a C double.
+		If you run into trouble along these lines, I recommend changing the array access to something like:
+			d = PyFloat_AsDouble(PyArray_GETITEM(zs, PyArray_GetPtr(zs, &i))); 
+		where zs is a numpy array object.
+		Laura Kreidberg 07/2015
+	*/
+
 
[…]

### 2015-07-28 c_src/_custom_ld.c
make notation consistent with paper

@@ -50,15 +50,15 @@ static PyObject *_custom_ld(PyObject *self, PyObject *args)
 	double rprs, d, fac, A_i, r, I, dr, A_f, r_in, r_out, delta, c1, c2, c3, c4, c5, c6;
 	int nthreads;
 	npy_intp i, dims[1];
-	PyArrayObject *zs, *flux;
+	PyArrayObject *ds, *flux;
 
-  	if(!PyArg_ParseTuple(args,"Oddddddddi", &zs, &rprs, &c1, &c2, &c3, &c4, &c5, &c6, &fac, &nthreads)) return NULL; //parses input arguments
+  	if(!PyArg_ParseTuple(args,"Oddddddddi", &ds, &rprs, &c1, &c2, &c3, &c4, &c5, &c6, &fac, &nthreads)) return NULL; //parses input arguments
 	
-	dims[0] = PyArray_DIMS(zs)[0]; 
-	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
+	dims[0] = PyArray_DIMS(ds)[0]; 
+	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
 
 	double *f_array = PyArray_DATA(flux);
-	double *z_array = PyArray_DATA(zs);
+	double *d_array = PyArray_DATA(ds);
 
 	/*
 		NOTE:  the safest way to access numpy arrays is to use the PyArray_GETITEM and PyArray_SETITEM functions.
@@ -66,8 +66,8 @@ static PyObject *_custom_ld(PyObject *self, PyObject *args)
 		beginning of the array with the PyArray_DATA (e.g., f_array) and access elements with e.g., f_array[i].
 		Success of this operation depends on the numpy array storing data in blocks equal in size to a C double.
 		If you run into trouble along these lines, I recommend changing the array access to something like:
-			d = PyFloat_AsDouble(PyArray_GETITEM(zs, PyArray_GetPtr(zs, &i))); 
-		where zs is a numpy array object.
+			d = PyFloat_AsDouble(PyArray_GETITEM(ds, PyArray_GetPtr(ds, &i))); 
+		where ds is a numpy array object.
 
 
 		Laura Kreidberg 07/2015
[…]

### 2015-07-28 c_src/_exponential_ld.c

@@ -58,14 +58,14 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 	npy_intp i, dims[1];
 	double dr, A_f, r_in, r_out, delta, c1, c2;
 	
-	PyArrayObject *zs, *flux;
-  	if(!PyArg_ParseTuple(args,"Oddddi", &zs, &rprs, &c1, &c2, &fac, &nthreads)) return NULL;
+	PyArrayObject *ds, *flux;
+  	if(!PyArg_ParseTuple(args,"Oddddi", &ds, &rprs, &c1, &c2, &fac, &nthreads)) return NULL;
 
-	dims[0] = PyArray_DIMS(zs)[0]; 
-	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
+	dims[0] = PyArray_DIMS(ds)[0]; 
+	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
 
 	double *f_array = PyArray_DATA(flux);
-	double *z_array = PyArray_DATA(zs);
+	double *d_array = PyArray_DATA(ds);
 
 	/*
 		NOTE:  the safest way to access numpy arrays is to use the PyArray_GETITEM and PyArray_SETITEM functions.
@@ -73,8 +73,8 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 		beginning of the array with the PyArray_DATA (e.g., f_array) and access elements with e.g., f_array[i].
 		Success of this operation depends on the numpy array storing data in blocks equal in size to a C double.
 		If you run into trouble along these lines, I recommend changing the array access to something like:
-			d = PyFloat_AsDouble(PyArray_GETITEM(zs, PyArray_GetPtr(zs, &i))); 
-		where zs is a numpy array object.
+			d = PyFloat_AsDouble(PyArray_GETITEM(ds, PyArray_GetPtr(ds, &i))); 
+		where ds is a numpy array object.
 
 
 		Laura Kreidberg 07/2015
@@ -91,7 +91,7 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 	#endif
 	for(i = 0; i < dims[0]; i++)
 	{
-		d = z_array[i];
+		d = d_array[i];
 		r_in = MAX(d - rprs, 0.);					//lower bound for integration
[…]

### 2015-07-28 c_src/_logarithmic_ld.c

@@ -57,14 +57,14 @@ static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 	npy_intp i, dims[1];
 	double dr, A_f, r_in, r_out, delta, c1, c2;
 	
-	PyArrayObject *zs, *flux;
-  	if(!PyArg_ParseTuple(args,"Oddddi", &zs, &rprs, &c1, &c2, &fac, &nthreads)) return NULL;
+	PyArrayObject *ds, *flux;
+  	if(!PyArg_ParseTuple(args,"Oddddi", &ds, &rprs, &c1, &c2, &fac, &nthreads)) return NULL;
 
-	dims[0] = PyArray_DIMS(zs)[0]; 
-	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
+	dims[0] = PyArray_DIMS(ds)[0]; 
+	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
 
 	double *f_array = PyArray_DATA(flux);
-	double *z_array = PyArray_DATA(zs);
+	double *d_array = PyArray_DATA(ds);
 
 	/*
 		NOTE:  the safest way to access numpy arrays is to use the PyArray_GETITEM and PyArray_SETITEM functions.
@@ -72,8 +72,8 @@ static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 		beginning of the array with the PyArray_DATA (e.g., f_array) and access elements with e.g., f_array[i].
 		Success of this operation depends on the numpy array storing data in blocks equal in size to a C double.
 		If you run into trouble along these lines, I recommend changing the array access to something like:
-			d = PyFloat_AsDouble(PyArray_GETITEM(zs, PyArray_GetPtr(zs, &i))); 
-		where zs is a numpy array object.
+			d = PyFloat_AsDouble(PyArray_GETITEM(ds, PyArray_GetPtr(ds, &i))); 
+		where ds is a numpy array object.
 
 
 		Laura Kreidberg 07/2015
@@ -90,7 +90,7 @@ static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 	#endif
 	for(i = 0; i < dims[0]; i++)
 	{
-		d = z_array[i];
+		d = d_array[i];
[…]

### 2015-07-28 c_src/_nonlinear_ld.c

@@ -57,14 +57,14 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 	npy_intp i, dims[1];
 	double dr, A_f, r_in, r_out, delta, c1, c2, c3, c4;
 	
-	PyArrayObject *zs, *flux;
-  	if(!PyArg_ParseTuple(args,"Oddddddi", &zs, &rprs, &c1, &c2, &c3, &c4, &fac, &nthreads)) return NULL;
+	PyArrayObject *ds, *flux;
+  	if(!PyArg_ParseTuple(args,"Oddddddi", &ds, &rprs, &c1, &c2, &c3, &c4, &fac, &nthreads)) return NULL;
 
-	dims[0] = PyArray_DIMS(zs)[0]; 
-	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
+	dims[0] = PyArray_DIMS(ds)[0]; 
+	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
 
 	double *f_array = PyArray_DATA(flux);
-	double *z_array = PyArray_DATA(zs);
+	double *d_array = PyArray_DATA(ds);
 
 	/*
 		NOTE:  the safest way to access numpy arrays is to use the PyArray_GETITEM and PyArray_SETITEM functions.
@@ -72,8 +72,8 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 		beginning of the array with the PyArray_DATA (e.g., f_array) and access elements with e.g., f_array[i].
 		Success of this operation depends on the numpy array storing data in blocks equal in size to a C double.
 		If you run into trouble along these lines, I recommend changing the array access to something like:
-			d = PyFloat_AsDouble(PyArray_GETITEM(zs, PyArray_GetPtr(zs, &i))); 
-		where zs is a numpy array object.
+			d = PyFloat_AsDouble(PyArray_GETITEM(ds, PyArray_GetPtr(ds, &i))); 
+		where ds is a numpy array object.
 
 
 		Laura Kreidberg 07/2015
@@ -90,7 +90,7 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 	#endif
 	for(i = 0; i < dims[0]; i++)
 	{
-		d = z_array[i];
+		d = d_array[i];
[…]

### 2015-07-28 c_src/_quadratic_ld.c

@@ -36,33 +36,33 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 {
 /*	Input: *************************************
  
-	 zs   	impact parameter in units of rs
+	 ds   	array of impact parameters in units of rs
 	 c1   	linear    limb-darkening coefficient (gamma_1 in Mandel & Agol 2002)
 	 c2   	quadratic limb-darkening coefficient (gamma_2)
 	 p    	occulting star size in units of rs
 	
 	 Output: ***********************************
 	
-	 flux 	fraction of flux at each zs for a limb-darkened source
+	 flux 	fraction of flux at each ds for a limb-darkened source
 	
 	 Limb darkening has the form:
 	 I(r) = [1 - c1 * (1 - sqrt(1 - (r/rs)^2)) - c2*(1 - sqrt(1 - (r/rs)^2))^2]/(1 - c1/3 - c2/6)/pi
 */
 	int nz, nthreads;
 	double c1, c2, p, *mu,  *lambdad, *etad, \
-		*lambdae, lam, x1, x2, x3, z, omega, kap0 = 0.0, kap1 = 0.0, \
+		*lambdae, lam, x1, x2, x3, d, omega, kap0 = 0.0, kap1 = 0.0, \
 		q, Kk, Ek, Pk, n;
-	PyArrayObject *zs, *flux;
+	PyArrayObject *ds, *flux;
 	npy_intp i, dims[1];
 
-  	if(!PyArg_ParseTuple(args,"Odddi", &zs, &p, &c1, &c2, &nthreads)) return NULL;
+  	if(!PyArg_ParseTuple(args,"Odddi", &ds, &p, &c1, &c2, &nthreads)) return NULL;
 
-	dims[0] = PyArray_DIMS(zs)[0]; 
-	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
+	dims[0] = PyArray_DIMS(ds)[0]; 
[…]

### 2015-07-28 c_src/_uniform_ld.c

@@ -27,18 +27,18 @@
 static PyObject *_uniform_ld(PyObject *self, PyObject *args)
 {
 	int nthreads;
-	double z, p, kap0, kap1;
+	double d, p, kap0, kap1;
 
-	PyArrayObject *zs, *flux;
+	PyArrayObject *ds, *flux;
 	npy_intp i, dims[1];
 	
-  	if(!PyArg_ParseTuple(args, "Odi", &zs, &p, &nthreads)) return NULL;		//parses function input
+  	if(!PyArg_ParseTuple(args, "Odi", &ds, &p, &nthreads)) return NULL;		//parses function input
 
-	dims[0] = PyArray_DIMS(zs)[0]; 
-	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(zs));	//creates numpy array to store return flux values
+	dims[0] = PyArray_DIMS(ds)[0]; 
+	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
 	
 	double *f_array = PyArray_DATA(flux);
-	double *z_array = PyArray_DATA(zs);
+	double *d_array = PyArray_DATA(ds);
 
 	if(fabs(p - 0.5) < 1.e-3) p = 0.5;
 
@@ -47,20 +47,20 @@ static PyObject *_uniform_ld(PyObject *self, PyObject *args)
 	#endif
 
 	#if defined (_OPENMP)
-	#pragma omp parallel for private(z, kap1, kap0)
+	#pragma omp parallel for private(d, kap1, kap0)
 	#endif
 	for(i=0; i<dims[0]; i++)
 	{
-		z = z_array[i]; 				// separation of centers
+		d = d_array[i]; 				// separation of centers
 		
-		if(z >= 1. + p) f_array[i] = 1.;		//no overlap
-		if(p >= 1. && z <= p - 1.) f_array[i] = 0.;	//total eclipse of the star
-		else if(z <= 1. - p) f_array[i] = 1. - p*p;	//planet is fully in transit
+		if(d >= 1. + p) f_array[i] = 1.;		//no overlap
+		if(p >= 1. && d <= p - 1.) f_array[i] = 0.;	//total eclipse of the star
[…]

### 2015-07-28 c_src/_custom_ld.c
make notation consistent with paper

@@ -29,25 +29,25 @@
 
 static PyObject *_custom_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double c1, double c2, double c3, double c4, double c5, double c6);
+double intensity(double x, double c1, double c2, double c3, double c4, double c5, double c6);
 
-double area(double d, double r, double R)
+double area(double d, double x, double R)
 {
 	/*
-	Returns area of overlapping circles with radii r and R; separated by a distance d
+	Returns area of overlapping circles with radii x and R; separated by a distance d
 	*/
-	double arg1 = (d*d + r*r - R*R)/(2.*d*r);
-	double arg2 = (d*d + R*R - r*r)/(2.*d*R); 
-	double arg3 = MAX((-d + r + R)*(d + r - R)*(d - r + R)*(d + r + R), 0.);
+	double arg1 = (d*d + x*x - R*R)/(2.*d*x);
+	double arg2 = (d*d + R*R - x*x)/(2.*d*R); 
+	double arg3 = MAX((-d + x + R)*(d + x - R)*(d - x + R)*(d + x + R), 0.);
 
-	if(r <= R - d) return M_PI*r*r;							//planet completely overlaps stellar circle
-	else if(r >= R + d) return M_PI*R*R;						//stellar circle completely overlaps planet
-	else return r*r*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);			//partial overlap
+	if(x <= R - d) return M_PI*x*x;							//planet completely overlaps stellar circle
+	else if(x >= R + d) return M_PI*R*R;						//stellar circle completely overlaps planet
+	else return x*x*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);			//partial overlap
 }
 
 static PyObject *_custom_ld(PyObject *self, PyObject *args)
 {
-	double rprs, d, fac, A_i, r, I, dr, A_f, r_in, r_out, delta, c1, c2, c3, c4, c5, c6;
+	double rprs, d, fac, A_i, x, I, dx, A_f, x_in, x_out, delta, c1, c2, c3, c4, c5, c6;
 	int nthreads;
 	npy_intp i, dims[1];
 	PyArrayObject *ds, *flux;
@@ -78,37 +78,37 @@ static PyObject *_custom_ld(PyObject *self, PyObject *args)
 	#endif
 
 	#if defined (_OPENMP)
[…]

### 2015-07-28 c_src/_exponential_ld.c

@@ -29,34 +29,34 @@
 
 static PyObject *_exponential_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double c1, double c2, double norm)
+double intensity(double x, double c1, double c2, double norm)
 {
-	if(r > 0.99995) r = 0.99995;
-	if(r < 0.00005) r = 0.00005;
-	double mu = sqrt(1. - r*r);
+	if(x > 0.99995) x = 0.99995;
+	if(x < 0.00005) x = 0.00005;
+	double mu = sqrt(1. - x*x);
 	return (1. - c1*(1. - mu) - c2/(1. - exp(mu)))/norm; 
 }
 
-double area(double d, double r, double R)
+double area(double d, double x, double R)
 {
 	/*
 	Returns area of overlapping circles with radii r and R; separated by a distance d
 	*/
-	double arg1 = (d*d + r*r - R*R)/(2.*d*r); 	
-	double arg2 = (d*d + R*R - r*r)/(2.*d*R); 
-	double arg3 = MAX((-d + r + R)*(d + r - R)*(d - r + R)*(d + r + R), 0.);
+	double arg1 = (d*d + x*x - R*R)/(2.*d*x); 	
+	double arg2 = (d*d + R*R - x*x)/(2.*d*R); 
+	double arg3 = MAX((-d + x + R)*(d + x - R)*(d - x + R)*(d + x + R), 0.);
 
-	if(r <= R - d) return M_PI*r*r;						//planet completely overlaps stellar circle
-	else if(r >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
-	else return r*r*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
+	if(x <= R - d) return M_PI*x*x;						//planet completely overlaps stellar circle
+	else if(x >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
+	else return x*x*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
 }
 
 static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 {
-	double rprs, d, fac, A_i, r, I; 
+	double rprs, d, fac, A_i, x, I; 
 	int nthreads;
 	npy_intp i, dims[1];
-	double dr, A_f, r_in, r_out, delta, c1, c2;
[…]

### 2015-07-28 c_src/_logarithmic_ld.c

@@ -29,33 +29,33 @@
 
 static PyObject *_logarithmic_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double c1, double c2, double norm)
+double intensity(double x, double c1, double c2, double norm)
 {
-	if(r > 0.99995) r = 0.99995;
-	double mu = sqrt(1. - r*r);
+	if(x > 0.99995) x = 0.99995;
+	double mu = sqrt(1. - x*x);
 	return (1. - c1*(1.-mu) - c2*mu*log(mu))/norm; 
 }
 
-double area(double d, double r, double R)
+double area(double d, double x, double R)
 {
 	/*
 	Returns area of overlapping circles with radii r and R; separated by a distance d
 	*/
-	double arg1 = (d*d + r*r - R*R)/(2.*d*r); 	
-	double arg2 = (d*d + R*R - r*r)/(2.*d*R); 
-	double arg3 = MAX((-d + r + R)*(d + r - R)*(d - r + R)*(d + r + R), 0.);
+	double arg1 = (d*d + x*x - R*R)/(2.*d*x); 	
+	double arg2 = (d*d + R*R - x*x)/(2.*d*R); 
+	double arg3 = MAX((-d + x + R)*(d + x - R)*(d - x + R)*(d + x + R), 0.);
 
-	if(r <= R - d) return M_PI*r*r;						//planet completely overlaps stellar circle
-	else if(r >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
-	else return r*r*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
+	if(x <= R - d) return M_PI*x*x;						//planet completely overlaps stellar circle
+	else if(x >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
+	else return x*x*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
 }
 
 static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 {
-	double rprs, d, fac, A_i, r, I; 
+	double rprs, d, fac, A_i, x, I; 
 	int nthreads;
 	npy_intp i, dims[1];
-	double dr, A_f, r_in, r_out, delta, c1, c2;
+	double dx, A_f, x_in, x_out, delta, c1, c2;
 	
 	PyArrayObject *ds, *flux;
[…]

### 2015-07-28 c_src/_nonlinear_ld.c

@@ -29,33 +29,33 @@
 
 static PyObject *_nonlinear_ld(PyObject *self, PyObject *args);
 
-double intensity(double r, double c1, double c2, double c3, double c4, double norm)
+double intensity(double x, double c1, double c2, double c3, double c4, double norm)
 {
-	if(r > 0.99995) r = 0.99995;
-	double sqrtmu = pow(1. - r*r,0.25);
+	if(x > 0.99995) x = 0.99995;
+	double sqrtmu = pow(1. - x*x,0.25);
 	return (1. - c1*(1. - sqrtmu) - c2*(1. - pow(sqrtmu,2.)) - c3*(1. - pow(sqrtmu, 3.)) - c4*(1. - pow(sqrtmu,4.)))/norm; 	
 }
 
-double area(double d, double r, double R)
+double area(double d, double x, double R)
 {
 	/*
 	Returns area of overlapping circles with radii r and R; separated by a distance d
 	*/
-	double arg1 = (d*d + r*r - R*R)/(2.*d*r); 	
-	double arg2 = (d*d + R*R - r*r)/(2.*d*R); 
-	double arg3 = MAX((-d + r + R)*(d + r - R)*(d - r + R)*(d + r + R), 0.);
+	double arg1 = (d*d + x*x - R*R)/(2.*d*x); 	
+	double arg2 = (d*d + R*R - x*x)/(2.*d*R); 
+	double arg3 = MAX((-d + x + R)*(d + x - R)*(d - x + R)*(d + x + R), 0.);
 
-	if(r <= R - d) return M_PI*r*r;						//planet completely overlaps stellar circle
-	else if(r >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
-	else return r*r*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
+	if(x <= R - d) return M_PI*x*x;						//planet completely overlaps stellar circle
+	else if(x >= R + d) return M_PI*R*R;					//stellar circle completely overlaps planet
+	else return x*x*acos(arg1) + R*R*acos(arg2) - 0.5*sqrt(arg3);		//partial overlap
 }
 
 static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 {
-	double rprs, d, fac, A_i, r, I; 
+	double rprs, d, fac, A_i, x, I; 
 	int nthreads;
[…]

### 2015-07-28 c_src/_quadratic_ld.c
bug fix to case 5.1

@@ -136,10 +136,10 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 			lambdae[i] = p*p;
 		}
 		//edge of the occulting star lies at the origin
-		if(fabs(d - p) < 1.0e-3*(d + p))		
+		if(fabs(d - p) < 1.0e-4*(d + p))		
 		{
 			d = p;
-		//	printf("zone 5\n");
+			//printf("zone 5\n");
 			if(p == 0.5)	
 			{
 				//printf("zone 6\n");
@@ -159,11 +159,11 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 				 	 	(32.0*pow(p, 4.0) - 20.0*p*p + 3.0)/9.0/M_PI/p*Kk;
 				etad[i] = 1.0/2.0/M_PI*(kap1 + p*p*(p*p + 2.0*d*d)*kap0 -  \
 				              	(1.0 + 5.0*p*p + d*d)/4.0*sqrt((1.0 - x1)*(x2 - 1.0)));
-				continue;
+			//	continue;
 			}
 			else if(d<0.5)	
 			{
-			//	printf("zone 5.2\n");
+				//printf("zone 5.2\n");
 				lam = 0.50*M_PI;
 				q = 2.0*p;
 				Kk = ellk(q);
@@ -278,7 +278,6 @@ double ellpic_bulirsch(double n, double k)
 		{
 			return 0.5*M_PI*(c*m0+d)/(m0*(m0+p));
 		}
-		//printf("nit %i\n", nit);
 		nit++;
 	}
 	printf("Convergence failure in ellpic_bulirsch\n");

### 2015-07-28 batman/tests.py
test ld models, error tolerance, and multithreading

@@ -18,42 +18,83 @@ from __future__ import print_function
 import numpy as np
 import math
 import matplotlib.pyplot as plt
+import timeit
 from .transitmodel import *
+from .openmp import *
+
+def wrapper(func, *args, **kwargs):
+    def wrapped():
+        return func(*args, **kwargs)
+    return wrapped
 
 def test():
-	print("Starting tests...")
+	print("\nStarting tests...\n")
 	failures = 0
 
 	params = TransitParams()
 	params.t0 = 0.
 	params.per = 1.0
 	params.rp = 0.1
-	params.a = 15.23
-	params.inc = 1.555*180./math.pi
+	params.a = 15.0
+	params.inc = 90.
 	params.ecc = 0.
 	params.w = 90. 
+	t = np.linspace(-0.01, 0.01, 1000)
+
+	u = [[], [0.1], [0.1,0.3], [0.1,-0.3], [0.1,-0.03], [0.0, 0.7, 0.0, -0.3]]
+	ld = ["uniform", "linear", "quadratic", "exponential", "logarithmic", "nonlinear"]
+	
+	#Testing different limb darkening models
+	for i in range(len(u)):
+		params.u = u[i]
+		params.limb_dark = ld[i]
+		print("Testing "+ld[i]+" limb darkening model...\t")
+		m = TransitModel(params, t)
+		lc = m.light_curve(params)
+		if np.min(lc) <= 1.-params.rp**2: print("\ttest passed")
+		else: 
+			print("test failed")
+			failures += 1
+
+	#Testing truncation error tolerance
+	params.u = [0.1, 0.3]
+	params.limb_dark = "quadratic"
+	m = TransitModel(params, t)
+	f_ref = m.light_curve(params)
+
 	params.u = np.array([0.0, 0.7, 0.0, -0.3])
 	params.limb_dark = "nonlinear"
 
-	t = np.linspace(0.01, 0.05, 1000)
-	err_max = 0.7
-	
-	m = TransitModel(params, t, err_max)
-	nonlinear_lc = m.LightCurve(params)
-	err = m.calc_err()
-	if err > err_max: failures += 1
+	print("\nTesting error tolerance...")
+	max_err = np.logspace(1, -2, 4) 
+	for i in range(len(max_err)):
+		m = TransitModel(params, t, max_err = max_err[i])
+		f = m.light_curve(params)
[…]

### 2015-07-28 docs/demo.py
change to light_curve() method

@@ -39,10 +39,10 @@ t = np.linspace(-0.02, 0.02, 1000)           #times at which to calculate light 
 
 m = batman.TransitModel(params, t)
 
-flux = m.LightCurve(params)
+flux = m.light_curve(params)
 
 params.rp = 0.11                             #updates the planet radius
-new_flux = m.LightCurve(params)              #recalculates model light curve
+new_flux = m.light_curve(params)              #recalculates model light curve
 
 plt.plot(t*24., flux, label = "rp = 0.1")
 plt.plot(t*24., new_flux, label = "rp = 0.11")
@@ -62,7 +62,7 @@ for i in range(4):
 	params.limb_dark = ld_options[i]             #specifies the limb darkening profile
        	params.u = ld_coefficients[i]		     #updates limb darkening coefficients
 	m = batman.TransitModel(params, t)  	     #initializes model with new limb darkening
-	flux = m.LightCurve(params)		     #calculates light curve
+	flux = m.light_curve(params)		     #calculates light curve
 	plt.plot(t*24., flux, label = ld_options[i])
 
 plt.xlim((-0.3, 0.3))

### 2015-07-28 docs/quickstart.py

@@ -38,7 +38,7 @@ params.limb_dark = "quadratic"          #limb darkening model
    
 t = np.linspace(-0.025, 0.025, 1000)    	#times at which to calculate light curve	
 m = batman.TransitModel(params, t)	        #initializes model
-flux = m.LightCurve(params)
+flux = m.light_curve(params)
 
 plt.plot(t, flux)
 plt.xlabel("Time from central transit (days)")

### 2015-07-28 docs/tutorial.py

@@ -38,12 +38,12 @@ params.u = [0.5, 0.1, 0.1, -0.1]       	#limb darkening coefficients
 t = np.linspace(-0.025, 0.025, 1000)  	#times at which to calculate light curve	
 m = batman.TransitModel(params, t)      #initializes model
 
-flux = m.LightCurve(params)		#calculates light curve
+flux = m.light_curve(params)		#calculates light curve
 
 radii = np.linspace(0.09, 0.11, 20)
 for r in radii:
 	params.rp = r				#updates planet radius
-	new_flux = m.LightCurve(params)         #recalculates light curve
+	new_flux = m.light_curve(params)        #recalculates light curve
 	plt.plot(t, new_flux)
 
 plt.xlabel("Time from central transit (days)")
@@ -59,10 +59,10 @@ ld_coefficients = [[], [0.3], [0.1, 0.3], [0.5, 0.1, 0.1, -0.1]]
 plt.figure()
 
 for i in range(4):
-	params.limb_dark = ld_options[i]             #specifies the limb darkening profile
-	params.u = ld_coefficients[i]	         #updates limb darkening coefficients
-	m = batman.TransitModel(params, t)	         #initializes the model
-	flux = m.LightCurve(params)		         #calculates light curve
+	params.limb_dark = ld_options[i]             	#specifies the limb darkening profile
+	params.u = ld_coefficients[i]	         	#updates limb darkening coefficients
+	m = batman.TransitModel(params, t)	        #initializes the model
+	flux = m.light_curve(params)		        #calculates light curve
 	plt.plot(t, flux, label = ld_options[i])
 
 plt.xlim((-0.025, 0.025))

### 2015-07-28 manuscript/figs/fig2.py
change to light_curve() method

@@ -25,10 +25,10 @@ params.u = [0.5, 0.1, 0.1, -0.1]       	#limb darkening coefficients
 t = np.linspace(-0.025, 0.025, 1000)  	#times at which to calculate light curve	
 m = batman.TransitModel(params, t)      #initializes model
 
-flux = m.LightCurve(params)		#calculates light curve
+flux = m.light_curve(params)		#calculates light curve
 
-m.set_fac(1.0e-4)
-flux_precise = m.LightCurve(params)
+m.fac = 1.0e-4
+flux_precise = m.light_curve(params)
 
 gs = gridspec.GridSpec(2, 1, height_ratios =[3,1], hspace=0.05)
 

### 2015-07-28 manuscript/figs/fig3.py

@@ -46,7 +46,7 @@ t = np.linspace(-0.013, 0.013, 100)  	#times at which to calculate light curve
 params.limb_dark = "quadratic"          #limb darkening model
 params.u = [0.1, 0.3]       	#limb darkening coefficients
 m = batman.TransitModel(params, t)      #initializes model
-wrapped = wrapper(m.LightCurve, params)
+wrapped = wrapper(m.light_curve, params)
 time = timeit.timeit(wrapped,number=100)/100.
 #plt.axvline(time, color='0.5', linestyle='dashed')
 plt.gca().annotate("", xy=(time, .1), xycoords='data', xytext=(time, 1), textcoords='data', arrowprops=dict(arrowstyle="->")) 
@@ -56,16 +56,16 @@ print("quadratic", time)
 params.limb_dark = "nonlinear"          #limb darkening model
 params.u = [0.0, 0.7, 0.0, -0.3]       	#limb darkening coefficients
 m = batman.TransitModel(params, t)      #initializes model
-flux = m.LightCurve(params)		#calculates light curve
+flux = m.light_curve(params)		#calculates light curve
 
-#generates Figure FIXME: max err as a function of function call time
+#generates Figure 3: max err as a function of function call time
 n = 100
 ts = []
 errs = []
 fac = np.logspace(-2, 0.3, n)
 for i in range(n):
-	m.set_fac(fac[i])
-	wrapped = wrapper(m.LightCurve, params)
+	m.fac = fac[i]
+	wrapped = wrapper(m.light_curve, params)
 	if i<10: t = timeit.timeit(wrapped,number=200)/200.
 	else: t = timeit.timeit(wrapped,number=1000)/1000.
 	ts.append(t)

### 2015-07-28 batman/transitmodel.py
change fac_lo in _get_fac()

@@ -49,8 +49,10 @@ class TransitModel:
 	:param nthreads: (optional)
 		Number of threads to use for parallelization. 
 
+	:param fac: (optional)
+		Scale factor for integration step size
 	"""
-	def __init__(self, params, t, max_err=1.0, nthreads = 1):
+	def __init__(self, params, t, max_err=1.0, nthreads = 1, fac = None):
 		#checking for invalid input
 		if  (params.limb_dark == "uniform" and len(params.u) != 0) or (params.limb_dark == "linear" and len(params.u) != 1) or \
 		    (params.limb_dark == "quadratic" and len(params.u) != 2) or (params.limb_dark == "logarithmic" and len(params.u) != 2) or \
@@ -78,8 +80,9 @@ class TransitModel:
 		self.u = params.u
 		self.max_err = max_err
 		self.limb_dark = params.limb_dark
-		self.zs= _rsky._rsky(t, params.t0, params.per, params.a, params.inc*pi/180., params.ecc, params.w*pi/180.)
-		self.fac = self._get_fac()
+		self.ds= _rsky._rsky(t, params.t0, params.per, params.a, params.inc*pi/180., params.ecc, params.w*pi/180.)
+		if fac != None: self.fac = fac
+		else: self.fac = self._get_fac()
 		if nthreads==None or nthreads == 1: self.nthreads=1
 		else:
 			if nthreads <= multiprocessing.cpu_count()and nthreads >1 and openmp.detect(): self.nthreads = nthreads
@@ -98,27 +101,27 @@ class TransitModel:
 
 		"""
 		if self.limb_dark in ["logarithmic", "exponential", "nonlinear", "squareroot", "custom"]:
-			zs = np.linspace(0., 1.1, 500)
-			fac_lo = 1.0e-4
+			ds = np.linspace(0., 1.1, 500)
+			fac_lo = 5.0e-4
 			if self.limb_dark == "nonlinear":
-				f0 = _nonlinear_ld._nonlinear_ld(zs, self.rp, self.u[0], self.u[1], self.u[2], self.u[3], fac_lo, self.nthreads)
-				f = _nonlinear_ld._nonlinear_ld(zs, self.rp, self.u[0], self.u[1], self.u[2], self.u[3], self.fac, self.nthreads)
[…]

### 2015-07-28 c_src/_exponential_ld.c
change all r to x

@@ -87,7 +87,7 @@ static PyObject *_exponential_ld(PyObject *self, PyObject *args)
 	double norm = 2.*M_PI*(0.5 - 0.1666666667*c1 + 0.77750463*c2); 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
 
 	#if defined (_OPENMP)
-	#pragma omp parallel for private(d, x_in, x_out, delta, r, dx, A_i, A_f, I)
+	#pragma omp parallel for private(d, x_in, x_out, delta, x, dx, A_i, A_f, I)
 	#endif
 	for(i = 0; i < dims[0]; i++)
 	{

### 2015-07-28 c_src/_logarithmic_ld.c

@@ -86,7 +86,7 @@ static PyObject *_logarithmic_ld(PyObject *self, PyObject *args)
 	double norm = (-3.*c1 + 2.*c2 + 9.)*M_PI/9.0; 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
 
 	#if defined (_OPENMP)
-	#pragma omp parallel for private(d, x_in, x_out, delta, r, dx, A_i, A_f, I)
+	#pragma omp parallel for private(d, x_in, x_out, delta, x, dx, A_i, A_f, I)
 	#endif
 	for(i = 0; i < dims[0]; i++)
 	{

### 2015-07-28 c_src/_nonlinear_ld.c

@@ -86,7 +86,7 @@ static PyObject *_nonlinear_ld(PyObject *self, PyObject *args)
 	double norm = (-c1/10. - c2/6. - 3.*c3/14. - c4/4. + 0.5)*2.*M_PI; 	//normalization for intensity profile (faster to calculate it once, rather than every time intensity is called)		
 
 	#if defined (_OPENMP)
-	#pragma omp parallel for private(d, x_in, x_out, delta, r, dx, A_i, A_f, I)
+	#pragma omp parallel for private(d, x_in, x_out, delta, x, dx, A_i, A_f, I)
 	#endif
 	for(i = 0; i < dims[0]; i++)
 	{

### 2015-07-29 c_src/_quadratic_ld.c
fix declaration of nz

@@ -59,7 +59,7 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 
 	dims[0] = PyArray_DIMS(ds)[0]; 
 	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
-	nz = (int)dims;
+	nz = (int)dims[0];
 
 	double *f_array = PyArray_DATA(flux);
 	double *d_array = PyArray_DATA(ds);

### 2015-07-29 c_src/_quadratic_ld.c
change nz to nd

@@ -48,7 +48,7 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 	 Limb darkening has the form:
 	 I(r) = [1 - c1 * (1 - sqrt(1 - (r/rs)^2)) - c2*(1 - sqrt(1 - (r/rs)^2))^2]/(1 - c1/3 - c2/6)/pi
 */
-	int nz, nthreads;
+	int nd, nthreads;
 	double c1, c2, p, *mu,  *lambdad, *etad, \
 		*lambdae, lam, x1, x2, x3, d, omega, kap0 = 0.0, kap1 = 0.0, \
 		q, Kk, Ek, Pk, n;
@@ -59,7 +59,7 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 
 	dims[0] = PyArray_DIMS(ds)[0]; 
 	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
-	nz = (int)dims[0];
+	nd = (int)dims[0];
 
 	double *f_array = PyArray_DATA(flux);
 	double *d_array = PyArray_DATA(ds);
@@ -76,10 +76,10 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 	*/
 
 
-	lambdad = (double *)malloc(nz*sizeof(double));
-	lambdae = (double *)malloc(nz*sizeof(double));
-	etad = (double *)malloc(nz*sizeof(double));
-	mu = (double *)malloc(nz*sizeof(double));
+	lambdad = (double *)malloc(nd*sizeof(double));
+	lambdae = (double *)malloc(nd*sizeof(double));
+	etad = (double *)malloc(nd*sizeof(double));
+	mu = (double *)malloc(nd*sizeof(double));
 
 	if(fabs(p - 0.5) < 1.0e-3) p = 0.5;
 

### 2015-07-29 batman/tests.py
fix bug in multithreading test

@@ -85,7 +85,7 @@ def test():
 		m = TransitModel(params, t, nthreads = 2)
 		wrapped = wrapper(m.light_curve, params)
 		t2 = timeit.timeit(wrapped,number=100)
-		if(t1 < 0.6*t2): print("\ttest passed")
+		if(t1 > 0.9*t2): print("\ttest passed")
 		else:
 			print("\ttest failed")
 			failures +=1

### 2015-07-29 batman/transitmodel.py
raise exception if max_err < 0.001

@@ -67,6 +67,7 @@ class TransitModel:
 		if params.limb_dark not in ["uniform", "linear", "quadratic", "logarithmic", "exponential", "squareroot", "nonlinear", "custom"]: 
 			raise Exception("\""+params.limb_dark+"\""+" limb darkening not supported; allowed options are:\n \
 				uniform, linear, quadratic, logarithmic, exponential, squareroot, nonlinear, custom")
+		if max_err < 0.001: raise Exception("The lowest allowed value for max_err is 0.001. For more accurate calculation, set the integration step size explicitly with the fac parameter.")
 
 		#initializes model parameters
 		self.t = t

### 2015-07-29 manuscript/ms.tex
update intro
add refs

@@
 The transit technique has revolutionized the study of exoplanetary systems.
-Thanks largely to the \emph{Kepler} mission, over a thousand planets have been discovered with this method \citep{rowe15}.
+Thanks largely to the \emph{Kepler} mission, thousands of planets have been discovered with this method \citep{rowe15}.
 These discoveries have yielded transformative constraints on planet occurrence rates over a wide range of planet sizes, orbital periods, and host star properties \citep{borucki11, youdin11, howard12,fressin13, dong13, morton14}.
-The have also enabled the first estimates of the occurrence of habitable planets \citep{traub12, dressing13, petigura13, kopparapu13, foremanmackey14, dressing15}.
-Transit light curves can even reveal planets' atmospheric temperature structure and composition \citep[e.g.][]{seager00, lecavelier08, sing11, deming13, kreidberg14b, knutson14a, fraine14, kreidberg15}.
-A number of current and planned observational facilities -- including \emph{K2}, \emph{TESS}, \emph{CHEOPS}, \emph{JWST}, and \emph{PLATO} -- will measure precise transit light curves for thousands of exoplanets that will further transform our understanding of planet formation, evolution, and habitability.
+They have also enabled the first estimates of the occurrence of habitable planets \citep{traub12, dressing13, petigura13, kopparapu13, foremanmackey14, dressing15}.
+Transit light curves can also reveal planets' atmospheric temperature structure and composition \citep[e.g.][]{seager00, charbonneau02, lecavelier08, sing11, deming13, knutson14a, fraine14, kreidberg15}.
+A number of current and planned observational facilities -- including \emph{K2}, \emph{TESS}, \emph{CHEOPS}, \emph{JWST}, and \emph{PLATO} -- will measure precise transit light curves for thousands of exoplanets that will further advance our understanding of planet formation, evolution, and habitability.
 
@@
 In this paper, I introduce the open source Python package \texttt{batman}.
-This package is based on code that was used to model high-precision light curves obtained for atmosphere characterization \citep{kreidberg14a, kreidberg14b, stevenson14a, kreidberg15}.
+This package is based on code that was used to model high-precision light curves obtained for atmosphere characterization \citep{kreidberg14b, kreidberg14a, kreidberg15, stevenson14c, stevenson14b, stevenson14a, stevenson14d}.
 The \texttt{batman} package enables fast computation of transit light curves for any radially symmetric limb darkening law, and currently supports uniform, linear, quadratic, logarithmic, exponential, and nonlinear limb darkening.
@@
 \end{equation}
-where $x$ is the normalized radial coordinate $0 < x < 1$, $I(x)$ is the 1D stellar intensity profile, $r_p$ is the planetary radius, $d$ is the separation of centers between the star and the planet, and $A(x, r_p, d)$ is the area of intersection between two circles of radii $x$ and $r_p$, separated by a distance $d$.
-The sum is carried out over the range $x_0 = d - r_p$ to $x_n = \textrm{MIN}(d + r_p, r_s)$, where $r_s = 1$ is the stellar radius.
+where $x$ is the normalized radial coordinate $0 < x < 1$, $I(x)$ is the 1D stellar intensity profile, $r_p$ is the planetary radius (in units of stellar radii), $d$ is the separation of centers between the star and the planet (in stellar radii), and $A(x, r_p, d)$ is the area of intersection between two circles of radii $x$ and $r_p$, separated by a distance $d$.
+The sum is carried out over the range $x_0 = \textrm{MAX}(d - r_p, 0)$ to $x_n = \textrm{MIN}(d + r_p, 1)$.
 The intersecting area is given by:
@@
 x^2\cos^{-1}{u} + r_p^2\cos^{-1}{v} - 0.5\sqrt{w}, & r_p -d < x < r_p + d\\
-x^2, & x \le r_p - d\\
-r_p^2 & x \ge r_p +d \\
+\pi x^2, & x \le r_p - d\\
+\pi r_p^2 & x \ge r_p +d \\
 \end{cases}
@@
 To ensure that the truncation error is below the specified threshold, the integration step size is tuned during model initialization.
-Truncation error is measured relative to a model calculated with a very small step size ($f = 10^{-4}$).
-For typical limb darkening profiles, this method is reliable for truncation errors down to $\sim 10^{-4}$ ppm.
+Truncation error is measured relative to a model calculated with a very small step size ($f = 5\times10^{-4}$).
+For typical limb darkening profiles, this method is reliable for truncation errors down to $\sim 10^{-3}$ ppm.
 However, tuning the step size is a slow operation because it requires computing several light curve models ($\sim10$).
@@
 \resizebox{\hsize}{!}{\includegraphics{f3.pdf}}
-\caption{Execution time as a function of truncation error tolerance for a light curve modeled with the nonlinear limb darkening law (black line).
+\caption{Truncation error as a function of execution time for a light curve modeled with the nonlinear limb darkening law (black line).
 The execution time for a quadratic model (computed analytically to better than 0.05 ppm accuracy) is indicated by the arrow.
@@
 I tested \texttt{batman}'s performance over a range of typical use cases with a 1.7 GHz Intel Core i5 processor.
-In Figure\,\ref{fig:performance}, I show the execution time versus error tolerance for a single transit light curve calculation using a nonlinear intensity profile, compared to the execution time for a quadratic model computed analytically.
+In Figure\,\ref{fig:performance}, I show the truncation error versus execution time for a single transit light curve calculation using a nonlinear intensity profile, compared to the execution time for a quadratic model computed analytically.
 The test case consists of 100 points evenly sampled in time during the planet's transit.
@@
 Even for cases where F1 is correct, the computation is over an order of magnitude slower than numeric integration (for an error tolerance of 0.1 ppm).
-I concluded that integration is a faster and easier solution than analytic models for nonlinear limb darkening.
+I concluded that integration is a faster and easier solution than analytic models for the four-parameter nonlinear limb darkening law.
 

### 2015-07-29 batman/transitmodel.py
better docstrings

@@ -37,21 +37,27 @@ class TransitModel:
 	"""
 	Class for generating model transit light curves.	
 
-	:param params:
-		A :attr:`TransitParams` object containing the physical parameters of the transit
+	:param params: A :attr:`TransitParams` object containing the physical parameters of the transit
+	:type params: a `TransitParams` instance
 
-	:param t:
-		Array of times at which to calculate the model.
+	:param t: Array of times at which to calculate the model.
+	:type t: ndarray 
 
-	:param max_err: (optional)
-		Error tolerance (in parts per million) for the model.
+	:param max_err: Error tolerance (in parts per million) for the model.
+	:type max_err: float, optional
 
-	:param nthreads: (optional)
-		Number of threads to use for parallelization. 
+	:param nthreads: Number of threads to use for parallelization. 
+	:type nthreads: int, optional
 
-	:param fac: (optional)
-		Scale factor for integration step size
+	:param fac: Scale factor for integration step size
+	:type fac: float, optional
+
+
+	:Example:
+	
+	>>> m = batman.TransitModel(params, max_err = 0.5, nthreads=4)
 	"""
+
 	def __init__(self, params, t, max_err=1.0, nthreads = 1, fac = None):
 		#checking for invalid input
 		if  (params.limb_dark == "uniform" and len(params.u) != 0) or (params.limb_dark == "linear" and len(params.u) != 1) or \
@@ -95,10 +101,13 @@ class TransitModel:
 	def calc_err(self, plot = False):
 		"""
 
-		Calculate maximum error (in parts per million) for transit light curve calculation.
+		Calculate maximum error for transit light curve calculation.
 			
-		:param plot: 
-			If set to ``True``, plots the error in the light curve model as a function of separation of centers.
+		:param plot: If ``True``, plots the error in the light curve model as a function of separation of centers.
+		:type plot: bool
+
+		:return: Truncation error (parts per million)
+		:rtype: float
 
 		"""
 		if self.limb_dark in ["logarithmic", "exponential", "nonlinear", "squareroot", "custom"]:
@@ -161,11 +170,17 @@ class TransitModel:
 	
 	def light_curve(self, params):
 		"""
-		Return a model light curve.
+		Calculate a model light curve.
 
-		:param params:
-			Transit parameter object.
[…]

### 2015-07-29 setup.py
update development status

@@ -73,7 +73,7 @@ _custom_ld   = Extension('batman._custom_ld', ['c_src/_custom_ld.c', 'c_src/_cus
 _rsky = Extension('batman._rsky', ['c_src/_rsky.c'])
 
 setup(	name='batman-package', 
-	version="0.9.1", 
+	version="1.0.0", 
 	author='Laura Kreidberg',
 	author_email = 'laura.kreidberg@uchicago.edu',
 	url = 'https://github.com/lkreidberg/batman',
@@ -81,7 +81,7 @@ setup(	name='batman-package',
 	license = ['GNU GPLv3'],
 	description ='Fast transit light curve modeling',
 	classifiers = [
-		'Development Status :: 4 - Beta',
+		'Development Status :: 5 - Production/Stable',
 		'Intended Audience :: Science/Research',
 		'Topic :: Scientific/Engineering',
 		'Programming Language :: Python'

### 2015-07-29 manuscript/ms.tex
change figs to eps format

@@
 The code uses C extension modules to speed up model calculation and is parallelized with OpenMP.
-For a typical light curve with 100 data points in transit, \texttt{batman} can calculate one million quadratic limb-darkened models in three minutes with a single 1.7 GHz Intel Core i5 processor.
+For a typical light curve with 100 data points in transit, \texttt{batman} can calculate one million quadratic limb-darkened models in 30 seconds with a single 1.7 GHz Intel Core i5 processor.
 The same calculation takes seven minutes using the four-parameter nonlinear limb darkening model (computed to 1 ppm accuracy).
@@
 \begin{figure}
-\resizebox{\hsize}{!}{\includegraphics{f1.pdf}}
+\resizebox{\hsize}{!}{\includegraphics{f1.eps}}
 \caption{Schematic illustration of the integration scheme.
@@
 \begin{figure}
-\resizebox{\hsize}{!}{\includegraphics{f2.pdf}}
+\resizebox{\hsize}{!}{\includegraphics{f2.eps}}
 \caption{An example transit light curve for a nonlinear stellar intensity profile (top panel) and truncation error for the calculation (bottom panel).
@@
 \begin{figure}
-\resizebox{\hsize}{!}{\includegraphics{f3.pdf}}
+\resizebox{\hsize}{!}{\includegraphics{f3.eps}}
 \caption{Truncation error as a function of execution time for a light curve modeled with the nonlinear limb darkening law (black line).
@@
 The stellar intensity profile is the same for the nonlinear and quadratic models: the nonlinear limb darkening coefficients are $(0.0, 0.7, 0.0, -0.3)$ and the quadratic coefficients are $(0.1, 0.3)$.
-For this case, reaching a truncation error of 1.0 ppm via integration is just two times slower than the analytic calculation.
 Decreasing the truncation error by a factor of 10 increases the computation time by a factor of three.
@@
 
+\input{ms.bbl}
 \bibliographystyle{apj}

### 2015-07-29 manuscript/figs/fig1.py

@@ -103,5 +103,5 @@ plt.gca().text(0.3,0.04, "d", fontsize=16, weight='bold')
 plt.xlabel("x (stellar radii)")
 plt.ylabel("y (stellar radii)")
 
-plt.savefig("f1.pdf", dpi=300)
+plt.savefig("f1.eps", dpi=300)
 plt.show()

### 2015-07-29 manuscript/figs/fig2.py

@@ -46,6 +46,6 @@ plt.ylabel("Error (ppm)")
 plt.xlim((-0.5, 0.5))
 plt.ylim((-1.1,0.1))
 
-plt.savefig("f2.pdf", dpi=300)
+plt.savefig("f2.eps", dpi=300)
 plt.show()
 

### 2015-07-29 manuscript/figs/fig3.py

@@ -46,6 +46,7 @@ t = np.linspace(-0.013, 0.013, 100)  	#times at which to calculate light curve
 params.limb_dark = "quadratic"          #limb darkening model
 params.u = [0.1, 0.3]       	#limb darkening coefficients
 m = batman.TransitModel(params, t)      #initializes model
+f1 = m.light_curve(params)
 wrapped = wrapper(m.light_curve, params)
 time = timeit.timeit(wrapped,number=100)/100.
 #plt.axvline(time, color='0.5', linestyle='dashed')
@@ -87,7 +88,7 @@ plt.xscale('log')
 plt.xlabel("Execution time (s)")
 plt.ylabel("Truncation error (ppm)")
 plt.tight_layout()
-plt.savefig("f3.pdf", dpi=300)
+plt.savefig("f3.eps", dpi=300)
 plt.show()
 
 

### 2015-08-04 c_src/_eclipse.c
add secondary eclipse module

@@ -1,0 +1,112 @@
+/* The batman package: fast computation of exoplanet transit light curves
+ * Copyright (C) 2015 Laura Kreidberg	 
+ * 
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ * 
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+#define NPY_NO_DEPRECATED_API NPY_1_7_API_VERSION
+#include <Python.h>
+#include "numpy/arrayobject.h"
+#include <math.h>
+
+#if defined (_OPENMP)
+#  include <omp.h>
+#endif
+
+static PyObject *_eclipse(PyObject *self, PyObject *args)
+{
+	int nthreads;
+	double d, p, kap0, kap1, fp, alpha_t, alpha_o;
+
+	PyArrayObject *ds, *flux;
+	npy_intp i, dims[1];
+	
+  	if(!PyArg_ParseTuple(args, "Oddi", &ds, &p, &fp, &nthreads)) return NULL;		//parses function input
+
+	dims[0] = PyArray_DIMS(ds)[0]; 
+	flux = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ds));	//creates numpy array to store return flux values
+	
+	double *f_array = PyArray_DATA(flux);
+	double *d_array = PyArray_DATA(ds);
+
+	if(fabs(p - 0.5) < 1.e-3) p = 0.5;
+
+	#if defined (_OPENMP)
+	omp_set_num_threads(nthreads);
+	#endif
+
+	#if defined (_OPENMP)
+	#pragma omp parallel for private(d, kap1, kap0)
+	#endif
+	for(i=0; i<dims[0]; i++)
+	{
+		d = d_array[i]; 						// separation of centers
+		
[…]

### 2015-08-04 setup.py

@@ -71,6 +71,7 @@ _logarithmic_ld   = Extension('batman._logarithmic_ld', ['c_src/_logarithmic_ld.
 _exponential_ld   = Extension('batman._exponential_ld', ['c_src/_exponential_ld.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
 _custom_ld   = Extension('batman._custom_ld', ['c_src/_custom_ld.c', 'c_src/_custom_intensity.c'], extra_compile_args = parallel_args, libraries = parallel_libraries) 
 _rsky = Extension('batman._rsky', ['c_src/_rsky.c'])
+_eclipse = Extension('batman._eclipse', ['c_src/_eclipse.c'])
 
 setup(	name='batman-package', 
 	version="1.0.0", 
@@ -88,5 +89,5 @@ setup(	name='batman-package',
 		],
 	include_dirs = [np.get_include()],
 	install_requires = ['numpy'],
-	ext_modules=[_nonlinear_ld, _quadratic_ld, _uniform_ld, _logarithmic_ld, _exponential_ld, _custom_ld, _rsky]
+	ext_modules=[_nonlinear_ld, _quadratic_ld, _uniform_ld, _logarithmic_ld, _exponential_ld, _custom_ld, _rsky, _eclipse]
 )

### 2015-08-04 batman/tests.py
add joker

@@ -93,8 +93,20 @@ def test():
 
 	if failures == 0: 
 		print("\nCongratulations! all tests passed")
-		print("  ____       ____\n  )   \     /   (\n   )_  \_V_/  _(\n     )__   __(\n        `-'")
-	else: print("Uh oh; tests finished with " + "{0}".format(failures) + " failures")
-
+		print("""
+  ____       ____
+  )   \     /   (
+   )_  \_V_/  _(
+     )__   __(
+        `-'""")
+	else: 
+		print("Uh oh; tests finished with " + "{0}".format(failures) + " failures")
 	
-
+		print("""
+         *
+    __  /|
+   /_ \/ |__
+  *" \.--._ \           
+     ( ^^ )\/
+      \__/ *""")
+	         

### 2015-08-04 c_src/_rsky.c
redefine t0 as time of inferior conjunction

@@ -39,11 +39,12 @@ static PyObject *_rsky(PyObject *self, PyObject *args)
 		(see the section by Murray, and Winn eq. 5).  In the Mandel & Agol 
 		(2002) paper, this quantity is denoted d.
 	*/
-	double ecc, E, inc, a, d, f, omega, per, M, n, t0, eps, t, BIGD = 100.;
+	double ecc, E, inc, a, d, f, omega, per, M, n, tp, tc, eps, t, BIGD = 100.;
+	int transittype;
 	npy_intp i, dims[1];
 	PyArrayObject *ts, *ds;
 
-  	if(!PyArg_ParseTuple(args,"Odddddd", &ts, &t0, &per, &a, &inc, &ecc, &omega)) return NULL; 
+  	if(!PyArg_ParseTuple(args,"Oddddddi", &ts, &tc, &per, &a, &inc, &ecc, &omega, &transittype)) return NULL; 
 
 	dims[0] = PyArray_DIMS(ts)[0]; 
 	ds = (PyArrayObject *) PyArray_SimpleNew(1, dims, PyArray_TYPE(ts));
@@ -57,18 +58,25 @@ static PyObject *_rsky(PyObject *self, PyObject *args)
 	for(i = 0; i < dims[0]; i++)
 	{
 		t = t_array[i];
+		
+		//calculates time of periastron passage from time of inferior conjunction 
+		f = M_PI/2. - omega;								//true anomaly corresponding to time of primary transit center
+		E = 2.*atan(sqrt((1. - ecc)/(1. + ecc))*tan(f/2.));				//corresponding eccentric anomaly
+		M = E - ecc*sin(E);						
+		tp = tc - per*M/2./M_PI;							//time of periastron 
 
 		if(ecc < 1.0e-5)
 		{
-			f = ((t - t0)/per - (int)((t - t0)/per))*2.*M_PI;			//calculates f for a circular orbit
+			f = ((t - tp)/per - (int)((t - tp)/per))*2.*M_PI;			//calculates f for a circular orbit
 		}
 		else
 		{
-			M = n*(t - t0);
+			M = n*(t - tp);
 			E = getE(M, ecc);
 			f = 2.*atan(sqrt((1.+ecc)/(1.-ecc))*tan(E/2.));
 		}
-		if ((omega+fabs(f)) >= M_PI && (omega+fabs(f)) <= 2.*M_PI) d = BIGD; 						// removes secondary transits
[…]

### 2015-08-04 manuscript/ms.tex
add description of eclipse model

@@
 \newcommand*{\blue}{\textcolor{blue}}
+\sloppy
 
@@
 \begin{abstract}
-I introduce \texttt{batman}, a Python package for modeling exoplanet transit light curves.
+I introduce \texttt{batman}, a Python package for modeling exoplanet transit \blue{and eclipse} light curves.
 The \texttt{batman} package supports calculation of light curves for any radially symmetric stellar limb darkening law, using a new integration algorithm for models that cannot be quickly calculated analytically.
@@
 Models for the remaining cases are computed with an efficient new integration scheme, described in \S\,\ref{section:algorithm}.
-I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:summary}.
+\blue{The package also supports secondary eclipse modeling.} I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:summary}.
 
@@
 
+\subsection{Secondary Eclipse Model}
+\blue{\texttt{batman} can also model secondary eclipses. Eclipse light curves are generated with 
+$$
+f = 1 - \frac{\alpha f_p}{1 + f_p}
+$$
+where $f$ is normalized flux, $f_p$ is the planet-to-star flux ratio, and $\alpha$ is the fraction of the planet disk that is occulted by the star.
+The model is normalized such that the sum of the stellar flux and the planet flux is unity.
+For a separation $d$, the occultation fraction $\alpha(d) = \alpha_t(d)/r_p^2$, where $1 - \alpha_t(d)$ is the transit light curve with uniform limb darkening.
+Note that this model assumes the planet flux is constant for all orbital phases.}
+
 \subsection{Utilities}
 \texttt{batman} includes a utility function to calculate the separation of centers $d$ between the star and the planet based on orbital parameters of the system.
-The input parameters are the planet semi-major axis $a$, inclination $i$, eccentricity $e$, longitude of periastron $\omega$, period $P$, and transit ephemeris $t_0$.
-For eccentric orbits, the transit ephemeris is the time of periastron; for circular orbits, $t_0$ is the mid-transit time.
+The input parameters are the planet semi-major axis $a$, inclination $i$, eccentricity $e$, longitude of periastron $\omega$, period $P$, and \blue{time of inferior conjunction $t_0$}.
 The separation of centers is given by:
@@
 $$
-where $f$ is defined as the orbital phase (for circular orbits) or the true anomaly (for eccentric orbits).
+where $f$ is the true anomaly.
 The true anomaly is calculated with the algorithm provided by Murray \& Correia in Chapter 1 of \cite{seager10}.
+\blue{For circular orbits, we adopt the convention $f = \pi/2$.}
 
-For the case $\pi <= \omega + |f| <= 2\pi$, $d$ is set to a large value so that the code does not produce a transit model when the planet is being eclipsed by the star.
-To model a secondary eclipse, use a uniform limb darkening profile and set $t_0$ equal to the mid-eclipse time.
+\blue{\texttt{batman} also provides utilities to calculate the time of periastron and the secondary eclipse time from the other transit parameters, using the method described in \S 3.1 of \cite{eastman13}.}
 
@@
 I introduced a new algorithm for computing transit light curves for any radially symmetric stellar limb darkening law.
-I also described the open-source Python package \texttt{batman}, a versatile code to generate model light curves.
+I also described the open-source Python package \texttt{batman}, a versatile code to generate model \blue{transit and eclipse} light curves.
 Uniform, linear, quadratic, logarithmic, exponential, and four-parameter nonlinear limb darkening laws are currently supported.
@@
 I also thank contributors to SciPy, Matplotlib, and the Python Programming Language for software and services.
+I am grateful for helpful suggestions from the referee, Jason Eastman, that improved the code and the manuscript.
 Support for this work was provided by a grant from the National Science Foundation through a Graduate Research Fellowship to the author.
 
-\input{ms.bbl}
+%\input{ms.bbl}
 \bibliographystyle{apj}

### 2015-08-04 c_src/_eclipse.c
better comments

@@ -65,8 +65,9 @@ static PyObject *_eclipse(PyObject *self, PyObject *args)
 		{
 			kap1=acos(fmin((1. - p*p + d*d)/2./d, 1.));
 			kap0=acos(fmin((p*p + d*d - 1.)/2./p/d, 1.));
-			alpha_t = (p*p*kap0 + kap1 - 0.5*sqrt(fmax(4.*d*d - pow(1. + d*d - p*p, 2.), 0.)))/M_PI;
-			alpha_o = alpha_t/p/p;
+			alpha_t = (p*p*kap0 + kap1 - 0.5*sqrt(fmax(4.*d*d \
+				- pow(1. + d*d - p*p, 2.), 0.)))/M_PI;		//transit depth
+			alpha_o = alpha_t/p/p;				 	//fraction of planet disk that is eclipsed by the star
 			f_array[i] = 1. - alpha_o*fp/(1. + fp);
 			
 		}

### 2015-08-04 c_src/_rsky.c

@@ -75,8 +75,8 @@ static PyObject *_rsky(PyObject *self, PyObject *args)
 			E = getE(M, ecc);
 			f = 2.*atan(sqrt((1.+ecc)/(1.-ecc))*tan(E/2.));
 		}
-		if (transittype == 1 && sin(f + omega)*sin(inc) <= 0.) d = BIGD;						//masks secondary
-		else if (transittype == 2 && sin(f + omega)*sin(inc) >= 0.) d = BIGD;						//masks primary
+		if (transittype == 1 && sin(f + omega)*sin(inc) <= 0.) d = BIGD;						//z < 0, so d is set to large value in order to not model primary transit during secondary eclipse
+		else if (transittype == 2 && sin(f + omega)*sin(inc) >= 0.) d = BIGD;						//z > 0, so d is set to large value in order not to model secondary eclipse during primary transit
 		else d = a*(1.0 - ecc*ecc)/(1.0 + ecc*cos(f))*sqrt(1.0 - sin(omega + f)*sin(omega + f)*sin(inc)*sin(inc));	//calculates separation of centers 
 		d_array[i] = d;
 	}

### 2015-08-04 batman/transitmodel.py
add get_t_periastron() and get_t_secondary() methods

@@ -23,6 +23,7 @@ from . import _logarithmic_ld
 from . import _exponential_ld
 from . import _custom_ld
 from . import _rsky
+from . import _eclipse
 #from . import _mandelagol_nonlinear_ld
 from math import pi
 import multiprocessing
@@ -52,13 +53,15 @@ class TransitModel:
 	:param fac: Scale factor for integration step size
 	:type fac: float, optional
 
+	:param transittype: Type of transit ("primary" or "secondary")
+	:type transittype: string, optional
 
 	:Example:
 	
 	>>> m = batman.TransitModel(params, max_err = 0.5, nthreads=4)
 	"""
 
-	def __init__(self, params, t, max_err=1.0, nthreads = 1, fac = None):
+	def __init__(self, params, t, max_err=1.0, nthreads = 1, fac = None, transittype = "primary"):
 		#checking for invalid input
 		if  (params.limb_dark == "uniform" and len(params.u) != 0) or (params.limb_dark == "linear" and len(params.u) != 1) or \
 		    (params.limb_dark == "quadratic" and len(params.u) != 2) or (params.limb_dark == "logarithmic" and len(params.u) != 2) or \
@@ -74,6 +77,7 @@ class TransitModel:
 			raise Exception("\""+params.limb_dark+"\""+" limb darkening not supported; allowed options are:\n \
 				uniform, linear, quadratic, logarithmic, exponential, squareroot, nonlinear, custom")
 		if max_err < 0.001: raise Exception("The lowest allowed value for max_err is 0.001. For more accurate calculation, set the integration step size explicitly with the fac parameter.")
+		if transittype not in ["primary", "secondary"]: raise Exception("Allowed transit types are \"primary\" and \"secondary\".")
 
 		#initializes model parameters
 		self.t = t
@@ -87,7 +91,10 @@ class TransitModel:
 		self.u = params.u
 		self.max_err = max_err
 		self.limb_dark = params.limb_dark
-		self.ds= _rsky._rsky(t, params.t0, params.per, params.a, params.inc*pi/180., params.ecc, params.w*pi/180.)
+		self.fp = params.fp
+		if transittype == "primary": self.transittype = 1
+		else: self.transittype = 2
[…]

### 2015-08-04 c_src/_quadratic_ld.c
bug fix for rp > 1; d - (p-1) < eps

@@ -177,7 +177,9 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 			continue;
 		}
 		 //occulting star partly occults the source and crosses the limb:
-		if((d > 0.5 + fabs(p  - 0.5) && d < 1.0 + p) || (p > 0.5 && d > fabs(1.0 - p)*1.0001 \
+		//if((d > 0.5 + fabs(p  - 0.5) && d < 1.0 + p) || (p > 0.5 && d > fabs(1.0 - p)*1.0001 \
+		//&& d < p))  //the factor of 1.0001 is from the Mandel/Agol Fortran routine, but gave bad output for d near fabs(1-p)
+		if((d > 0.5 + fabs(p  - 0.5) && d < 1.0 + p) || (p > 0.5 && d > fabs(1.0 - p) \
 			&& d < p))
 		{
 			//printf("zone 3.1\n");
@@ -226,7 +228,6 @@ static PyObject *_quadratic_ld(PyObject *self, PyObject *args)
 
 	return PyArray_Return((PyArrayObject *)flux);
 }
-
 	
 /*
 

### 2015-08-04 batman/transitmodel.py
move get_t_periastron and get_t_secondary to TransitModel

@@ -218,6 +218,26 @@ class TransitModel:
 			else: raise Exception("Invalid limb darkening option")
 		else: return _eclipse._eclipse(self.ds, params.rp, params.fp, self.nthreads)			
 
+	def get_t_periastron(self, params):
+		"""
+		Return the time of periastron passage.
+		"""
+		TA = pi/2. - params.w*pi/180.
+		E = 2.*np.arctan(np.sqrt((1. - params.ecc)/(1. + params.ecc))*np.tan(TA/2.))
+		M = E - params.ecc*np.sin(E)
+		t_periastron = params.t0 - params.per*M/2./pi
+		return t_periastron
+
+	def get_t_secondary(self, params):
+		"""
+		Return the time of secondary eclipse center.
+		"""
+		TA = 3.*pi/2. - params.w*pi/180.
+		E = 2.*np.arctan(np.sqrt((1. - params.ecc)/(1. + params.ecc))*np.tan(TA/2.))
+		M = E - params.ecc*np.sin(E)
+		t_secondary = self.get_t_periastron(params) + params.per*M/2./pi
+		return t_secondary
+
 
 class TransitParams(object):
 	"""
@@ -277,22 +297,3 @@ class TransitParams(object):
 		self.limb_dark = None
 		self.fp = None
 
-	def get_t_periastron(self):
-		"""
-		Return the time of periastron passage.
-		"""
-		TA = pi/2. - self.w*pi/180.
-		E = 2.*np.arctan(np.sqrt((1. - self.ecc)/(1. + self.ecc))*np.tan(TA/2.))
-		M = E - self.ecc*np.sin(E)
-		t_periastron = self.t0 - self.per*M/2./pi
-		return t_periastron
-
-	def get_t_secondary(self):
-		"""
-		Return the time of secondary eclipse center.
-		"""
-		TA = 3.*pi/2. - self.w*pi/180.
-		E = 2.*np.arctan(np.sqrt((1. - self.ecc)/(1. + self.ecc))*np.tan(TA/2.))
-		M = E - self.ecc*np.sin(E)
-		t_secondary = self.get_t_periastron() + self.per*M/2./pi
-		return t_secondary

### 2015-08-04 docs/demo.py

@@ -1,74 +1,0 @@
-# The batman package: fast computation of exoplanet transit light curves
-# Copyright (C) 2015 Laura Kreidberg	 
-# 
-# This program is free software: you can redistribute it and/or modify
-# it under the terms of the GNU General Public License as published by
-# the Free Software Foundation, either version 3 of the License, or
-# (at your option) any later version.
-# 
-# This program is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-# 
-# You should have received a copy of the GNU General Public License
-# along with this program.  If not, see <http://www.gnu.org/licenses/>.
-
-import batman
-import numpy as np
-import matplotlib.pyplot as plt
-from pylab import *
-from matplotlib import rc
-
-rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
-matplotlib.rcParams.update({'font.size':14})
-
-params = batman.TransitParams()
-params.t0 = 0.                               #time of periastron passage (for eccentric orbits), OR
-                                             #mid-transit time (for circular orbits)
-params.per = 1.                              #orbital period
-params.rp = 0.1                              #planet radius (in stellar radii)
-params.a = 12.                               #semi-major axis (in stellar radii)
-params.inc = 87.                             #orbital inclination
-params.ecc = 0.                              #eccentricity
-params.w = 90.                               #longitude of periastron  #FIXME check if this makes sense
-params.u = [0., 0.7, 0., -0.3]               #limb darkening coefficients
-params.limb_dark = "nonlinear"       	     #limb darkening model
-
-t = np.linspace(-0.02, 0.02, 1000)           #times at which to calculate light curve
-
-m = batman.TransitModel(params, t)
-
-flux = m.light_curve(params)
-
-params.rp = 0.11                             #updates the planet radius
-new_flux = m.light_curve(params)              #recalculates model light curve
-
-plt.plot(t*24., flux, label = "rp = 0.1")
-plt.plot(t*24., new_flux, label = "rp = 0.11")
-plt.xlabel("Time from central transit (hours)")
-plt.ylabel("Flux")
-plt.legend()
-plt.savefig("new_rp.png")
[…]

### 2015-08-04 docs/quickstart.py
update docs to include eclipse modeling

@@ -25,8 +25,7 @@ rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
 matplotlib.rcParams.update({'font.size':14})
 
 params = batman.TransitParams()
-params.t0 = 0. 				#time of periastron passage (for eccentric orbits), OR
-					#mid-transit time (for circular orbits)
+params.t0 = 0. 				#time of inferior conjunction 
 params.per = 1.				#orbital period	
 params.rp = 0.1				#planet radius (in units of stellar radii)
 params.a = 15.				#semi-major axis (in units of stellar radii)

### 2015-08-04 docs/tutorial.py

@@ -24,8 +24,7 @@ rc('font',**{'family':'sans-serif','sans-serif':['Arial']})
 matplotlib.rcParams.update({'font.size':14})
 
 params = batman.TransitParams()	        #object to store transit parameters
-params.t0 = 0. 				#time of periastron passage (for eccentric orbits), OR
-					#mid-transit time (for circular orbits)
+params.t0 = 0. 				#time of inferior conjunction 
 params.per = 1.				#orbital period	
 params.rp = 0.1				#planet radius (in units of stellar radii)
 params.a = 15.				#semi-major axis (in units of stellar radii)
@@ -35,7 +34,7 @@ params.w = 90.				#longitude of periastron (in degrees)
 params.limb_dark = "nonlinear"          #limb darkening model
 params.u = [0.5, 0.1, 0.1, -0.1]       	#limb darkening coefficients
    
-t = np.linspace(-0.025, 0.025, 1000)  	#times at which to calculate light curve	
+"""t = np.linspace(-0.025, 0.025, 1000)  	#times at which to calculate light curve	
 m = batman.TransitModel(params, t)      #initializes model
 
 flux = m.light_curve(params)		#calculates light curve
@@ -68,7 +67,7 @@ for i in range(4):
 plt.xlim((-0.025, 0.025))
 plt.ylim((0.987, 1.001))
 plt.legend()
-plt.xlabel("Time from central transit (days)")
+plt.xlabel("Time (days)")
 plt.ylabel("Relative flux")
 plt.savefig("lightcurves.png")
 #plt.show()
@@ -77,4 +76,21 @@ m = batman.TransitModel(params, t, max_err = 0.5)
 plt.clf()
 m.calc_err(plot = True) 
 
-m = batman.TransitModel(params, t, nthreads = 4)
+#m = batman.TransitModel(params, t, nthreads = 4)"""
+
+params.fp = 0.001
+t = np.linspace(-0.52, -0.48, 1000)
+m = batman.TransitModel(params, t, transittype="secondary")	       
+flux = m.light_curve(params)
+
+plt.plot(t, flux)
+plt.ylim((0.9989, 1.0001))
+plt.xlim((-0.52, -0.48))
+plt.xlabel("Time (days)") 
+plt.ylabel("Relative flux")
+plt.savefig("eclipse.png")
+#plt.show()
+
+t_secondary = m.get_t_secondary(params)
+print(t_secondary)
+

### 2015-08-04 manuscript/ms.tex
update manuscript in response to referee

@@
 
-Computationally intensive sections of code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
+Computationally intensive sections of code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by \blue{a factor of 30} over a pure Python implementation \blue{for quadratic limb darkening}.
 \texttt{batman} also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.

### 2015-08-04 docs/conf.py
add link to paper

@@ -59,9 +59,9 @@ copyright = u'2015, Laura Kreidberg'
 # built documents.
 #
 # The short X.Y version.
-version = '1.0'
+version = '2.0'
 # The full version, including alpha/beta/rc tags.
-release = '1.0'
+release = '2.0.0'
 
 # The language for content autogenerated by Sphinx. Refer to documentation
 # for a list of supported languages.

### 2015-08-04 setup.py
update version number

@@ -74,7 +74,7 @@ _rsky = Extension('batman._rsky', ['c_src/_rsky.c'])
 _eclipse = Extension('batman._eclipse', ['c_src/_eclipse.c'])
 
 setup(	name='batman-package', 
-	version="1.0.0", 
+	version="2.0.0", 
 	author='Laura Kreidberg',
 	author_email = 'laura.kreidberg@uchicago.edu',
 	url = 'https://github.com/lkreidberg/batman',

### 2015-08-04 batman/transitmodel.py
update docstrings

@@ -270,12 +270,14 @@ class TransitParams(object):
 	:param limb_dark: Limb darkening model (choice of "nonlinear", "quadratic", "exponential", "logarithmic", "squareroot", "linear", "uniform", or "custom")
 	:type limb_dark: str
 
+	:param fp: Planet-to-star flux ratio (for secondary eclipse models).
+	:type fp: float, optional
+
 	:Example:
 	
 	>>> import batman
 	>>> params = batman.TransitParams()
-	>>> params.t0 = 0. 				#time of periastron passage (for eccentric orbits), OR
-	>>>						#mid-transit time (for circular orbits)
+	>>> params.t0 = 0. 				#time of inferior conjunction
 	>>> params.per = 1.				#orbital period	
 	>>> params.rp = 0.1				#planet radius (in units of stellar radii)
 	>>> params.a = 15.				#semi-major axis (in units of stellar radii)

### 2015-08-07 batman/transitmodel.py
add get_t_conjunction() method

@@ -93,7 +93,9 @@ class TransitModel:
 		self.limb_dark = params.limb_dark
 		self.fp = params.fp
 		if transittype == "primary": self.transittype = 1
-		else: self.transittype = 2
+		else: 
+			self.transittype = 2
+			params.t0 = self.get_t_conjunction(params)
 		self.ds= _rsky._rsky(t, params.t0, params.per, params.a, params.inc*pi/180., params.ecc, params.w*pi/180., self.transittype)
 		if fac != None: self.fac = fac
 		else: self.fac = self._get_fac()
@@ -218,35 +220,51 @@ class TransitModel:
 			else: raise Exception("Invalid limb darkening option")
 		else: return _eclipse._eclipse(self.ds, params.rp, params.fp, self.nthreads)			
 
-	def get_t_periastron(self, params):
-		"""
-		Return the time of periastron passage.
-		"""
-		TA = pi/2. - params.w*pi/180.
+	def _get_phase(self, params, position):
+		if position == "periastron": TA = 0.
+		elif position == "primary": TA = pi/2. - params.w*pi/180.
+		elif position == "secondary": TA = 3.*pi/2. - params.w*pi/180.
+		
 		E = 2.*np.arctan(np.sqrt((1. - params.ecc)/(1. + params.ecc))*np.tan(TA/2.))
 		M = E - params.ecc*np.sin(E)
-		t_periastron = params.t0 - params.per*M/2./pi
-		return t_periastron
+		return M/2./pi
+	
+	def get_t_periastron(self, params):
+		"""
+		Return the time of periastron passage (calculated based on the time of inferior conjunction, params.t0).
+		"""
+		phase = self._get_phase(params, "primary")
+		return params.t0 - params.per*phase
 
 	def get_t_secondary(self, params):
 		"""
-		Return the time of secondary eclipse center.
+		Return the time of secondary eclipse center (calculated based on the time of inferior conjunction, params.t0).
 		"""
-		TA = 3.*pi/2. - params.w*pi/180.
-		E = 2.*np.arctan(np.sqrt((1. - params.ecc)/(1. + params.ecc))*np.tan(TA/2.))
-		M = E - params.ecc*np.sin(E)
-		t_secondary = self.get_t_periastron(params) + params.per*M/2./pi
-		return t_secondary
+		phase = self._get_phase(params, "primary")
+		phase2 = self._get_phase(params, "secondary")
+		return params.t0 + params.per*(phase2-phase)
+
+	def get_t_conjunction(self, params):
+		"""
[…]

### 2015-08-07 c_src/_eclipse.c
normalize eclipse so f_star = 1

@@ -55,11 +55,11 @@ static PyObject *_eclipse(PyObject *self, PyObject *args)
 		
 		if(d >= 1. + p) 
 		{
-			f_array[i] = 1.;					//no overlap
+			f_array[i] = 1. + fp;					//planet fully visible
 		}
 		else if(d < 1. - p)
 		{
-			f_array[i] = 1./(1.+fp);
+			f_array[i] = 1.;					//planet fully occulted
 		}
 		else								//planet is crossing the limb
 		{
@@ -68,11 +68,9 @@ static PyObject *_eclipse(PyObject *self, PyObject *args)
 			alpha_t = (p*p*kap0 + kap1 - 0.5*sqrt(fmax(4.*d*d \
 				- pow(1. + d*d - p*p, 2.), 0.)))/M_PI;		//transit depth
 			alpha_o = alpha_t/p/p;				 	//fraction of planet disk that is eclipsed by the star
-			f_array[i] = 1. - alpha_o*fp/(1. + fp);
-			
+			f_array[i] = 1. + fp*(1. - alpha_o);			//planet partially occulted
 		}
 	}
-
 	return PyArray_Return((PyArrayObject *)flux);
 }
 

### 2015-08-07 manuscript/ms.tex
updated in response to ref
describe change to eclipse normalization

--- 2015-08-04 manuscript/ms.tex
@@
 \begin{abstract}
-I introduce \texttt{batman}, a Python package for modeling exoplanet transit \blue{and eclipse} light curves.
+I introduce \texttt{batman}, a Python package for modeling exoplanet transit and eclipse light curves.
 The \texttt{batman} package supports calculation of light curves for any radially symmetric stellar limb darkening law, using a new integration algorithm for models that cannot be quickly calculated analytically.
@@
 Models for the remaining cases are computed with an efficient new integration scheme, described in \S\,\ref{section:algorithm}.
-\blue{The package also supports secondary eclipse modeling.} I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:summary}.
+The package also supports secondary eclipse modeling.
+I discuss \texttt{batman}'s features and performance in \S\,\ref{section:batman} and conclude in \S\,\ref{section:summary}.
 
@@
 \subsection{Secondary Eclipse Model}
-\blue{\texttt{batman} can also model secondary eclipses. Eclipse light curves are generated with 
+\texttt{batman} can also model secondary eclipses. Eclipse light curves are generated with 
 $$
-f = 1 - \frac{\alpha f_p}{1 + f_p}
+f = 1 + f_p(1-\alpha)
 $$
 where $f$ is normalized flux, $f_p$ is the planet-to-star flux ratio, and $\alpha$ is the fraction of the planet disk that is occulted by the star.
-The model is normalized such that the sum of the stellar flux and the planet flux is unity.
+The model is normalized such that the stellar flux is unity.
 For a separation $d$, the occultation fraction $\alpha(d) = \alpha_t(d)/r_p^2$, where $1 - \alpha_t(d)$ is the transit light curve with uniform limb darkening.
-Note that this model assumes the planet flux is constant for all orbital phases.}
+Note that this model assumes the planet flux is constant for all orbital phases.
 
@@
 \texttt{batman} includes a utility function to calculate the separation of centers $d$ between the star and the planet based on orbital parameters of the system.
-The input parameters are the planet semi-major axis $a$, inclination $i$, eccentricity $e$, longitude of periastron $\omega$, period $P$, and \blue{time of inferior conjunction $t_0$}.
+The input parameters are the planet semi-major axis $a$, inclination $i$, eccentricity $e$, longitude of periastron $\omega$, period $P$, and time of inferior conjunction $t_0$.
 The separation of centers is given by:
@@
 The true anomaly is calculated with the algorithm provided by Murray \& Correia in Chapter 1 of \cite{seager10}.
-\blue{For circular orbits, we adopt the convention $f = \pi/2$.}
+For circular orbits, we adopt the convention $f = \pi/2$.
 
-\blue{\texttt{batman} also provides utilities to calculate the time of periastron and the secondary eclipse time from the other transit parameters, using the method described in \S 3.1 of \cite{eastman13}.}
+\texttt{batman} also provides utilities to calculate the time of periastron and the secondary eclipse time from the other transit parameters, using the method described in \S 3.1 of \cite{eastman13}.
 
@@
 
-Computationally intensive sections of code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by roughly two orders of magnitude over a pure Python implementation.
+Computationally intensive sections of code (including all of the transit model calculation) are written as C extension modules with the Python/C API, which improves the performance by a factor of 30 over a pure Python implementation for quadratic limb darkening.
 \texttt{batman} also includes the option to parallelize at the C level with OpenMP, which further speeds up the calculation.
@@
 I introduced a new algorithm for computing transit light curves for any radially symmetric stellar limb darkening law.
-I also described the open-source Python package \texttt{batman}, a versatile code to generate model \blue{transit and eclipse} light curves.
+I also described the open-source Python package \texttt{batman}, a versatile code to generate model transit and eclipse light curves.
 Uniform, linear, quadratic, logarithmic, exponential, and four-parameter nonlinear limb darkening laws are currently supported.

### 2015-08-07 setup.py
update version number

@@ -74,7 +74,7 @@ _rsky = Extension('batman._rsky', ['c_src/_rsky.c'])
 _eclipse = Extension('batman._eclipse', ['c_src/_eclipse.c'])
 
 setup(	name='batman-package', 
-	version="2.0.0", 
+	version="2.1.0", 
 	author='Laura Kreidberg',
 	author_email = 'laura.kreidberg@uchicago.edu',
 	url = 'https://github.com/lkreidberg/batman',

### 2015-08-07 c_src/_custom_intensity.c
change variable r to x

@@ -18,17 +18,17 @@
 #include <math.h>
 
 /*
-	- The intensity function returns the stellar intensity at a radius r; where 0 <= r <= 1
-	- The function arguments are the normalized radius (r), and limb darkening coefficients c1, ..., un
-	- see below for an example intensity profile for I(r) \propto 1 - c1*(1-sqrt(1-r^2)) - c2*ln((sqrt(1-r^2)+c)/(1+c))
+	- The intensity function returns the stellar intensity at a radius x; where 0 <= x <= 1
+	- The function arguments are the normalized radius (x), and limb darkening coefficients c1, ..., un
+	- see below for an example intensity profile for I(x) \propto 1 - c1*(1-sqrt(1-x^2)) - c2*ln((sqrt(1-x^2)+c)/(1+c))
 	- The normalization constant is calculated by constraining the integrated intensity to equal 1:
-		\int_r \int_theta {I(r)*r*dr*dtheta}/norm = 1
+		\int_x \int_theta {I(x)*x*dx*dtheta}/norm = 1
 */
 
-double intensity(double r, double c1, double c2, double c3, double c4, double c5, double c6) 
+double intensity(double x, double c1, double c2, double c3, double c4, double c5, double c6) 
 {
-	if(r > 0.99995) r = 0.99995;
-	double mu = sqrt(1. - r*r);
+	if(x > 0.99995) x = 0.99995;
+	double mu = sqrt(1. - x*x);
 	double norm = 2.*M_PI*(-c1/6. - c2*c3/2. + c2/4. + 0.5 + c2*c3*c3*log(1. + 1./c3)/2.); 
 	return (1. - c1*(1. - mu) - c2*log((mu + c3)/(1. + c3)))/norm;
 }

### 2015-08-07 batman/transitmodel.py
typo fix in docstring

@@ -231,14 +231,14 @@ class TransitModel:
 	
 	def get_t_periastron(self, params):
 		"""
-		Return the time of periastron passage (calculated based on the time of inferior conjunction, params.t0).
+		Return the time of periastron passage (calculated using `params.t0`).
 		"""
 		phase = self._get_phase(params, "primary")
 		return params.t0 - params.per*phase
 
 	def get_t_secondary(self, params):
 		"""
-		Return the time of secondary eclipse center (calculated based on the time of inferior conjunction, params.t0).
+		Return the time of secondary eclipse center (calculated using `params.t0`).
 		"""
 		phase = self._get_phase(params, "primary")
 		phase2 = self._get_phase(params, "secondary")
@@ -246,7 +246,7 @@ class TransitModel:
 
 	def get_t_conjunction(self, params):
 		"""
-		Return the time of primary transit center (calculated based on the time of secondary eclipse, params.t_secondary).
+		Return the time of primary transit center (calculated using `params.t_secondary`).
 		"""
 		phase = self._get_phase(params, "primary")
 		phase2 = self._get_phase(params, "secondary")

### 2015-08-07 docs/tutorial.py
add supersampling example

@@ -79,18 +79,19 @@ m.calc_err(plot = True)
 #m = batman.TransitModel(params, t, nthreads = 4)"""
 
 params.fp = 0.001
-t = np.linspace(-0.52, -0.48, 1000)
+params.t_secondary = 0.5
+t = np.linspace(0.48, 0.52, 1000)
 m = batman.TransitModel(params, t, transittype="secondary")	       
 flux = m.light_curve(params)
 
 plt.plot(t, flux)
-plt.ylim((0.9989, 1.0001))
-plt.xlim((-0.52, -0.48))
+plt.ylim((0.9995, 1.0015))
+plt.xlim((0.48, 0.52))
+plt.gca().set_yticklabels([0.9995, 1.000, 1.0005, 1.001, 1.0015])
 plt.xlabel("Time (days)") 
 plt.ylabel("Relative flux")
 plt.savefig("eclipse.png")
-#plt.show()
+plt.show()
 
-t_secondary = m.get_t_secondary(params)
-print(t_secondary)
 
+m = batman.TransitModel(params, t, supersample_factor = 7, exp_time = 0.001)

### 2015-08-07 batman/transitmodel.py
add supersampling

@@ -56,12 +56,18 @@ class TransitModel:
 	:param transittype: Type of transit ("primary" or "secondary")
 	:type transittype: string, optional
 
+	:param supersample_factor:	Number of points subdividing exposure
+	:type supersample_factor: integer, optional
+
+	:param exp_time: Exposure time (in same units as `t`)
+	:type exp_time: double, optional
+
 	:Example:
 	
 	>>> m = batman.TransitModel(params, max_err = 0.5, nthreads=4)
 	"""
 
-	def __init__(self, params, t, max_err=1.0, nthreads = 1, fac = None, transittype = "primary"):
+	def __init__(self, params, t, max_err=1.0, nthreads = 1, fac = None, transittype = "primary", supersample_factor = 1, exp_time = 0.):
 		#checking for invalid input
 		if  (params.limb_dark == "uniform" and len(params.u) != 0) or (params.limb_dark == "linear" and len(params.u) != 1) or \
 		    (params.limb_dark == "quadratic" and len(params.u) != 2) or (params.limb_dark == "logarithmic" and len(params.u) != 2) or \
@@ -89,16 +95,27 @@ class TransitModel:
 		self.ecc = params.ecc
 		self.w = params.w
 		self.u = params.u
-		self.max_err = max_err
 		self.limb_dark = params.limb_dark
 		self.fp = params.fp
+		self.max_err = max_err
+		
+		self.supersample_factor = supersample_factor
+		self.exp_time = exp_time
+		if self.supersample_factor > 1: 
+			ts = []
+			for i in range(len(self.t)): ts.append(np.linspace(self.t[i] - self.exp_time/2., self.t[i] + self.exp_time/2., self.supersample_factor))
+			self.t_supersample = np.array(ts).flatten()
+		else: self.t_supersample = self.t
+		
 		if transittype == "primary": self.transittype = 1
 		else: 
 			self.transittype = 2
 			params.t0 = self.get_t_conjunction(params)
-		self.ds= _rsky._rsky(t, params.t0, params.per, params.a, params.inc*pi/180., params.ecc, params.w*pi/180., self.transittype)
+		self.ds = _rsky._rsky(self.t_supersample, params.t0, params.per, params.a, params.inc*pi/180., params.ecc, params.w*pi/180., self.transittype)
+		
 		if fac != None: self.fac = fac
 		else: self.fac = self._get_fac()
+		
[…]

### 2015-08-07 batman/transitmodel.py
better docstrings

@@ -309,6 +309,9 @@ class TransitParams(object):
 	:param fp: Planet-to-star flux ratio (for secondary eclipse models).
 	:type fp: float, optional
 
+	.. note::  
+		- Units for the orbital period and ephemeris can be anything as long as they are consistent (e.g. both in days). 
+		- The orbital path is calculated based on `t0` for primary transits and `t_secondary` for secondary eclipses.
 
 	:Example:
 	
@@ -323,9 +326,6 @@ class TransitParams(object):
 	>>> params.w = 90.				#longitude of periastron (in degrees) 
 	>>> params.u = [0.1, 0.3] 	      	        #limb darkening coefficients
 	>>> params.limb_dark = "quadratic"          	#limb darkening model
-
-	.. note::  Units for the orbital period and ephemeris can be anything as long as they are consistent (e.g. both in days). 
-
 	"""
 	def __init__(self):
 		self.t0 = None

### 2015-08-07 manuscript/ms.tex
add warning about light travel time

@@
 They have also enabled the first estimates of the occurrence of habitable planets \citep{traub12, dressing13, petigura13, kopparapu13, foremanmackey14, dressing15}.
-Transit light curves can also reveal planets' atmospheric temperature structure and composition \citep[e.g.][]{seager00, charbonneau02, lecavelier08, sing11, deming13, knutson14a, fraine14, kreidberg15}.
+Transit light curves can even reveal planets' atmospheric temperature structure and composition \citep[e.g.][]{seager00, charbonneau02, lecavelier08, sing11, deming13, knutson14a, fraine14, kreidberg15}.
 A number of current and planned observational facilities -- including \emph{K2}, \emph{TESS}, \emph{CHEOPS}, \emph{JWST}, and \emph{PLATO} -- will measure precise transit light curves for thousands of exoplanets that will further advance our understanding of planet formation, evolution, and habitability.
@@
 
-\texttt{batman} also provides utilities to calculate the time of periastron and the secondary eclipse time from the other transit parameters, using the method described in \S 3.1 of \cite{eastman13}.
+\texttt{batman} provides utilities to calculate the time of periastron, time of inferior conjunction, and the secondary eclipse time from the other transit parameters, using the method described in \S 3.1 of \cite{eastman13}.
+Note however that ``batman`` does not correct for the effects of light travel time.
+
+An additional utility is light curve supersampling, which allows the user to calculate the average value of the light curve over a specified number of evenly spaced points during an exposure.
 

### 2015-10-05 batman/__init__.py
Adding __all__ statements for py3 compatibility

@@ -1,7 +1,7 @@
 __all__ = ['transitmodel', 'tests', 'plots']
 
 
-__version__ = "0.9.0"
+__version__ = "0.9.1"
 
 from .transitmodel import *
 from .tests import *

### 2015-10-05 batman/plots.py

@@ -24,6 +24,7 @@ from . import _nonlinear_ld
 import timeit
 import batman
 
+__all__ = ['make_plots']
 
 def wrapper(func, *args, **kwargs):
     def wrapped():

### 2015-10-05 batman/tests.py

@@ -20,7 +20,9 @@ import math
 import matplotlib.pyplot as plt
 import timeit
 from .transitmodel import *
-from .openmp import *
+from .openmp import detect
+
+__all__ = ['test']
 
 def wrapper(func, *args, **kwargs):
     def wrapped():
@@ -76,7 +78,7 @@ def test():
 			failures += 1
 
 	print("\nTesting multithreading...")
-	if openmp.detect():
+	if detect():
 		params.u = [0.1, 0.3]
 		params.limb_dark = "quadratic"
 		m = TransitModel(params, t, nthreads = 1)

### 2015-10-05 batman/transitmodel.py

@@ -29,12 +29,14 @@ from math import pi
 import multiprocessing
 from . import openmp
 
+__all__ = ['TransitModel', 'TransitParams']
+
 def wrapper(func, *args, **kwargs):
     def wrapped():
         return func(*args, **kwargs)
     return wrapped
 
-class TransitModel:
+class TransitModel(object):
 	"""
 	Class for generating model transit light curves.	
 

### 2015-10-06 batman/__init__.py
update version number

@@ -1,7 +1,7 @@
 __all__ = ['transitmodel', 'tests', 'plots']
 
 
-__version__ = "0.9.1"
+__version__ = "2.1.1"
 
 from .transitmodel import *
 from .tests import *

### 2015-10-06 setup.py

@@ -74,7 +74,7 @@ _rsky = Extension('batman._rsky', ['c_src/_rsky.c'])
 _eclipse = Extension('batman._eclipse', ['c_src/_eclipse.c'])
 
 setup(	name='batman-package', 
-	version="2.1.0", 
+	version="2.1.1", 
 	author='Laura Kreidberg',
 	author_email = 'laura.kreidberg@uchicago.edu',
 	url = 'https://github.com/lkreidberg/batman',
